### 2015-09-16 iclr-paper.tex
First Commit of Paper

\documentclass{article} % For LaTeX2e
\usepackage{iclr-style/iclr2016_conference,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}

\newcommand{\sigmoid}{\boldsymbol{\sigma}}
\newcommand{\hlang}{h^{lang}}
\newcommand{\hlangall}{\boldsymbol{h^{lang}}}
\newcommand{\hdec}{h^{dec}}
\newcommand{\henc}{h^{enc}}
\newcommand{\readop}{\mathit{read}}
\newcommand{\writeop}{\mathit{write}}
\newcommand{\encoder}{\mathit{LSTM}^{enc}}
\newcommand{\decoder}{\mathit{LSTM}^{dec}}
\newcommand{\canv}{c}
\newcommand{\lat}{z}
\newcommand{\Lat}{Z}
\newcommand{\post}{Q}
\newcommand{\prior}{P}
\newcommand{\loss}{\mathcal{L}}
\newcommand{\lloss}{\mathcal{L}^{z}}
\newcommand{\rloss}{\mathcal{L}^{x}}

\title{Generating Images From Captions\\ With Attention}

\author{
Elman Mansimov\textsuperscript{*}, Emilio Parisotto\thanks{Equal Contribution} , Jimmy Lei Ba \& Ruslan Salakhutdinov\\
Department of Computer Science\\
University of Toronto\\
Toronto, Ontario, Canada \\
\texttt{\{emansim,eparisotto,rsalakhu\}@cs.toronto.edu}, \texttt{jimmy@psi.utoronto.ca}
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\begin{document}

\maketitle

\begin{abstract}
At the end.
\end{abstract}

\section{Introduction}

Generating realistic images from their descriptions is a very hard task that combines together two challenging problems of language modelling and image generation. Generating images from captions is a significantly harder problem than a reverse problem of generating captions from images which has recenly attracted a lot of attention. The model has to capture the semantic knowledge expressed in the description and then use it to generate the values of RGB channels of each pixel in the image ...

\section{Related Work}

Deep Neural Networks have achieved a remarkable performance in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech} and etc. While most of the recent success has been achieved by discriminative models, the generative models have not yet enjoyed the same level of success. Most of the previous work in generative models has been focused on variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}. While these models are very powerful, each iteration of training requires a computationally costly step of MCMC to approximate an intractable normalization constant that makes it hard to scale them to large datasets.

\cite{kingma_vae} have introduced the Variational Auto-Encoder (VAE) which can be seen as a neural network with continous latent variables. The encoder is used to approximate posterior distribution and the decoder is used to stochastically reconstruct the data from latent variables. The model performs an efficient inference and learning that allows it to scale to large datasets. \cite{gregor_draw} have introduced Deep Recurrent Attention Writer (DRAW), where they have incorporated a novel differentiable attention mechanism into the VAE which significantly improved its performance as well as quality of generated samples. While most of the samples of VAE and DRAW resemble a clear structure of objects, the generated images are blurry most of the time.

Generative Adversarial Networks (GANs) \citep{goodfellow_gan} are another type of generative models that use noise-contrastive estimation \citep{gutmann_nce} to avoid calculating intractable normalization constant. The model consists of generator that generates samples from uniform distribution and discriminator that discriminates between real and generated images. Both networks are playing the game, where generator tries to produce samples that look real and discriminator tries not to be fooled by generator. Recently, \cite{denton_lapgan} have scaled those models, by training GANs at each level of Laplacian pyramid of images. While their model has generated sharp looking samples, the generated images have lacked a clear structure. Compared to other mentioned generative models, GANs are more unstable and are harder to train.

While all of the previous work has been focused on unconditional models or models conditioned on labels, to the best of our knowledge this paper is the first to introduce the generative model of images conditioned on captions.

\section{Model}

Our proposed model can be seen as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$. Let $y$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $x$ be the image corresponding to that caption.

\subsection{Encoder}

The encoder is a deterministic Bidirectional LSTM that encodes the variable size sentences into the vector representation $s$. Bidirectional LSTM consists of one Forward LSTM and Backward LSTM which combine information from past and future respectively. The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$. Then these hidden states are concatenated together into the sequence $[\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.

The final representation of the sentence is calculated as follows:
\begin{align}
s_{t} = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}
\end{align}
where $\hlang_{0}$ is initialized to the learned bias.
Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. We will describe how $\alpha_{1...N}$ are calculated in the next section.

\subsection{Decoder}

The decoder is a DRAW model, which is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$. Formally, the decoder iteratively computes the following equations for $t=1,...,T$
\begin{align}
\label{eq:x_hat}
\hat{x}_t &= x-\sigmoid(\canv_{t-1})\\
\label{eq:read}
r_t &= \readop(x_t, \hat{x}_t, \hdec_{t-1})\\
\henc_t &= \encoder(\henc_{t-1}, [r_t, \hdec_{t-1}])\\
\lat_t &\sim \post(\Lat_t|\henc_t)\\
\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1})\\
s_{t} &= align(\hdec_{t-1}, \hlangall)\\
\label{eq:write}
\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
\end{align}

where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{n}]$ and $\canv_{0}, \hdec_{0}, \henc_{0}$ are initialized to the learned biases.

The $align$ function introduced by \cite{bahdanau_mt} is used to compute probabilities $\alpha_{1...n}$ by first computing the energies $e_{t,0}, e_{t,1}, ..., e_{t,n}$ and then rescaling them to the probability distribution $\alpha_{1}, \alpha_{2}, ..., \alpha_{n}$
\begin{align}
e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t})\\
\alpha_{j} &= softmax(e_{tj})
\end{align}

where $softmax(e_{tj}) = \frac{exp(e_{tj})}{\sum_{j=1}^{N}e_{tj}}$

\subsection{Learning}

The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. The model is trained to maximize the marginal likelihood $\loss$ of the correct image $x$ given the input caption $y$. The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$.

The reconstruction loss $\rloss$ equals to $\frac{1}{L}\sum_{l=1}^{L}(log\,p(x_{t}|y,z)$ where $L$ is the number of samples used during training, which was set to $1$ in our experiments.

The latent loss is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term. Since the patches drawn on canvas over time are not independent of each other, naturally the sufficient statistics of the prior distribution at time $t$ should be dependent on the sufficient statistics of the prior distribution at time $t-1$. Therefore, instead of setting $\prior(\Lat_1), ..., \prior(\Lat_T)$ to be independent unit gaussian distributions, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$, as in \citep{bachman_sdm}, where 
\begin{align}
\mu_{t}^{prior} &= tanh(W_{mu}\hdec_{t-1})\\
\sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
\end{align}
Overall, the likelihood $\loss$ is calculated as follows:
\begin{align}
\loss &= -\sum_{t=1}^{T}D_{KL}(\post(\Lat_t|\henc_t,s_{t-1})\,||\,\prior(\Lat_t)) + \frac{1}{L}\sum_{l=1}^{L}log\,p(x_{t}|y,z)\\
&=
\frac{1}{2}\sum_{t=1}^{T}(1 - 2\,log\,\sigma_{t}^{prior} + 2\,log\,\sigma_{t} - \frac{exp(2\,log\,\sigma_{t}) + (\mu_{t} - \mu_{t}^{prior})^{2}}{exp(2\,log\,\sigma_{t}^{prior})}) + \frac{1}{L}\sum_{l=1}^{L}log\,p(x_{t}|y,z)
\end{align}

\subsection{Image Generation}

During the image generation step, we throw away the Inference LSTM from the Decoder and instead sample from the prior distribution. Due to the samples from DRAW model being blurry, we do an additional post processing step, where we use adversarial net ...

%\section{Model}

%Our proposed model can be seen as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time. The encoder is a RNNSearch model \citep{bahdanau_mt}, which is a deterministic recurrent neural network that encodes the variable sized sentence into the vector representation. The decoder is a DRAW model, which is a stochastic recurrent neural network that consists of inference network that infers the distribution of latent variables of images and then the generator network uses the inferred latent variables in order to reconstruct the images. At the test time, the images are sampled from rr distribution given the vector representation of the caption.

%A soft attention mechanism in the encoder combines the hidden states of the rnn as a weighted sum, where the weights are updated at every timestep depending on the hidden states of the encoder and the hidden states of the generator network in the decoder. The dynamically updated vector representation of the captions helps the model to cope with long sentences and achieve a better performance, without using tricks such as reversing the source sentences. A selective attention mechanism in the decoder restricts both the input region obverved by inference network and the output region modified by the generator network.

%The use of attention in both the encoder and the decoder allows the model to generate the parts of image while focusing the relevant representations of words in the caption.

%Mathematically, the model is trained by maximizing the marginal likelihood $L$ of a correct image $x$ given the caption $y$, consisting of words $y_{1}, y_{2}, ..., y_{n}$. Marginal likelihood is defined as 
%$L = -\sum_{t=1}^{T}D_{KL}(q(z_{t}|x,y) || p(z_{t})) + \frac{1}{s}\sum_{s=1}^{S}(logp(x|y,z)$, where $s$ is the number of samples which was $1$ during training.

%Due to generated images from DRAW being blurry, as a post processing step we have used ...

\section{Experiments}

\bibliography{iclr-paper}
\bibliographystyle{iclr-style/iclr2016_conference}

\end{document}

### 2015-09-18 iclr-paper.tex
various fixes here and there

@@
 Generating realistic images from their descriptions is a very hard task that combines together two challenging problems of language modelling and image generation.
+The model has to capture the semantic knowledge expressed in the description and then use it to generate the values of three color channels of each pixel in the image.
 Generating images from captions is a significantly harder problem than a reverse problem of generating captions from images which has recenly attracted a lot of attention.
-The model has to capture the semantic knowledge expressed in the description and then use it to generate the values of RGB channels of each pixel in the image ...
 
@@
 \begin{align}
-e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t})\\
+e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)\\
 \alpha_{j} &= softmax(e_{tj})
@@
 
-where $softmax(e_{tj}) = \frac{exp(e_{tj})}{\sum_{j=1}^{N}e_{tj}}$
+where $softmax(e_{tj}) = \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}$
 
@@
 The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}.
-The model is trained to maximize the marginal likelihood $\loss$ of the correct image $x$ given the input caption $y$.
+The model is trained to maximize the lower bound of marginal likelihood $\loss$ of the correct image $x$ given the input caption $y$.
 The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$.

### 2015-09-21 iclr-paper.tex
more changes. added figures for mnist

@@
 \usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{caption}
 
@@
 \newcommand{\rloss}{\mathcal{L}^{x}}
+
+\newenvironment{myfont}{\fontfamily{[pcr]}\selectfont}{\par}
+\DeclareTextFontCommand{\textmyfont}{\myfont}
+
+\setlength{\abovecaptionskip}{5pt} % Chosen fairly arbitrarily
+\setlength{\belowcaptionskip}{5pt} % Chosen fairly arbitrarily
 
@@
 Generating realistic images from their descriptions is a very hard task that combines together two challenging problems of language modelling and image generation.
-The model has to capture the semantic knowledge expressed in the description and then use it to generate the values of three color channels of each pixel in the image.
-Generating images from captions is a significantly harder problem than a reverse problem of generating captions from images which has recenly attracted a lot of attention.
+Effectively, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate the pixel intensities of the image.
+Generating images from captions is a significantly harder problem than a reverse problem of generating captions from images which has recenly attracted a lot of attention from machine learning research community \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
+
+By using a sequence to sequence framework to approach the problem of image generation from captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
 
@@
 
-During the image generation step, we throw away the Inference LSTM from the Decoder and instead sample from the prior distribution.
-Due to the samples from DRAW model being blurry, we do an additional post processing step, where we use adversarial net ...
-
-%\section{Model}
-
-%Our proposed model can be seen as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time.
-The encoder is a RNNSearch model \citep{bahdanau_mt}, which is a deterministic recurrent neural network that encodes the variable sized sentence into the vector representation.
-The decoder is a DRAW model, which is a stochastic recurrent neural network that consists of inference network that infers the distribution of latent variables of images and then the generator network uses the inferred latent variables in order to reconstruct the images.
-At the test time, the images are sampled from rr distribution given the vector representation of the caption.
-
-%A soft attention mechanism in the encoder combines the hidden states of the rnn as a weighted sum, where the weights are updated at every timestep depending on the hidden states of the encoder and the hidden states of the generator network in the decoder.
-The dynamically updated vector representation of the captions helps the model to cope with long sentences and achieve a better performance, without using tricks such as reversing the source sentences.
-A selective attention mechanism in the decoder restricts both the input region obverved by inference network and the output region modified by the generator network.
-
-%The use of attention in both the encoder and the decoder allows the model to generate the parts of image while focusing the relevant representations of words in the caption.
-
-%Mathematically, the model is trained by maximizing the marginal likelihood $L$ of a correct image $x$ given the caption $y$, consisting of words $y_{1}, y_{2}, ..., y_{n}$.
-Marginal likelihood is defined as
-%$L = -\sum_{t=1}^{T}D_{KL}(q(z_{t}|x,y) || p(z_{t})) + \frac{1}{s}\sum_{s=1}^{S}(logp(x|y,z)$, where $s$ is the number of samples which was $1$ during training.
-
-%Due to generated images from DRAW being blurry, as a post processing step we have used ...
+During the image generation step, we throw away the inference network from the decoder and instead sample from the prior distribution.
+Due to the blurriness of samples generated by DRAW model, we do an additional post processing step, where we use the generator of the adversarial network trained on residuals of laplacian pyramid to sharpen the generated images, similar to \citep{denton_lapgan}.
+By fixing the prior of adversarial generator to the mean of uniform distribution, it gets treated as a deterministic neural network which allows us to calculate the lower bound of likelihood.
+The reconstuction loss becomes the loss between sharpened image and correct image, whereas the latent loss stays the same.
+We also noticed that sampling from the mean of uniform distribution allowed us to generate much less noisy samples than by sampling from the uniform distribution itself.
 
 \section{Experiments}
+\subsection{MNIST With Captions}
+As a toy experiment, we have trained our model on the two digit MNIST dataset with captions constructed on the fly.
+Two random digits were either placed horizontally or vertically, so that they don't overlap, on the black $60 \times 60$ background.
+The caption indicated the way digits were placed in the image and had the following template: \textit{the digit (1) is (2) the (3) of the digit (4) \textless eos\textgreater}.
+(1) and (4) were the digit numbers, (2) was either on or at, (3) was one of the adjectives top, bottom, left or right.
+For example, the generated images corresponding to the caption \textit{the digit seven is at the bottom of the digit five \textless eos\textgreater} as well as other captions are shown on~\ref{fig:fig1}.
+While most of the generative models were trained on the version of binarized MNIST dataset \citep{russ_binarizedmnist}, our model was trained directly on pixel intensities with binary cross entropy cost function.
+
+\subsection{COCO}
+TODO:
+1) show table of SSI and PR curve
+2) show samples unfolded over time as well as the words the model is attending to (both success and failure cases)
+3) show generated samples with colour flipped, background changed etc.
+4) compare samples with LAPGAN, VAE etc.
+5) show tsne plot of types of images and well as topic of samples where the model works well where works badly
+
+\subsection{CIFAR}
+
+\begin{figure}[t]
+\label{fig1}
+\captionsetup{labelformat=empty}
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/5-7-4.png}
+  \caption{\textit{is digit 5}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/5-7-6.png}
+  \caption{\textit{is digit 5}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/5-7-9.png}
+  \caption{\textit{is digit 5}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/5-7-13.png}
+  \caption{\textit{is digit 5}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/5-7-17.png}
+  \caption{\mbox{\textit{eos 7 digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/5-7-22.png}
+  \caption{\mbox{\textit{eos 7 digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/5-7-26.png}
+  \caption{\mbox{\textit{eos 7 digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/5-7-27.png}
+  \caption{\mbox{\textit{eos 7 digit}}}
+\endminipage\hfill
+
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/9-2-4.png}
+  \caption{\mbox{\textit{eos digit 2}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/9-2-6.png}
+  \caption{\mbox{\textit{eos digit 2}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/9-2-9.png}
+  \caption{\mbox{\textit{eos digit 2}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/9-2-13.png}
+  \caption{\mbox{\textit{eos digit 2}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/9-2-17.png}
+  \caption{\mbox{\textit{of top digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/9-2-22.png}
+  \caption{\mbox{\textit{of top digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/9-2-26.png}
+  \caption{\mbox{\textit{of top digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/9-2-27.png}
+  \caption{\mbox{\textit{of top digit}}}
+\endminipage\hfill
+
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/1-0-4.png}
+  \caption{\mbox{\textit{eos 0 of}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/1-0-6.png}
+  \caption{\mbox{\textit{eos digit 0}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/1-0-9.png}
+  \caption{\mbox{\textit{eos digit 0}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/1-0-12.png}
+  \caption{\mbox{\textit{eos 0 of}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/1-0-17.png}
+  \caption{\mbox{\textit{1 digit the}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/1-0-22.png}
+  \caption{\mbox{\textit{1 digit the}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/1-0-25.png}
+  \caption{\mbox{\textit{1 digit on}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/1-0-28.png}
+  \caption{\mbox{\textit{digit the eos}}}
+\endminipage\hfill
+
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/4-8-4.png}
+  \caption{\textit{left 8 4}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/4-8-6.png}
+  \caption{\textit{left 8 4}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/4-8-9.png}
+  \caption{\mbox{\textit{left digit 4}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/4-8-14.png}
+  \caption{\mbox{\textit{eos 8 digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/4-8-18.png}
+  \caption{\mbox{\textit{eos 8 digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/4-8-22.png}
+  \caption{\mbox{\textit{8 digit the}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/4-8-25.png}
+  \caption{\mbox{\textit{8 digit digit}}}
+\endminipage\hfill
+\minipage{0.1\textwidth}
+  \includegraphics[width=\linewidth]{figures/4-8-28.png}
+  \caption{\mbox{\textit{digit digit the}}}
+\endminipage\hfill
+
+\captionsetup{labelformat=empty}
+\caption{Figure 1: Four examples show the generated images unfolded over several timesteps as well as the top three words the model attends to while generating images.}
+\end{figure}
 

### 2015-09-28 iclr-paper.tex
Made many changes, added lots of pictures. Still have a lot to do

@@
 \usepackage{caption}
+%\usepackage{subcaption}
+\usepackage{subfig}
 
@@
 \author{
-Elman Mansimov\textsuperscript{*}, Emilio Parisotto\thanks{Equal Contribution} , Jimmy Lei Ba \& Ruslan Salakhutdinov\\
+Elman Mansimov, Emilio Parisotto, Jimmy Lei Ba \& Ruslan Salakhutdinov\\
 Department of Computer Science\\
@@
 While most of the generative models were trained on the version of binarized MNIST dataset \citep{russ_binarizedmnist}, our model was trained directly on pixel intensities with binary cross entropy cost function.
-
-\subsection{COCO}
-TODO:
-1) show table of SSI and PR curve
-2) show samples unfolded over time as well as the words the model is attending to (both success and failure cases)
-3) show generated samples with colour flipped, background changed etc.
-4) compare samples with LAPGAN, VAE etc.
-5) show tsne plot of types of images and well as topic of samples where the model works well where works badly
-
-\subsection{CIFAR}
 
@@
 
+\subsection{COCO}
+
+Microsoft COCO \citep{mscoco} is a largest dataset of images annotated with captions consisting of 83k images.
+The rich collection of images with variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on caption very challenging.
+Since some of the images have more than five captions attached to them, for consistency with related work on caption generation we disregard extra captions.
+
+In the following subsections we make both qualitative and quantitative analysis of our model as well as compare its performance with other related generative models.
+
+\subsubsection{Analysis of Generated Images}
+The main goal of this work is to learn a model that can understand the semantic meaning expressed in description of the image, such as the properties of objects, the relationships between them, etc.
+and then use that knowledge to generate relevant images.
+To verify that, we generated a set of captions inspired by COCO dataset and changed some words in the captions to see whether the model made the relevant changed in the generated samples.
+
+First, we wanted to see whether the model understood one of the simplest properties of any object, the color.
+We generated images of school buses with four different colors: yellow, red, green and blue.
+Although, there are images of buses with different colors in the training set, all school buses specifically are colored yellow.
+As you can see below, the model managed to generate images with an object that looked like a bus with the correct color.
+
+\vspace{-0.3cm}
+\begin{figure}[h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[A \underline{yellow} school bus parked in a parking lot.
+]{\includegraphics[width=0.22\textwidth]{figures/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[A \underline{red} school bus parked in a parking lot.
+]{\includegraphics[width=0.22\textwidth]{figures/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[A \underline{green} school bus parked in a parking lot.
+]{\includegraphics[width=0.22\textwidth]{figures/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[A \underline{blue} school bus parked in a parking lot.
+]{\includegraphics[width=0.22\textwidth]{figures/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\end{center}
+%\caption{whole image caption}
+\label{fig:label}
+\vspace{-0.2cm}
+\end{figure}
+
+Apart from changing the colors of objects, we were curious whether changing the background of the scene described in the caption would result in the appropriate changes in the generated samples.
+It is a somewhat harder task for a model than changing the color, due to larger number of changes that have to be made in the generated samples.
+Nevertheless as shown below, changing the grass type from dry to green as well as adding clouds in the caption has resulted in the appropriate changes.
+
+\vspace{-0.3cm}
+\begin{figure}[h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[A very large commercial plane flying in the \underline{blue} skies.
+]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-blue-skies-sharp.png}}\quad
+%
+\subfloat[A very large commercial plane flying in the \underline{cloudy} skies.
+]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-cloudy-skies-sharp.png}}\quad
+%
+%
+\end{center}
+%\caption{whole image caption}
+\label{fig:label}
+\vspace{-0.8cm}
+\end{figure}
+
+\begin{figure}[h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[A heard of elephants walking across a \underline{dry} grass field.
+]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
+%
+\subfloat[A heard of elephants walking across a \underline{green} grass field.
+]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
+%
+%
+\end{center}
+%\caption{whole image caption}
+\label{fig:label}
+\vspace{-0.2cm}
+\end{figure}
+
+Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we have found that the model made appropriate changes as long as some similar patter was present in the training set.
+However, it wasn't always the case when changing an object itself in the description.
+In cases, when objects did't have a noticeable differences in their properties, such as shape or color, the changes in the generated samples weren't very clearly seen.
+This could be attributed to the limitation of the model to grasp the detailed understanding of each object.
+
+\vspace{-0.3cm}
+\begin{figure}[h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[The \underline{decadent chocolate dessert} is on the table.
+]{\includegraphics[width=0.48\textwidth]{figures/the-decadent-chocolate-dessert-is-on-the-table-2-sharp.png}}\quad
+%
+\subfloat[The \underline{bowl of banas} is on the table.
+]{\includegraphics[width=0.48\textwidth]{figures/a-bowl-of-bananas-is-on-the-table-2-sharp.png}}\quad
+%
+%
+\end{center}
+%\caption{whole image caption}
+\label{fig:label}
+\vspace{-0.8cm}
+\end{figure}
+
+\begin{figure}[!h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[A vintage photo of a \underline{cat}.
+]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-cat-sharp.png}}\quad
+%
+\subfloat[A vintage photo of a \underline{dog}.
+]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-dog-sharp.png}}\quad
+%
+%
+\end{center}
+%\caption{whole image caption}
+\label{fig:labelr}
+\vspace{-0.2cm}
+\end{figure}
+
+\subsubsection{Analysis of Attention}
+
+To be added soon.
+
+\subsubsection{Comparison With Other Models}
+The quantitative evaluation of generative models has been a subject of ambiguity within a machine learning community.
+Compared to reporting classification accuracies, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
+To avoid ambiguity, we report results on two different metrics as well as do qualitative comparison of generative models.
+
+As shown on Figure blah, we have sampled the images from the same caption in four of the best models.
+
+First, to compare the the peformance of our model with other generative models trained with variational objective, we report the Recall$@$K, namely the mean number of sentences for which the correct image is ranked within the top-K retrieved results, based on the lower bound of likelihood $\loss$.
+If you want to get a really good results build a model like Ryan did.
+Do deal away with easy images we take ratio with ...
+with mean caption.
+To deal with large computational complexity of looping over each image we only consider closest neightbors.
+The precision Recall curve is reported there...
+
+Second,
+
+\vspace{-0.3cm}
+\begin{figure}[!t]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat{\includegraphics[width=0.22\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp.png}}\quad
+%
+\subfloat{\includegraphics[width=0.22\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan.png}}\quad
+%
+\subfloat{\includegraphics[width=0.22\textwidth]{to-do.png}}\quad
+%
+\subfloat{\includegraphics[width=0.22\textwidth]{to-do.png}}\quad
+%
+\end{center}
+\caption{Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
+\label{fig:label}
+\vspace{-0.2cm}
+\end{figure}
+
+\subsection{CIFAR}
+
 \bibliography{iclr-paper}

### 2015-10-05 iclr-paper.tex
update intro

@@
 
-Generating realistic images from their descriptions is a very hard task that combines together two challenging problems of language modelling and image generation.
-Effectively, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate the pixel intensities of the image.
-Generating images from captions is a significantly harder problem than a reverse problem of generating captions from images which has recenly attracted a lot of attention from machine learning research community \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
+Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
+This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components.
+Image generative models studied previously are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
+Despite the advances in generative models, learning the highly structured natural image distribution alone in the high dimensional pixel space alone proves to be a difficult task.
+In real world, however, images rarely appear in isolation.
+They often accompanied by their unstructured textual descriptions on web pages and in books.
+One domain presents substantial amount of relating information of the other domain.
+The additional information from image and unstructured text description could be used to simplify the image modelling task.
 
-By using a sequence to sequence framework to approach the problem of image generation from captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
+There are two directions in learning a generative model of image and text.
+One approach is to learn a text generative model conditioned on the images.
+Significant amount of recent works have been focusing on generating captions from images\citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
+The models take an image descriptor and generate unstructured texts through a recurrent decoder.
+By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description.
+Generating high dimensional realistic images from their descriptions is a more difficult approach that combines two challenging components of language modelling and image generation.
+Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image.
+Although, the interesting high dimensional natural images lay on a small manifold that is difficult to capture, the additional text description cues of a target image may simplify the learning problem by focusing on the conditional distribution.
+
+In this paper, we illustrate how simple sequential deep learning techniques can be used to build a conditional probabilistic model over natural image space effectively.
+By using a sequence to sequence framework to approach the problem of image generation from unstructured natural language captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
 

### 2015-10-05 iclr-paper.tex
update

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{graphicx}
@@
 
+\newcommand{\comm}[1]{}
+\newcommand{\given}{\,|\,}
+\newcommand{\expectation}{\mathbb{E}}
+\newcommand{\kldiv}{\mathrm{D}_{\rm KL}}
+\newcommand{\klBars}{\,\|\,}
 \newcommand{\sigmoid}{\boldsymbol{\sigma}}
@@
 \newcommand{\Lat}{Z}
+\newcommand{\numSamples}{L}
+\newcommand{\sampleIdx}{l}
+\newcommand{\LatSample}{\tilde{Z}}
+\newcommand{\icaption}{{\bf{y}}}
+\newcommand{\oimage}{{\bf{x}}}
 \newcommand{\post}{Q}
@@
 By using a sequence to sequence framework to approach the problem of image generation from unstructured natural language captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
+Overall, the main contribution of this work are the following:
 
@@
 \section{Model}
+Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $\oimage$ be the image corresponding to that caption.
 
-Our proposed model can be seen as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
-Let $y$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $x$ be the image corresponding to that caption.
+\subsection{Modelling Language: the Bidirectional Attention RNN}
 
-\subsection{Encoder}
-
-The encoder is a deterministic Bidirectional LSTM that encodes the variable size sentences into the vector representation $s$.
+The input caption sentences are fed into a deterministic Bidirectional LSTM\cite{} that encodes the variable size sentences into the vector representation $s$.
 Bidirectional LSTM consists of one Forward LSTM and Backward LSTM which combine information from past and future respectively.
@@
 
-\subsection{Decoder}
+\subsection{Modelling Images: the Conditional DRAW Network}
 
-The decoder is a DRAW model, which is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
+The image generative model is a DRAW model, which is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
 Formally, the decoder iteratively computes the following equations for $t=1,...,T$
@@
 Since the patches drawn on canvas over time are not independent of each other, naturally the sufficient statistics of the prior distribution at time $t$ should be dependent on the sufficient statistics of the prior distribution at time $t-1$.
-Therefore, instead of setting $\prior(\Lat_1), ..., \prior(\Lat_T)$ to be independent unit gaussian distributions, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$, as in \citep{bachman_sdm}, where
+Therefore, instead of setting $\prior(\Lat_1), ..., \prior(\Lat_T)$ to be independent unit gaussian distributions, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$, which forms a Markov chain $\prior(\Lat_1), \prior(\Lat_2\given\Lat_1), ..., \prior(\Lat_T\given\Lat_{T-1})$ as in \citep{bachman_sdm}, where
 \begin{align}
-\mu_{t}^{prior} &= tanh(W_{mu}\hdec_{t-1})\\
+\mu_{t}^{prior} &= tanh(W_{\mu}\hdec_{t-1})\\
 \sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
 \end{align}
-Overall, the likelihood $\loss$ is calculated as follows:
+Overall, the loss function $\loss$ is calculated as follows:
+\begin{align}
+\loss =  &\expectation_{Q(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
+\end{align}
+The expectation can be approximate by $\numSamples$ Monte Carlo samples $\LatSample_{1:T}$ from $\post(\Lat_{1:T}\given\icaption, \oimage)$:
+\begin{align}
+\loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ - \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
+\end{align}
+\comm{
 \begin{align}
@@
 \end{align}
+}
 
-\subsection{Image Generation}
+\subsection{Generating Images from Captions}
 

### 2015-10-05 iclr-paper.tex
update
update equation

@@
 By using a sequence to sequence framework to approach the problem of image generation from unstructured natural language captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
-Overall, the main contribution of this work are the following:
+Overall, the main contributions of this work are the following:
 
@@
 
-\subsection{Modelling Language: the Bidirectional Attention RNN}
+\subsection{Language Model: the Bidirectional Attention RNN}
 
@@
 
-\subsection{Modelling Images: the Conditional DRAW Network}
+\subsection{Image Model: the Conditional DRAW Network}
 
 The image generative model is a DRAW model, which is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
+We extend the DRAW network to include the $align$ function by \cite{bahdanau_mt} which used to compute the alignment between the input caption and intermediate image generative steps:
+
 Formally, the decoder iteratively computes the following equations for $t=1,...,T$
@@
 \hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1})\\
-s_{t} &= align(\hdec_{t-1}, \hlangall)\\
+s_{t} &= align(\hdec_{t-1}, \hlangall) \\
 \label{eq:write}
@@
 
-The $align$ function introduced by \cite{bahdanau_mt} is used to compute probabilities $\alpha_{1...n}$ by first computing the energies $e_{t,0}, e_{t,1}, ..., e_{t,n}$ and then rescaling them to the probability distribution $\alpha_{1}, \alpha_{2}, ..., \alpha_{n}$
+The cooresponding alignment probabilities $\alpha_{1...n}$ at each step are obtained by:
 \begin{align}
 e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)\\
-\alpha_{j} &= softmax(e_{tj})
+\alpha_{j} &= \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}.
 \end{align}
-
-where $softmax(e_{tj}) = \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}$
 

### 2015-10-06 iclr-paper.tex
update model sec

@@
 \subsection{Language Model: the Bidirectional Attention RNN}
-
+\label{sec:lang}
 The input caption sentences are fed into a deterministic Bidirectional LSTM\cite{} that encodes the variable size sentences into the vector representation $s$.
@@
 
-The final representation of the sentence is calculated as follows:
-\begin{align}
-s_{t} = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}
-\end{align}
-where $\hlang_{0}$ is initialized to the learned bias.
-Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention.
-We will describe how $\alpha_{1...N}$ are calculated in the next section.
-
 \subsection{Image Model: the Conditional DRAW Network}
 
-The image generative model is a DRAW model, which is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
-We extend the DRAW network to include the $align$ function by \cite{bahdanau_mt} which used to compute the alignment between the input caption and intermediate image generative steps:
+The DRAW network \cite{gregor_draw} is a sequential probabilistic model generating images by accumulating the output at each iterative step.
+While the original DRAW network assumes the latent variables are independent, it has shown in \citep{bachman_sdm} the model performance is improved by including the dependencies of latent variables.
 
-Formally, the decoder iteratively computes the following equations for $t=1,...,T$
+We extended the architecture the DRAW network generative process to include additional input caption from the language model described in Sec.
+(\ref{sec:lang}).
+Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
+The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}:
+
+Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
 \begin{align}
@@
 
-where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{n}]$ and $\canv_{0}, \hdec_{0}, \henc_{0}$ are initialized to the learned biases.
+where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
+Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $algin$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
+\begin{align}
+s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.
+\end{align}
 
-The cooresponding alignment probabilities $\alpha_{1...n}$ at each step are obtained by:
+The corresponding alignment probabilities $\alpha_{1...n}$ at each step are obtained by:
 \begin{align}
@@
 \end{align}
+
+Here $\hlang_{0}$ is initialized to the learned bias.
+Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention.
 
@@
 
-The latent loss is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term.
+The latent loss $\lloss$ is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term.
 Since the patches drawn on canvas over time are not independent of each other, naturally the sufficient statistics of the prior distribution at time $t$ should be dependent on the sufficient statistics of the prior distribution at time $t-1$.

### 2015-10-07 figures/iclr-paper.tex
merged jimmys and elmans changes

--- 2015-10-05 iclr-paper.tex
@@
 \usepackage{amsmath}
-\usepackage{amsfonts}
 \usepackage{graphicx}
@@
 \usepackage{subfig}
+\usepackage{tabulary}
+\usepackage{soul}
+\usepackage{color}
+\usepackage{xcolor}
+%\usepackage{floatrow}
+% Table float box with bottom caption, box width adjusted to content
+%\newfloatcommand{capbtabbox}{table}[][\FBwidth]
 
-\newcommand{\comm}[1]{}
-\newcommand{\given}{\,|\,}
-\newcommand{\expectation}{\mathbb{E}}
-\newcommand{\kldiv}{\mathrm{D}_{\rm KL}}
-\newcommand{\klBars}{\,\|\,}
 \newcommand{\sigmoid}{\boldsymbol{\sigma}}
@@
 \newcommand{\Lat}{Z}
-\newcommand{\numSamples}{L}
-\newcommand{\sampleIdx}{l}
-\newcommand{\LatSample}{\tilde{Z}}
-\newcommand{\icaption}{{\bf{y}}}
-\newcommand{\oimage}{{\bf{x}}}
 \newcommand{\post}{Q}
@@
 \newcommand{\rloss}{\mathcal{L}^{x}}
+\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
+\definecolor{0}{RGB}{247,251,255}
+\definecolor{1}{RGB}{222,235,247}
+\definecolor{2}{RGB}{198,219,239}
+\definecolor{3}{RGB}{158,202,225}
+\definecolor{4}{RGB}{107,174,214}
+\definecolor{5}{RGB}{66,146,198}
 
@@
 
-Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
-This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components.
-Image generative models studied previously are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
-Despite the advances in generative models, learning the highly structured natural image distribution alone in the high dimensional pixel space alone proves to be a difficult task.
-In real world, however, images rarely appear in isolation.
-They often accompanied by their unstructured textual descriptions on web pages and in books.
-One domain presents substantial amount of relating information of the other domain.
-The additional information from image and unstructured text description could be used to simplify the image modelling task.
+Generating realistic images from their descriptions is a very hard task that combines together two challenging problems of language modelling and image generation.
+Effectively, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate the pixel intensities of the image.
+Generating images from captions is a significantly harder problem than a reverse problem of generating captions from images which has recenly attracted a lot of attention from machine learning research community \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
 
-There are two directions in learning a generative model of image and text.
-One approach is to learn a text generative model conditioned on the images.
-Significant amount of recent works have been focusing on generating captions from images\citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
-The models take an image descriptor and generate unstructured texts through a recurrent decoder.
-By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description.
-Generating high dimensional realistic images from their descriptions is a more difficult approach that combines two challenging components of language modelling and image generation.
-Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image.
-Although, the interesting high dimensional natural images lay on a small manifold that is difficult to capture, the additional text description cues of a target image may simplify the learning problem by focusing on the conditional distribution.
-
-In this paper, we illustrate how simple sequential deep learning techniques can be used to build a conditional probabilistic model over natural image space effectively.
-By using a sequence to sequence framework to approach the problem of image generation from unstructured natural language captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
-Overall, the main contribution of this work are the following:
+By using a sequence to sequence framework to approach the problem of image generation from captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
 
@@
 
-Deep Neural Networks have achieved a remarkable performance in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech} and etc.
+\hlc[1]{Deep} \hlc[2]{Neural} \hlc[3]{Networks} \hlc[4]{have} \hlc[0]{achieved} a remarkable performance in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech} and etc.
 While most of the recent success has been achieved by discriminative models, the generative models have not yet enjoyed the same level of success.
@@
 \section{Model}
-Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
-Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $\oimage$ be the image corresponding to that caption.
 
-\subsection{Modelling Language: the Bidirectional Attention RNN}
+Our proposed model can be seen as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+Let $y$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $x$ be the image corresponding to that caption.
 
-The input caption sentences are fed into a deterministic Bidirectional LSTM\cite{} that encodes the variable size sentences into the vector representation $s$.
+\subsection{Encoder}
+
+The encoder is a deterministic Bidirectional LSTM that encodes the variable size sentences into the vector representation $s$.
 Bidirectional LSTM consists of one Forward LSTM and Backward LSTM which combine information from past and future respectively.
@@
 
-\subsection{Modelling Images: the Conditional DRAW Network}
+\subsection{Decoder}
 
-The image generative model is a DRAW model, which is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
+The decoder is a DRAW model, which is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
 Formally, the decoder iteratively computes the following equations for $t=1,...,T$
@@
 Since the patches drawn on canvas over time are not independent of each other, naturally the sufficient statistics of the prior distribution at time $t$ should be dependent on the sufficient statistics of the prior distribution at time $t-1$.
-Therefore, instead of setting $\prior(\Lat_1), ..., \prior(\Lat_T)$ to be independent unit gaussian distributions, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$, which forms a Markov chain $\prior(\Lat_1), \prior(\Lat_2\given\Lat_1), ..., \prior(\Lat_T\given\Lat_{T-1})$ as in \citep{bachman_sdm}, where
+Therefore, instead of setting $\prior(\Lat_1), ..., \prior(\Lat_T)$ to be independent unit gaussian distributions, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$, as in \citep{bachman_sdm}, where
 \begin{align}
-\mu_{t}^{prior} &= tanh(W_{\mu}\hdec_{t-1})\\
+\mu_{t}^{prior} &= tanh(W_{mu}\hdec_{t-1})\\
 \sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
 \end{align}
-Overall, the loss function $\loss$ is calculated as follows:
-\begin{align}
-\loss =  &\expectation_{Q(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
-\end{align}
-The expectation can be approximate by $\numSamples$ Monte Carlo samples $\LatSample_{1:T}$ from $\post(\Lat_{1:T}\given\icaption, \oimage)$:
-\begin{align}
-\loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ - \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
-\end{align}
-\comm{
+Overall, the likelihood $\loss$ is calculated as follows:
 \begin{align}
@@
 \end{align}
-}
 
-\subsection{Generating Images from Captions}
+\subsection{Image Generation}
 
@@
 \subfloat[A \underline{yellow} school bus parked in a parking lot.
-]{\includegraphics[width=0.22\textwidth]{figures/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
 %
 \subfloat[A \underline{red} school bus parked in a parking lot.
-]{\includegraphics[width=0.22\textwidth]{figures/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
 %
 \subfloat[A \underline{green} school bus parked in a parking lot.
-]{\includegraphics[width=0.22\textwidth]{figures/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
 %
 \subfloat[A \underline{blue} school bus parked in a parking lot.
-]{\includegraphics[width=0.22\textwidth]{figures/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
 %
@@
 \begin{center}
-\subfloat[A very large commercial plane flying in the \underline{blue} skies.
-]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-blue-skies-sharp.png}}\quad
+\subfloat[A very large commercial plane flying in \underline{blue} skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
 %
-\subfloat[A very large commercial plane flying in the \underline{cloudy} skies.
-]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-cloudy-skies-sharp.png}}\quad
+\subfloat[A very large commercial plane flying in \underline{rainy} skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
 %
+\subfloat[A herd of elephants walking across a \underline{dry} grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
 %
-\end{center}
-%\caption{whole image caption}
-\label{fig:label}
-\vspace{-0.8cm}
-\end{figure}
-
-\begin{figure}[h]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\subfloat[A heard of elephants walking across a \underline{dry} grass field.
-]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
-%
-\subfloat[A heard of elephants walking across a \underline{green} grass field.
-]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
-%
+\subfloat[A herd of elephants walking across a \underline{green} grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
 %
@@
 \end{figure}
+
+\vspace{-0.5cm}
+\caption{Closest images from the training set in the pixel space}
+\begin{figure}[h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\quad
+%
+\end{center}
+\label{fig:label}
+\vspace{-0.5cm}
+\end{figure}
+
+%\vspace{-0.3cm}
+%\begin{figure}[h]
+%\captionsetup[subfigure]{labelformat=empty}
+%\begin{center}
+%\subfloat[A very large commercial plane flying in the \underline{blue} skies.
+%]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-blue-skies-sharp.png}}\quad
+%
+%\subfloat[A very large commercial plane flying in the \underline{cloudy} skies.
+%]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-cloudy-skies-sharp.png}}\quad
+%
+%
+%\end{center}
+%\caption{whole image caption}
+%\label{fig:label}
+%\vspace{-0.8cm}
+%\end{figure}
+
+%\begin{figure}[h]
+%\captionsetup[subfigure]{labelformat=empty}
+%\begin{center}
+%\subfloat[A herd of elephants walking across a \underline{dry} grass field.
+%]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
+%
+%\subfloat[A herd of elephants walking across a \underline{green} grass field.
+%]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
+%
+%
+%\end{center}
+%\caption{whole image caption}
+%\label{fig:label}
+%\vspace{-0.2cm}
+%\end{figure}
 
@@
 \begin{center}
-\subfloat[The \underline{decadent chocolate dessert} is on the table.
-]{\includegraphics[width=0.48\textwidth]{figures/the-decadent-chocolate-dessert-is-on-the-table-2-sharp.png}}\quad
+\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
+]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
 %
-\subfloat[The \underline{bowl of banas} is on the table.
-]{\includegraphics[width=0.48\textwidth]{figures/a-bowl-of-bananas-is-on-the-table-2-sharp.png}}\quad
+\subfloat[\underline{A bowl of banas} is on the table.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
 %
+\subfloat[A vintage photo of a \underline{cat}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
+%
+\subfloat[A vintage photo of a \underline{dog}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
 %
@@
 \label{fig:label}
-\vspace{-0.8cm}
-\end{figure}
-
-\begin{figure}[!h]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\subfloat[A vintage photo of a \underline{cat}.
-]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-cat-sharp.png}}\quad
-%
-\subfloat[A vintage photo of a \underline{dog}.
-]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-dog-sharp.png}}\quad
-%
-%
-\end{center}
-%\caption{whole image caption}
-\label{fig:labelr}
 \vspace{-0.2cm}
@@
 
+%\vspace{-0.3cm}
+%\begin{figure}[h]
+%\captionsetup[subfigure]{labelformat=empty}
+%\begin{center}
+%\subfloat[\underline{The decadent chocolate dessert} is on the table.
+%]{\includegraphics[width=0.48\textwidth]{figures/the-decadent-chocolate-dessert-is-on-the-table-2-sharp.png}}\quad
+%
+%\subfloat[\underline{A bowl of banas} is on the table.
+%]{\includegraphics[width=0.48\textwidth]{figures/a-bowl-of-bananas-is-on-the-table-2-sharp.png}}\quad
+%
+%
+%\end{center}
+%\caption{whole image caption}
+%\label{fig:label}
+%\vspace{-0.8cm}
+%\end{figure}
+
+%\begin{figure}[!h]
+%\captionsetup[subfigure]{labelformat=empty}
+%\begin{center}
+%\subfloat[A vintage photo of a \underline{cat}.
+%]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-cat-sharp.png}}\quad
+%
+%\subfloat[A vintage photo of a \underline{dog}.
+%]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-dog-sharp.png}}\quad
+%
+%
+%\end{center}
+%\caption{whole image caption}
+%\label{fig:labelr}
+%\vspace{-0.2cm}
+%\end{figure}
+
 \subsubsection{Analysis of Attention}
 
-To be added soon.
+\subsubsection{Comparison With Other Models}
+The quantitative evaluation of generative models has been a subject of ambiguity in a machine learning community.
+Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
+To get a better comparisions between performances of generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
 
-\subsubsection{Comparison With Other Models}
-The quantitative evaluation of generative models has been a subject of ambiguity within a machine learning community.
-Compared to reporting classification accuracies, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To avoid ambiguity, we report results on two different metrics as well as do qualitative comparison of generative models.
+As you can see on Figure~\ref{fig:diffmodels}, we generated several samples using each of the current state-of-the-art generative models corresponding to the caption \textit{A group of people walk on a beach with surf boards.}.
+The samples generated by variational models were sharpened by Adversarial Network trained on edge residuals as explained in ...
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and have a less clear structure in them, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on them.
 
-As shown on Figure blah, we have sampled the images from the same caption in four of the best models.
+As for the quantitative comparison of different models we first compare the performance of our model with other generative models trained with variational objective.
+We rank the images conditioned on some caption in the test set based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model.
+As expected, the quality of image retrieval using generative models is worse compared to the models that were specifically build to do retrieval, like \cite{kiros_captions} which map captions and images to the same space.
+To deal with large computational complexity of looping through each image in the test set, we create a shortlist of 99 neighbors based on the convolutional features trained on CIFAR dataset.
+We found that the reconstruction error $\rloss$ increased for the sharpened version of generated images.
+Since the sharpening changes the color of image as well adds edges to it, computing reconstruction error for each pixel is not necessarily a good measure.
 
-First, to compare the the peformance of our model with other generative models trained with variational objective, we report the Recall$@$K, namely the mean number of sentences for which the correct image is ranked within the top-K retrieved results, based on the lower bound of likelihood $\loss$.
-If you want to get a really good results build a model like Ryan did.
-Do deal away with easy images we take ratio with ...
-with mean caption.
-To deal with large computational complexity of looping over each image we only consider closest neightbors.
-The precision Recall curve is reported there...
+Instead of calculating reconstruction error for each pixel separately, we also evaluate all models by reporting Structural Similarity Index (SSI) which incoporates perceptual phenomena, such as luminace and contrast masking into the error calculation.
+By also considering that the pixels have strong inter-dependencies when they are close to each other, the metric is calculated on the small windows of the image.
+We sampled 50 images from prior per each test caption for each generative model and calculated SSI between real image and the generated images.
+The SSI results are reported in Table~\ref{tab:ssi}
 
-Second,
+\begin{table}[!t]
+\begin{center}
+\begin{tabulary}{\linewidth}{c || c c c c c || c}
+\hline
+\multicolumn{7}{c}{\textbf{COCO}} \\
+\hline
+& \multicolumn{5}{c||}{Image Search} & Image Similarity \\
+%\cline{2-6}
+\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
+\hline
+\hline
+LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
+\hline
+Fully-conn. VAE (L2 cost) & 0.7 & 4.6 & 8.7 & 46.8 & 54 & 0.122 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv. VAE (L2 cost) & 0.7 & 4.4 & 8.5 & 46.6 & 54 & \textbf{0.131} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+Noalign DRAW & \textbf{1.0} & 6.1 & 11.2 & 52.1 & \textbf{48} & 0.114 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+Skipthought DRAW & \textbf{1.0} & \textbf{6.2} & \textbf{11.5} & \textbf{52.5} & \textbf{48} & 0.118 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+Align DRAW & 0.9 & \textbf{6.2} & 11.2 & \textbf{52.5} & \textbf{48} & 0.115 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+\end{tabulary}
+\end{center}
+\end{table}
+%\begin{tabular}{ccc}
+%\toprule
+%\textbf{Model} & \textbf{Original} & \textbf{Sharpened} \\ \hline
+%Align DRAW & 0.156 & 0.115 \\
+%Noalign DRAW & 0.155 & 0.114 \\
+%Skipthought DRAW & 0.157 & 0.118 \\
+%LAPGAN & 0.083 & - \\
+%Conv/Deconv VAE & \textbf{0.164} & \textbf{0.131}
+%\bottomrule
+%\label{tab:ssi}
+%\end{tabular}
 
@@
 \begin{center}
-\subfloat{\includegraphics[width=0.22\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp.png}}\quad
+\subfloat[Align DRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp.png}}\quad
 %
-\subfloat{\includegraphics[width=0.22\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan.png}}\quad
+\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan.png}}\quad
 %
-\subfloat{\includegraphics[width=0.22\textwidth]{to-do.png}}\quad
+\subfloat[Conv. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae.png}}\quad
 %
-\subfloat{\includegraphics[width=0.22\textwidth]{to-do.png}}\quad
+\subfloat{\includegraphics[width=0.23\textwidth]{to-do.png}}\quad
 %
@@
 \caption{Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
-\label{fig:label}
+\label{fig:diffmodels}
 \vspace{-0.2cm}
 \end{figure}
-
-\subsection{CIFAR}
 

### 2015-10-07 iclr-paper.tex
merged jimmys and elmans changes

@@
 \usepackage{subfig}
+\usepackage{tabulary}
+\usepackage{soul}
+\usepackage{color}
+\usepackage{xcolor}
+
+\captionsetup{font=small}
 
@@
 \newcommand{\rloss}{\mathcal{L}^{x}}
+\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
+
+\definecolor{0}{RGB}{247,251,255}
+\definecolor{1}{RGB}{222,235,247}
+\definecolor{2}{RGB}{198,219,239}
+\definecolor{3}{RGB}{158,202,225}
+\definecolor{4}{RGB}{107,174,214}
+\definecolor{5}{RGB}{66,146,198}
+
+\newcommand{\hlczero}[1]{\hlc[0]{#1}}
+\newcommand{\hlcone}[1]{\hlc[1]{#1}}
+\newcommand{\hlctwo}[1]{\hlc[2]{#1}}
+\newcommand{\hlcthree}[1]{\hlc[3]{#1}}
+\newcommand{\hlcfour}[1]{\hlc[4]{#1}}
+\newcommand{\hlcfive}[1]{\hlc[5]{#1}}
 
@@
 One approach is to learn a text generative model conditioned on the images.
-Significant amount of recent works have been focusing on generating captions from images\citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
+Significant amount of recent works has been focused on generating captions from images \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
 The models take an image descriptor and generate unstructured texts through a recurrent decoder.
@@
 
+\begin{figure}[!t]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\includegraphics[width=0.99\textwidth]{figures/alignDraw-cropped.pdf}\quad
+%
+\end{center}
+\caption{The image of the proposed model}
+\label{fig:figmodel}
+\vspace{-0.3cm}
+\end{figure}
+
 \section{Model}
@@
 (\ref{sec:lang}).
-Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
+Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
 The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}:
@@
 where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
-Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $algin$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
+Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
 \begin{align}
@@
 \section{Experiments}
-\subsection{MNIST With Captions}
-As a toy experiment, we have trained our model on the two digit MNIST dataset with captions constructed on the fly.
+% \subsection{MNIST With Captions}
+% As a toy experiment, we have trained our model on the two digit MNIST dataset with captions constructed on the fly.
 Two random digits were either placed horizontally or vertically, so that they don't overlap, on the black $60 \times 60$ background.
@@
 
-\begin{figure}[t]
-\label{fig1}
-\captionsetup{labelformat=empty}
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/5-7-4.png}
-  \caption{\textit{is digit 5}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/5-7-6.png}
-  \caption{\textit{is digit 5}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/5-7-9.png}
-  \caption{\textit{is digit 5}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/5-7-13.png}
-  \caption{\textit{is digit 5}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/5-7-17.png}
-  \caption{\mbox{\textit{eos 7 digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/5-7-22.png}
-  \caption{\mbox{\textit{eos 7 digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/5-7-26.png}
-  \caption{\mbox{\textit{eos 7 digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/5-7-27.png}
-  \caption{\mbox{\textit{eos 7 digit}}}
-\endminipage\hfill
+% \begin{figure}[t]
+% \label{fig1}
+% \captionsetup{labelformat=empty}
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/5-7-4.png}
+%   \caption{\textit{is digit 5}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/5-7-6.png}
+%   \caption{\textit{is digit 5}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/5-7-9.png}
+%   \caption{\textit{is digit 5}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/5-7-13.png}
+%   \caption{\textit{is digit 5}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/5-7-17.png}
+%   \caption{\mbox{\textit{eos 7 digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/5-7-22.png}
+%   \caption{\mbox{\textit{eos 7 digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/5-7-26.png}
+%   \caption{\mbox{\textit{eos 7 digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/5-7-27.png}
+%   \caption{\mbox{\textit{eos 7 digit}}}
+% \endminipage\hfill
 
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/9-2-4.png}
-  \caption{\mbox{\textit{eos digit 2}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/9-2-6.png}
-  \caption{\mbox{\textit{eos digit 2}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/9-2-9.png}
-  \caption{\mbox{\textit{eos digit 2}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/9-2-13.png}
-  \caption{\mbox{\textit{eos digit 2}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/9-2-17.png}
-  \caption{\mbox{\textit{of top digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/9-2-22.png}
-  \caption{\mbox{\textit{of top digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/9-2-26.png}
-  \caption{\mbox{\textit{of top digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/9-2-27.png}
-  \caption{\mbox{\textit{of top digit}}}
-\endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/9-2-4.png}
+%   \caption{\mbox{\textit{eos digit 2}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/9-2-6.png}
+%   \caption{\mbox{\textit{eos digit 2}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/9-2-9.png}
+%   \caption{\mbox{\textit{eos digit 2}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/9-2-13.png}
+%   \caption{\mbox{\textit{eos digit 2}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/9-2-17.png}
+%   \caption{\mbox{\textit{of top digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/9-2-22.png}
+%   \caption{\mbox{\textit{of top digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/9-2-26.png}
+%   \caption{\mbox{\textit{of top digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/9-2-27.png}
+%   \caption{\mbox{\textit{of top digit}}}
+% \endminipage\hfill
 
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/1-0-4.png}
-  \caption{\mbox{\textit{eos 0 of}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/1-0-6.png}
-  \caption{\mbox{\textit{eos digit 0}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/1-0-9.png}
-  \caption{\mbox{\textit{eos digit 0}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/1-0-12.png}
-  \caption{\mbox{\textit{eos 0 of}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/1-0-17.png}
-  \caption{\mbox{\textit{1 digit the}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/1-0-22.png}
-  \caption{\mbox{\textit{1 digit the}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/1-0-25.png}
-  \caption{\mbox{\textit{1 digit on}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/1-0-28.png}
-  \caption{\mbox{\textit{digit the eos}}}
-\endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/1-0-4.png}
+%   \caption{\mbox{\textit{eos 0 of}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/1-0-6.png}
+%   \caption{\mbox{\textit{eos digit 0}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/1-0-9.png}
+%   \caption{\mbox{\textit{eos digit 0}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/1-0-12.png}
+%   \caption{\mbox{\textit{eos 0 of}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/1-0-17.png}
+%   \caption{\mbox{\textit{1 digit the}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/1-0-22.png}
+%   \caption{\mbox{\textit{1 digit the}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/1-0-25.png}
+%   \caption{\mbox{\textit{1 digit on}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/1-0-28.png}
+%   \caption{\mbox{\textit{digit the eos}}}
+% \endminipage\hfill
 
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/4-8-4.png}
-  \caption{\textit{left 8 4}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/4-8-6.png}
-  \caption{\textit{left 8 4}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/4-8-9.png}
-  \caption{\mbox{\textit{left digit 4}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/4-8-14.png}
-  \caption{\mbox{\textit{eos 8 digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/4-8-18.png}
-  \caption{\mbox{\textit{eos 8 digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/4-8-22.png}
-  \caption{\mbox{\textit{8 digit the}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/4-8-25.png}
-  \caption{\mbox{\textit{8 digit digit}}}
-\endminipage\hfill
-\minipage{0.1\textwidth}
-  \includegraphics[width=\linewidth]{figures/4-8-28.png}
-  \caption{\mbox{\textit{digit digit the}}}
-\endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/4-8-4.png}
+%   \caption{\textit{left 8 4}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/4-8-6.png}
+%   \caption{\textit{left 8 4}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/4-8-9.png}
+%   \caption{\mbox{\textit{left digit 4}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/4-8-14.png}
+%   \caption{\mbox{\textit{eos 8 digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/4-8-18.png}
+%   \caption{\mbox{\textit{eos 8 digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/4-8-22.png}
+%   \caption{\mbox{\textit{8 digit the}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/4-8-25.png}
+%   \caption{\mbox{\textit{8 digit digit}}}
+% \endminipage\hfill
+% \minipage{0.1\textwidth}
+%   \includegraphics[width=\linewidth]{figures/4-8-28.png}
+%   \caption{\mbox{\textit{digit digit the}}}
+% \endminipage\hfill
 
-\captionsetup{labelformat=empty}
-\caption{Figure 1: Four examples show the generated images unfolded over several timesteps as well as the top three words the model attends to while generating images.}
-\end{figure}
+% \captionsetup{labelformat=empty}
+% \caption{Figure 1: Four examples show the generated images unfolded over several timesteps as well as the top three words the model attends to while generating images.}
+% \end{figure}
 
@@
 
-Microsoft COCO \citep{mscoco} is a largest dataset of images annotated with captions consisting of 83k images.
+Microsoft COCO \citep{mscoco} is the largest dataset of images annotated with captions consisting of roughly 83k images.
 The rich collection of images with variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on caption very challenging.
@@
 
-In the following subsections we make both qualitative and quantitative analysis of our model as well as compare its performance with other related generative models.
+In the following subsections we make both qualitative and quantitative analysis of our model as well as compare its performance with the performances of other related generative models.
 
 \subsubsection{Analysis of Generated Images}
-The main goal of this work is to learn a model that can understand the semantic meaning expressed in description of the image, such as the properties of objects, the relationships between them, etc.
+The main goal of this work is to learn a model that can understand the semantic meaning expressed in the descriptions of the images, such as the properties of objects, the relationships between them, etc.
 and then use that knowledge to generate relevant images.
-To verify that, we generated a set of captions inspired by COCO dataset and changed some words in the captions to see whether the model made the relevant changed in the generated samples.
+To verify that, we wrote a set of captions inspired by COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
 
-First, we wanted to see whether the model understood one of the simplest properties of any object, the color.
-We generated images of school buses with four different colors: yellow, red, green and blue.
+First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
+As shown in Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue.
 Although, there are images of buses with different colors in the training set, all school buses specifically are colored yellow.
-As you can see below, the model managed to generate images with an object that looked like a bus with the correct color.
-
-\vspace{-0.3cm}
-\begin{figure}[h]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\subfloat[A \underline{yellow} school bus parked in a parking lot.
-]{\includegraphics[width=0.22\textwidth]{figures/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
-%
-\subfloat[A \underline{red} school bus parked in a parking lot.
-]{\includegraphics[width=0.22\textwidth]{figures/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
-%
-\subfloat[A \underline{green} school bus parked in a parking lot.
-]{\includegraphics[width=0.22\textwidth]{figures/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
-%
-\subfloat[A \underline{blue} school bus parked in a parking lot.
-]{\includegraphics[width=0.22\textwidth]{figures/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
-%
-\end{center}
-%\caption{whole image caption}
-\label{fig:label}
-\vspace{-0.2cm}
-\end{figure}
-
-Apart from changing the colors of objects, we were curious whether changing the background of the scene described in the caption would result in the appropriate changes in the generated samples.
-It is a somewhat harder task for a model than changing the color, due to larger number of changes that have to be made in the generated samples.
-Nevertheless as shown below, changing the grass type from dry to green as well as adding clouds in the caption has resulted in the appropriate changes.
-
-\vspace{-0.3cm}
-\begin{figure}[h]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\subfloat[A very large commercial plane flying in the \underline{blue} skies.
-]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-blue-skies-sharp.png}}\quad
-%
-\subfloat[A very large commercial plane flying in the \underline{cloudy} skies.
-]{\includegraphics[width=0.48\textwidth]{figures/a-very-large-commercial-plane-flying-in-the-cloudy-skies-sharp.png}}\quad
-%
-%
-\end{center}
-%\caption{whole image caption}
-\label{fig:label}
-\vspace{-0.8cm}
-\end{figure}
-
-\begin{figure}[h]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\subfloat[A heard of elephants walking across a \underline{dry} grass field.
-]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
-%
-\subfloat[A heard of elephants walking across a \underline{green} grass field.
-]{\includegraphics[width=0.48\textwidth]{figures/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
-%
-%
-\end{center}
-%\caption{whole image caption}
-\label{fig:label}
-\vspace{-0.2cm}
-\end{figure}
-
-Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we have found that the model made appropriate changes as long as some similar patter was present in the training set.
-However, it wasn't always the case when changing an object itself in the description.
-In cases, when objects did't have a noticeable differences in their properties, such as shape or color, the changes in the generated samples weren't very clearly seen.
-This could be attributed to the limitation of the model to grasp the detailed understanding of each object.
-
-\vspace{-0.3cm}
-\begin{figure}[h]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\subfloat[The \underline{decadent chocolate dessert} is on the table.
-]{\includegraphics[width=0.48\textwidth]{figures/the-decadent-chocolate-dessert-is-on-the-table-2-sharp.png}}\quad
-%
-\subfloat[The \underline{bowl of banas} is on the table.
-]{\includegraphics[width=0.48\textwidth]{figures/a-bowl-of-bananas-is-on-the-table-2-sharp.png}}\quad
-%
-%
-\end{center}
-%\caption{whole image caption}
-\label{fig:label}
-\vspace{-0.8cm}
-\end{figure}
+Despite that, the model managed to generate images of an object that is reminiscent of the bus pained with the relevant color.
 
@@
 \begin{center}
+\subfloat[A \underline{yellow} school bus parked in a parking lot.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[A \underline{red} school bus parked in a parking lot.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[A \underline{green} school bus parked in a parking lot.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[A \underline{blue} school bus parked in a parking lot.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\end{center}
+\caption{Examples of changing the color while keeping the caption fixed.}
+\label{fig:genimages1}
+\vspace{-0.3cm}
+\end{figure}
+
+Apart from changing the colors of objects, we were curious whether changing the background of the scene described in the caption would result in the appropriate changes in the generated samples.
+Changing the background in images is a somewhat harder task for a model that changing the color, due to the larger visual area in images that is taken by the background.
+Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy as well as changing the grass type from dry to green resulted in the appropriate changes.
+The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.
+
+\begin{figure}[!h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\\
+\vspace{-0.45cm}
+%
+\subfloat[A very large commercial plane flying in \underline{blue} skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
+%
+\subfloat[A very large commercial plane flying in \underline{rainy} skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
+%
+\subfloat[A herd of elephants walking across a \underline{dry} grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
+%
+\subfloat[A herd of elephants walking across a \underline{green} grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
+%
+\end{center}
+\caption{Examples of changing the background while keeping the caption fixed. The respective nearest training images based on pixelwise distance are displayed on top.}
+\label{fig:genimages2}
+\vspace{-0.3cm}
+\end{figure}
+
+Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar patter was present in the training set.
+However, it wasn't always the case when changing an object itself in the description.
+As you can see in Figure~\ref{fig:genimages3}, when objects didn't have a clear fine grained differences, such as different shape or color, the relevant changes in the generated samples weren't very clearly seen.
+This highlighted the limitation of the model to grasp the detailed understanding of each object.
+
+\begin{figure}[!h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
+]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
+%
+\subfloat[\underline{A bowl of banas} is on the table.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
+%
 \subfloat[A vintage photo of a \underline{cat}.
-]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-cat-sharp.png}}\quad
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
 %
 \subfloat[A vintage photo of a \underline{dog}.
-]{\includegraphics[width=0.48\textwidth]{figures/a-vintage-photo-of-a-dog-sharp.png}}\quad
-%
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
 %
 \end{center}
-%\caption{whole image caption}
-\label{fig:labelr}
-\vspace{-0.2cm}
+\caption{Examples of changing the object while keeping the caption fixed.}
+\label{fig:genimages3}
+\vspace{-0.3cm}
 \end{figure}
@@
 
-To be added soon.
+Unfortunately, we found that there was no real connection between the patches drawn on canvas and most attended words at each timestep.
+During the generation, the model mostly focused on several words that were represented the idea in the caption.
+The most attended words allowed us to interpret whether the model would make relevant changes.
 
 \subsubsection{Comparison With Other Models}
-The quantitative evaluation of generative models has been a subject of ambiguity within a machine learning community.
-Compared to reporting classification accuracies, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To avoid ambiguity, we report results on two different metrics as well as do qualitative comparison of generative models.
+The quantitative evaluation of generative models has been a subject of ambiguity in a machine learning community.
+Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
+To get a better comparisions between performances of generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
 
-As shown on Figure blah, we have sampled the images from the same caption in four of the best models.
-
-First, to compare the the peformance of our model with other generative models trained with variational objective, we report the Recall$@$K, namely the mean number of sentences for which the correct image is ranked within the top-K retrieved results, based on the lower bound of likelihood $\loss$.
-If you want to get a really good results build a model like Ryan did.
-Do deal away with easy images we take ratio with ...
-with mean caption.
-To deal with large computational complexity of looping over each image we only consider closest neightbors.
-The precision Recall curve is reported there...
-
-Second,
-
-\vspace{-0.3cm}
 \begin{figure}[!t]
@@
 \caption{Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
-\label{fig:label}
+\label{fig:diffmodels}
 \vspace{-0.2cm}
@@
 
-\subsection{CIFAR}
+As you can see on Figure~\ref{fig:diffmodels}, we generated several samples from prior of each of the current state-of-the-art generative models corresponding to the caption ``A group of people walk on a beach with surf boards".
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and don't have a very clear structure in them, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
+
+As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods.
+We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model.
+As we expected, the quality of image retrieval using generative models is worse compared to the disriminative models that were specifically build for retrieval.
+To deal with large computational complexity of looping through each test image, we create a shortlist of hundred images including the correct one, based on the convolutional features of VGG like model trained on CIFAR dataset.
+Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likehood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
+We found that the reconstruction error $\rloss$ increased for the sharpened images that considerably hurt the retrieval results.
+Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.
+
+Instead of calculating error per each pixel, we turn to the smarter metric, such as Structural Similarity Index (SSI), which incorporates luminace and contrast masking into the error calculation.
+Due to strong inter-dependencies of closer pixels, the metric is calculated on the small windows of the image.
+To calculate SSI, we sampled fifty images from prior of each generative model per each caption in the test set.
+
+\begin{table}[!t]
+\begin{center}
+\begin{tabulary}{\linewidth}{c || c c c c c || c}
+\hline
+\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
+\hline
+& \multicolumn{5}{c||}{Image Search} & Image Similarity \\
+%\cline{2-6}
+\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
+\hline
+\hline
+LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
+\hline
+Fully-conn. VAE (L2 cost) & 0.7 & 4.6 & 8.7 & 46.8 & 54 & 0.122 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv. VAE (L2 cost) & 0.7 & 4.4 & 8.5 & 46.6 & 54 & \textbf{0.131} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+Noalign DRAW & \textbf{1.0} & 6.1 & 11.2 & 52.1 & \textbf{48} & 0.114 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+Skipthought DRAW & \textbf{1.0} & \textbf{6.2} & \textbf{11.5} & \textbf{52.5} & \textbf{48} & 0.118 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+Align DRAW & 0.9 & \textbf{6.2} & 11.2 & \textbf{52.5} & \textbf{48} & 0.115 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+\end{tabulary}
+\end{center}
+\end{table}
 

### 2015-10-07 iclr-paper.tex
new changes

@@
 
-\begin{figure}[!h]
+\begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
@@
 \subsubsection{Analysis of Attention}
+Unfortunately, we found that there was no connection between the patches drawn on canvas and most attended words at each timestep.
+During the image generation, the model mostly focused on several words that carried the semantic meaning of caption.
+The words which were mostly attended during all generation steps indicated the kind of scene the model would generate.
+For example, as shown in Figure~\ref{fig:attentimages1} by equally looking at words deset and forest allowed the model to make relevant changes in the scene.
+Whereas in the second example, the model completely ignored word sun and didn't make any changes.
 
-Unfortunately, we found that there was no real connection between the patches drawn on canvas and most attended words at each timestep.
-During the generation, the model mostly focused on several words that were represented the idea in the caption.
-The most attended words allowed us to interpret whether the model would make relevant changes.
+\begin{figure}[!h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{desert}}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
+%
+\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{forest}}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
+%
+\subfloat[\hlctwo{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlctwo{beach}}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
+%
+\subfloat[\hlcthree{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlczero{sun}}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
+%
+\end{center}
+\caption{Examples of most attended words while changing the background in the caption.}
+\label{fig:attentimages1}
+\vspace{-0.3cm}
+\end{figure}
 
@@
 \hline
-Fully-conn. VAE (L2 cost) & 0.7 & 4.6 & 8.7 & 46.8 & 54 & 0.122 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv. VAE (L2 cost) & 0.7 & 4.4 & 8.5 & 46.6 & 54 & \textbf{0.131} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-Noalign DRAW & \textbf{1.0} & 6.1 & 11.2 & 52.1 & \textbf{48} & 0.114 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-Skipthought DRAW & \textbf{1.0} & \textbf{6.2} & \textbf{11.5} & \textbf{52.5} & \textbf{48} & 0.118 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-Align DRAW & 0.9 & \textbf{6.2} & 11.2 & \textbf{52.5} & \textbf{48} & 0.115 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+Fully-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+Skipthought DRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+Noalign DRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+Align DRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
 \end{tabulary}
@@
 \end{table}
+
+% \begin{table}[!t]
+% \begin{center}
+% \begin{tabulary}{\linewidth}{c || c c c c c || c}
+% \hline
+% \multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
+% \hline
+% & \multicolumn{5}{c||}{Image Search} & Image Similarity \\
+% %\cline{2-6}
+% \textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
+% \hline
+% \hline
+% LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
+% \hline
+% Fully-conn. VAE (L2 cost) & 0.7 & 4.6 & 8.7 & 46.8 & 54 & 0.122 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+% Conv. VAE (L2 cost) & 0.7 & 4.4 & 8.5 & 46.6 & 54 & \textbf{0.131} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+% Noalign DRAW & \textbf{1.0} & 6.1 & 11.2 & 52.1 & \textbf{48} & 0.114 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+% Skipthought DRAW & \textbf{1.0} & \textbf{6.2} & \textbf{11.5} & \textbf{52.5} & \textbf{48} & 0.118 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+% Align DRAW & 0.9 & \textbf{6.2} & 11.2 & \textbf{52.5} & \textbf{48} & 0.115 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+% \end{tabulary}
+% \end{center}
+% \end{table}
 

### 2015-10-07 ram_workshop/ram-paper.tex
update ram workshop
update ram

--- 2015-10-07 iclr-paper.tex
@@
 \documentclass{article} % For LaTeX2e
-\usepackage{iclr-style/iclr2016_conference,times}
+\usepackage{nips15submit_e,times}
 \usepackage{hyperref}
@@
 \usepackage{xcolor}
+%\usepackage{natbib}
+%\usepackage[style=numeric,sorting=none,maxnames=2]{natbib}
+\usepackage[square,comma,sort&compress,numbers]{natbib} 
+\bibliographystyle{unsrt85}
 
@@
 \newcommand{\new}{\marginpar{NEW}}
-
+\graphicspath{{../}}
 \begin{document}
@@
 This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components.
-Image generative models studied previously are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
-Despite the advances in generative models, learning the highly structured natural image distribution alone in the high dimensional pixel space alone proves to be a difficult task.
+Image generative models studied previously, e.g.
+variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}, are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
+Despite the advances in the variational generative models \cite{kingma_vae}\cite{gregor_draw}, learning the highly structured natural image distribution alone in the high dimensional pixel space alone proves to be a difficult task.
 In real world, however, images rarely appear in isolation.
@@
 
-In this paper, we illustrate how simple sequential deep learning techniques can be used to build a conditional probabilistic model over natural image space effectively.
-By using a sequence to sequence framework to approach the problem of image generation from unstructured natural language captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
-Overall, the main contributions of this work are the following:
-
+In this paper, we address the problem of image generation from unstructured natural language captions.
+By extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw}, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
+Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images using soft attention mechanism attending to different input words while generating images.
+The images generated by our alignDRAW model are post-process by.
+We then illustrate how our method generalizes over new color combinations and environments.
+\comm{
 \section{Related Work}
 
-Deep Neural Networks have achieved a remarkable performance in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech} and etc.
-While most of the recent success has been achieved by discriminative models, the generative models have not yet enjoyed the same level of success.
-Most of the previous work in generative models has been focused on variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}.
-While these models are very powerful, each iteration of training requires a computationally costly step of MCMC to approximate an intractable normalization constant that makes it hard to scale them to large datasets.
-
-\cite{kingma_vae} have introduced the Variational Auto-Encoder (VAE) which can be seen as a neural network with continous latent variables.
+ which can be seen as a neural network with continous latent variables.
 The encoder is used to approximate posterior distribution and the decoder is used to stochastically reconstruct the data from latent variables.
@@
 While all of the previous work has been focused on unconditional models or models conditioned on labels, to the best of our knowledge this paper is the first to introduce the generative model of images conditioned on captions.
-
+}
 \begin{figure}[!t]
@@
 
-\bibliography{iclr-paper}
-\bibliographystyle{iclr-style/iclr2016_conference}
+{\footnotesize
+\bibliography{ram-paper}}
 

### 2015-10-07 ram_workshop/ram-paper.tex
reworded

@@
 
-Microsoft COCO \citep{mscoco} is the largest dataset of images annotated with captions consisting of roughly 83k images.
-The rich collection of images with variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on caption very challenging.
-Since some of the images have more than five captions attached to them, for consistency with related work on caption generation we disregard extra captions.
+Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
+The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on a caption very challenging.
+For consistency with related work on caption generation, we disregard four of the five captions when training our model.
+%Since some of the images have more than five captions attached to them,
 
-In the following subsections we make both qualitative and quantitative analysis of our model as well as compare its performance with the performances of other related generative models.
+In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models.
 
 \subsubsection{Analysis of Generated Images}
-The main goal of this work is to learn a model that can understand the semantic meaning expressed in the descriptions of the images, such as the properties of objects, the relationships between them, etc.
+The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
 and then use that knowledge to generate relevant images.
-To verify that, we wrote a set of captions inspired by COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
+To verify that, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
 
 First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
-As shown in Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue.
-Although, there are images of buses with different colors in the training set, all school buses specifically are colored yellow.
-Despite that, the model managed to generate images of an object that is reminiscent of the bus pained with the relevant color.
+In Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue.
+Although, there are images of buses with different colors in the training set, all mentioned school buses are specifically colored yellow.
+Despite that, the model managed to generate images of an object that is visually reminiscent of a school bus that is painted with the specified color.
 
@@
 
-Apart from changing the colors of objects, we were curious whether changing the background of the scene described in the caption would result in the appropriate changes in the generated samples.
-Changing the background in images is a somewhat harder task for a model that changing the color, due to the larger visual area in images that is taken by the background.
-Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy as well as changing the grass type from dry to green resulted in the appropriate changes.
+Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
+The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
+Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
 The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.
@@
 \end{center}
-\caption{Examples of changing the background while keeping the caption fixed. The respective nearest training images based on pixelwise distance are displayed on top.}
+\caption{Examples of changing the background while keeping the caption fixed. The respective nearest training images based on pixel-wise L2 distance are displayed on top.}
 \label{fig:genimages2}
@@
 
-Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar patter was present in the training set.
-However, it wasn't always the case when changing an object itself in the description.
-As you can see in Figure~\ref{fig:genimages3}, when objects didn't have a clear fine grained differences, such as different shape or color, the relevant changes in the generated samples weren't very clearly seen.
-This highlighted the limitation of the model to grasp the detailed understanding of each object.
+Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
+However, the model struggled when the visual difference between objects was very small, such as the objects having the same general shape and colors.
+In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
+This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
 
@@
 %
-\subfloat[\underline{A bowl of banas} is on the table.
+\subfloat[\underline{A bowl of bananas} is on the table.
 ]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad

### 2015-10-07 ram_workshop/ram-paper.tex
changes

@@
 \subsubsection{Comparison With Other Models}
-The quantitative evaluation of generative models has been a subject of ambiguity in a machine learning community.
+Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffer from its own specific drawbacks.
 Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To get a better comparisions between performances of generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
+To get a better comparision between performances of different generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
 
@@
 
-As you can see on Figure~\ref{fig:diffmodels}, we generated several samples from prior of each of the current state-of-the-art generative models corresponding to the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and don't have a very clear structure in them, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
+In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects don't have a clear structure in them, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
 

### 2015-10-07 ram_workshop/ram-paper.tex
added new changes to ram_workshop. mostly changed analysis of attention section

--- 2015-10-07 ram_workshop/ram-paper.tex
@@
 
-Unfortunately, we found that there was no real connection between the patches drawn on canvas and most attended words at each timestep.
-During the generation, the model mostly focused on several words that were represented the idea in the caption.
-The most attended words allowed us to interpret whether the model would make relevant changes.
+After flipping set of words in the captions, we were curious to see which words did the model attend to when generating images.
+It turned out that during the generation step, the model mostly focused on the specific words that carried out the main semantic meaning expressed in the sentences.
+The attention values in sentences helped us interpret the reasons why the model did or didn't make relevant changes in images when flipping words.
+For example, as shown in Figure~\ref{fig:attentimages1} despite flipping word ``desert'' to ``forest'', the model equally looked at these words and thus made the correct changes.
+Whereas, flipping between words ``beach'' and ``sun'' has resulted in no change, due to the model ignoring the word ``sun''.
+
+We also tried to analyze the way the model generated images.
+Unfortunately, we found that there was no connection between the patches drawn on canvas and the most attended words at particular timesteps.
+
+\begin{figure}[!h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{desert}}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
+%
+\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{forest}}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
+%
+\subfloat[\hlctwo{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlctwo{beach}}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
+%
+\subfloat[\hlcthree{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlczero{sun}}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
+%
+\end{center}
+\caption{Examples of most attended words while changing the background in the caption.}
+\label{fig:attentimages1}
+\vspace{-0.3cm}
+\end{figure}
 
@@
 %
-\subfloat{\includegraphics[width=0.23\textwidth]{to-do.png}}\quad
+\subfloat[Fully-conn. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae.png}}\quad
 %
@@
 
-\begin{table}[!t]
+\begin{table}[!b]
 \begin{center}

### 2015-10-07 ram_workshop/ram-paper.tex
cut1

--- 2015-10-07 ram_workshop/ram-paper.tex
@@
 \newcommand{\oimage}{{\bf{x}}}
-\newcommand{\post}{Q}
-\newcommand{\prior}{P}
+\newcommand{\post}{q}
+\newcommand{\prior}{p}
 \newcommand{\loss}{\mathcal{L}}
@@
 variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}, are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
-Despite the advances in the variational generative models \cite{kingma_vae}\cite{gregor_draw}, learning the highly structured natural image distribution alone in the high dimensional pixel space alone proves to be a difficult task.
+Despite the advances in the variational generative models \cite{kingma_vae}\cite{gregor_draw} and the Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, learning the highly structured natural image distribution alone in the high dimensional pixel space alone proves to be a difficult task.
 In real world, however, images rarely appear in isolation.
@@
 The additional information from image and unstructured text description could be used to simplify the image modelling task.
+In addition, learning conditional generative model from text allows better understanding of the generalization performance of the model to new scenes.
 
-There are two directions in learning a generative model of image and text.
-One approach is to learn a text generative model conditioned on the images.
+\comm{There are two directions in learning a generative model of image and text.  One approach is to learn a text generative model conditioned on the images.}
+
 Significant amount of recent works has been focused on generating captions from images \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
-The models take an image descriptor and generate unstructured texts through a recurrent decoder.
 By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description.
 Generating high dimensional realistic images from their descriptions is a more difficult approach that combines two challenging components of language modelling and image generation.
-Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image.
+
+\comm{The models take an image descriptor and generate unstructured texts through a recurrent decoder.}
+
+\comm{Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image.
 Although, the interesting high dimensional natural images lay on a small manifold that is difficult to capture, the additional text description cues of a target image may simplify the learning problem by focusing on the conditional distribution.
+}
 
@@
 Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images using soft attention mechanism attending to different input words while generating images.
-The images generated by our alignDRAW model are post-process by.
-We then illustrate how our method generalizes over new color combinations and environments.
+The images generated by our alignDRAW model are refined by post-processing deterministic Laplacian pyramid adversarial network \cite{denton_lapgan}.
+We then illustrate how our method, learnt on Microsoft COCO dataset, generalizes to the captions describing novel scenarios that is unseen from the training set.
+
 \comm{
@@
 \begin{figure}[!t]
+\vspace{-0.3in}
 \captionsetup[subfigure]{labelformat=empty}
@@
 \end{center}
+\vspace{-0.2in}
 \caption{The image of the proposed model}
 \label{fig:figmodel}
-\vspace{-0.3cm}
+\vspace{-1cm}
 \end{figure}
-
+\vspace{-0.1in}
 \section{Model}
+\vspace{-0.1in}
 Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
@@
 
-\subsection{Language Model: the Bidirectional Attention RNN}
+\subsection{Language Representation: the Bidirectional Attention RNN}
 \label{sec:lang}
@@
 
-\subsection{Image Model: the Conditional DRAW Network}
-
+\subsection{Image Generation: the Conditional DRAW Network}
+\comm{
 The DRAW network \cite{gregor_draw} is a sequential probabilistic model generating images by accumulating the output at each iterative step.
 While the original DRAW network assumes the latent variables are independent, it has shown in \citep{bachman_sdm} the model performance is improved by including the dependencies of latent variables.
-
-We extended the architecture the DRAW network generative process to include additional input caption from the language model described in Sec.
+}
+We extended the architecture the DRAW network \citep{gregor_draw} generative process to include additional input caption from the language model described in Sec.
 (\ref{sec:lang}).
@@
 \label{eq:x_hat}
-\hat{x}_t &= x-\sigmoid(\canv_{t-1})\\
-\label{eq:read}
-r_t &= \readop(x_t, \hat{x}_t, \hdec_{t-1})\\
-\henc_t &= \encoder(\henc_{t-1}, [r_t, \hdec_{t-1}])\\
-\lat_t &\sim \post(\Lat_t|\henc_t)\\
+\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1})\\
 \hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1})\\
@@
 \end{align}
-
-The corresponding alignment probabilities $\alpha_{1...n}$ at each step are obtained by:
+The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using caption representation $\hlang$ and hidden state of the generative model $\hdec_{t}$ as in \citep{bahdanau_mt}
+\comm{
 \begin{align}
@@
 \end{align}
-
+}
+\comm{
 Here $\hlang_{0}$ is initialized to the learned bias.
 Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. 
-
+}
 \subsection{Learning}
 
-The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}.
+The conditional generative model is trained to maximize the variational lower bound on the log-likelihood $\icaption$, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$ similar to \cite{kingma_vae}.
+The posterior inference is approximated by an inference network $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in red dash box in Figure (\ref{fig:figmodel}).
+Unlike the original DRAW network where latent variables are independent unit Gaussian $\mathcal{N}(0, I)$, the align DRAW model uses latent variables whose mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{}.
+In particular, the prior distribution $\prior(\Lat_t\given\Lat_{1:t-1})$ takes the form $\mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1}))$.
+
+\comm{
+is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}.
 The model is trained to maximize the lower bound of marginal likelihood $\loss$ of the correct image $x$ given the input caption $y$.
@@
 \end{align}
-Overall, the loss function $\loss$ is calculated as follows:
+}
+Overall, the variational objective $\loss$ is defined with the model parameters vector $\theta$ as follows:
 \begin{align}
-\loss =  &\expectation_{Q(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
+\loss_\theta =  &\expectation_{\post(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
 \end{align}
+
+\comm{
 The expectation can be approximate by $\numSamples$ Monte Carlo samples $\LatSample_{1:T}$ from $\post(\Lat_{1:T}\given\icaption, \oimage)$:
@@
 }
-
+}
 \subsection{Generating Images from Captions}
@@
 
-Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
-The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on a caption very challenging.
-For consistency with related work on caption generation, we disregard four of the five captions when training our model.
-%Since some of the images have more than five captions attached to them,
+Microsoft COCO \citep{mscoco} is the largest dataset of images annotated with captions consisting of roughly 83k images.
+The rich collection of images with variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on caption very challenging.
+Since some of the images have more than five captions attached to them, for consistency with related work on caption generation we disregard extra captions.
 
-In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models.
+In the following subsections we make both qualitative and quantitative analysis of our model as well as compare its performance with the performances of other related generative models.
 
 \subsubsection{Analysis of Generated Images}
-The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
+The main goal of this work is to learn a model that can understand the semantic meaning expressed in the descriptions of the images, such as the properties of objects, the relationships between them, etc.
 and then use that knowledge to generate relevant images.
-To verify that, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
+To verify that, we wrote a set of captions inspired by COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
 
 First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
-In Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue.
-Although, there are images of buses with different colors in the training set, all mentioned school buses are specifically colored yellow.
-Despite that, the model managed to generate images of an object that is visually reminiscent of a school bus that is painted with the specified color.
+As shown in Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue.
+Although, there are images of buses with different colors in the training set, all school buses specifically are colored yellow.
+Despite that, the model managed to generate images of an object that is reminiscent of the bus pained with the relevant color.
 
@@
 
-Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
-The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
-Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
+Apart from changing the colors of objects, we were curious whether changing the background of the scene described in the caption would result in the appropriate changes in the generated samples.
+Changing the background in images is a somewhat harder task for a model that changing the color, due to the larger visual area in images that is taken by the background.
+Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy as well as changing the grass type from dry to green resulted in the appropriate changes.
 The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.
@@
 \end{center}
-\caption{Examples of changing the background while keeping the caption fixed. The respective nearest training images based on pixel-wise L2 distance are displayed on top.}
+\caption{Examples of changing the background while keeping the caption fixed. The respective nearest training images based on pixelwise distance are displayed on top.}
 \label{fig:genimages2}
@@
 
-Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
-However, the model struggled when the visual difference between objects was very small, such as the objects having the same general shape and colors.
-In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
-This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
+Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar patter was present in the training set.
+However, it wasn't always the case when changing an object itself in the description.
+As you can see in Figure~\ref{fig:genimages3}, when objects didn't have a clear fine grained differences, such as different shape or color, the relevant changes in the generated samples weren't very clearly seen.
+This highlighted the limitation of the model to grasp the detailed understanding of each object.
 
@@
 %
-\subfloat[\underline{A bowl of bananas} is on the table.
+\subfloat[\underline{A bowl of banas} is on the table.
 ]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
@@
 \subsubsection{Comparison With Other Models}
-Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffer from its own specific drawbacks.
+The quantitative evaluation of generative models has been a subject of ambiguity in a machine learning community.
 Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To get a better comparision between performances of different generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
+To get a better comparisions between performances of generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
 
@@
 
-In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects don't have a clear structure in them, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
+As you can see on Figure~\ref{fig:diffmodels}, we generated several samples from prior of each of the current state-of-the-art generative models corresponding to the caption ``A group of people walk on a beach with surf boards".
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and don't have a very clear structure in them, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
 

### 2015-10-08 ram_workshop/ram-paper.tex
fixed intro

@@
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding. 
-This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components.
-Image generative models studied previously, e.g.
-variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}, are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
-Despite the advances in the variational generative models \cite{kingma_vae}\cite{gregor_draw} and the Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, learning the highly structured natural image distribution alone in the high dimensional pixel space alone proves to be a difficult task.
-In real world, however, images rarely appear in isolation.
-They often accompanied by their unstructured textual descriptions on web pages and in books.
-One domain presents substantial amount of relating information of the other domain.
-The additional information from image and unstructured text description could be used to simplify the image modelling task.
-In addition, learning conditional generative model from text allows better understanding of the generalization performance of the model to new scenes.
+%This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components. 
+Generative models of images studied previously, e.g. variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn},  
+often defined distributions that were restricted to being either unconditioned or conditioned on simple structured annotations, for example classification labels.
+%are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels. 
+Despite the recent advances in new methods of generative models, such as Variational Auto-Encoders (VAE) \cite{kingma_vae}\cite{gregor_draw} and Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, learning the highly structured natural image distribution in high-dimensional pixel space proves to be a difficult task.
+In real world circumstances, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
+%One domain presents substantial amount of relating information of the other domain. 
+The additional information from these descriptions could be used to simplify the image modelling task. %%
+Moreover, learning generative models conditioned on text allows a better understanding of the generalization performance of the model, as we can create textual descriptions of completely new scenes not seen at training time.
 
-\comm{There are two directions in learning a generative model of image and text.  One approach is to learn a text generative model conditioned on the images.}
+\comm{There are two directions in learning a generative model of images conditioned text.  One approach is to learn a text generative model conditioned on the images.}
 
-Significant amount of recent works has been focused on generating captions from images \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
-By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description.
-Generating high dimensional realistic images from their descriptions is a more difficult approach that combines two challenging components of language modelling and image generation.
+A significant amount of recent work has been focused on generating captions from images \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions}.
+A question that can be posed is whether the opposite is possible, i.e.
+generating an image given a caption.
+%By contrast, learning a generative model for images and text may also be studied by generating images correctly interpreting the text description. 
+Generating high dimensional realistic images from their descriptions is a more difficult problem as it requires both language modelling and image generation.
 
@@
 In this paper, we address the problem of image generation from unstructured natural language captions.
-By extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw}, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
-Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images using soft attention mechanism attending to different input words while generating images.
-The images generated by our alignDRAW model are refined by post-processing deterministic Laplacian pyramid adversarial network \cite{denton_lapgan}.
-We then illustrate how our method, learnt on Microsoft COCO dataset, generalizes to the captions describing novel scenarios that is unseen from the training set.
+By extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw}, our model iteratively draws patches on a canvas, while attending to relevant words in the description.
+Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images using a soft attention mechanism attending to different input words while generating images.
+The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \cite{denton_lapgan}.
+We then illustrate how our method, learnt on the Microsoft COCO dataset, generalizes to captions describing novel scenarios that are not seen in the training set.
 

### 2015-10-08 ram_workshop/ram-paper.tex
clean up

@@
 Despite the recent advances in new methods of generative models, such as Variational Auto-Encoders (VAE) \cite{kingma_vae}\cite{gregor_draw} and Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, learning the highly structured natural image distribution in high-dimensional pixel space proves to be a difficult task.
-In real world circumstances, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
+In real world applications, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
 %One domain presents substantial amount of relating information of the other domain. 
@@
 \vspace{-0.1in}
-Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+Our proposed model can be viewed as utilizing the sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}, where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
 Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $\oimage$ be the image corresponding to that caption.
@@
 The input caption sentences are fed into a deterministic Bidirectional LSTM\cite{} that encodes the variable size sentences into the vector representation $s$.
-Bidirectional LSTM consists of one Forward LSTM and Backward LSTM which combine information from past and future respectively.
+A Bidirectional LSTM consists of one Forward LSTM and one Backward LSTM which combine information from past and future, respectively.
 The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$.
@@
 }
-We extended the architecture the DRAW network \citep{gregor_draw} generative process to include additional input caption from the language model described in Sec.
-(\ref{sec:lang}).
-Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
-The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}:
+We extended the architecture of the generative DRAW network \citep{gregor_draw} 
+%generative process an input caption from the
+to include a language model, described in Sec.
+(\ref{sec:lang}), which process the conditioned caption.
+Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of an Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and a Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
+The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps, which can be seen as an adaptation of the translation model of \cite{bahdanau_mt}:
 
@@
 \end{align}
-The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using caption representation $\hlang$ and hidden state of the generative model $\hdec_{t}$ as in \citep{bahdanau_mt}
+The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlang$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}
 \comm{
@@
 
-The conditional generative model is trained to maximize the variational lower bound on the log-likelihood $\icaption$, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$ similar to \cite{kingma_vae}.
-The posterior inference is approximated by an inference network $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in red dash box in Figure (\ref{fig:figmodel}).
-Unlike the original DRAW network where latent variables are independent unit Gaussian $\mathcal{N}(0, I)$, the align DRAW model uses latent variables whose mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{}.
+The conditional generative model is trained to maximize the variational lower bound on the log-likelihood $\icaption$, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$ similarly to \cite{kingma_vae}.
+The posterior inference is approximated by an inference network $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in the red dashed box in Figure (\ref{fig:figmodel}).
+Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the align DRAW model uses latent variables whose mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{}.
 In particular, the prior distribution $\prior(\Lat_t\given\Lat_{1:t-1})$ takes the form $\mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1}))$.
@@
 
-During the image generation step, we throw away the inference network from the decoder and instead sample from the prior distribution.
-Due to the blurriness of samples generated by DRAW model, we do an additional post processing step, where we use the generator of the adversarial network trained on residuals of laplacian pyramid to sharpen the generated images, similar to \citep{denton_lapgan}.
-By fixing the prior of adversarial generator to the mean of uniform distribution, it gets treated as a deterministic neural network which allows us to calculate the lower bound of likelihood.
+During the image generation step, we detach the inference network from the decoder and discard it, and instead sample from the prior distribution.
+Due to the blurriness of samples generated by the DRAW model, we do an additional post processing step, where we use the generator of an adversarial network trained on the residuals of a laplacian pyramid to sharpen the generated images,
+similar to \citep{denton_lapgan}.
+By fixing the prior of the adversarial generator to be the mean of the original uniform distribution, it gets treated as a deterministic neural network which allows us to calculate the lower bound of the likelihood.
 The reconstuction loss becomes the loss between sharpened image and correct image, whereas the latent loss stays the same.
-We also noticed that sampling from the mean of uniform distribution allowed us to generate much less noisy samples than by sampling from the uniform distribution itself.
+We also noticed that inputting the mean of the uniform distribution into the edge generator network allowed us to generate samples with much less noise than if we had sampled from the uniform distribution.
 
@@
 \end{center}
-\caption{Examples of changing the background while keeping the caption fixed. The respective nearest training images based on pixel-wise L2 distance are displayed on top.}
+\caption{Bottom: Examples of changing the background while keeping the caption fixed. Top: The respective nearest training images based on pixel-wise L2 distance.}
 \label{fig:genimages2}
@@
 Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
-However, the model struggled when the visual difference between objects was very small, such as the objects having the same general shape and colors.
+However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
 In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
@@
 
-After flipping set of words in the captions, we were curious to see which words did the model attend to when generating images.
-It turned out that during the generation step, the model mostly focused on the specific words that carried out the main semantic meaning expressed in the sentences.
-The attention values in sentences helped us interpret the reasons why the model did or didn't make relevant changes in images when flipping words.
-For example, as shown in Figure~\ref{fig:attentimages1} despite flipping word ``desert'' to ``forest'', the model equally looked at these words and thus made the correct changes.
-Whereas, flipping between words ``beach'' and ``sun'' has resulted in no change, due to the model ignoring the word ``sun''.
+After flipping sets of words in the captions, we were curious to see which words the model attended to when generating images.
+It turned out that during the generation step, the model mostly focused on the specific words (or their neighbors) that carried the main semantic meaning expressed in the sentences.
+The attention values in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
+For example, in Figure~\ref{fig:attentimages1} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
+This suggests that, in their respective sentences, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
+In contrast, when we swap words ``beach'' and ``sun'', we can see a drastic change between sentences in the probability distribution over words.
+By noting that the model completely ignores the word ``sun'' in the second sentence, we can therefore gain a more thorough understanding of why we see no visual differences between the images generated by each caption.
 
@@
 Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To get a better comparision between performances of different generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
+To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
 
@@
 In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects don't have a clear structure in them, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects,
+%don't have a clear structure in them
+whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
 
@@
 We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model.
-As we expected, the quality of image retrieval using generative models is worse compared to the disriminative models that were specifically build for retrieval.
-To deal with large computational complexity of looping through each test image, we create a shortlist of hundred images including the correct one, based on the convolutional features of VGG like model trained on CIFAR dataset.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likehood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
-We found that the reconstruction error $\rloss$ increased for the sharpened images that considerably hurt the retrieval results.
+Unsurprisingly, the quality of image retrieval using generative models is worse than that of disriminative models that were specifically built for retrieval.
+To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on images having closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
+Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
+We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.
 Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.
 
-Instead of calculating error per each pixel, we turn to the smarter metric, such as Structural Similarity Index (SSI), which incorporates luminace and contrast masking into the error calculation.
-Due to strong inter-dependencies of closer pixels, the metric is calculated on the small windows of the image.
-To calculate SSI, we sampled fifty images from prior of each generative model per each caption in the test set.
+Instead of calculating error per each pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
+Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image.
+To calculate SSI, we sampled fifty images from the prior of each generative model for each caption in the test set.
 

### 2015-10-08 ram_workshop/ram-paper.tex
small errors

@@
 to include a language model, described in Sec.
-(\ref{sec:lang}), which process the conditioned caption.
+(\ref{sec:lang}), which processes the conditioned caption.
 Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of an Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and a Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
@@
 We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model.
-Unsurprisingly, the quality of image retrieval using generative models is worse than that of disriminative models that were specifically built for retrieval.
-To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on images having closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
+Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval.
+To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
 Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
@@
 
-Instead of calculating error per each pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
+Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
 Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image.
-To calculate SSI, we sampled fifty images from the prior of each generative model for each caption in the test set.
+To calculate SSI, we sampled fifty images from the prior of each generative model for every caption in the test set.
 

### 2015-10-08 ram_workshop/ram-paper.tex
update consolidate

--- 2015-10-08 ram_workshop/ram-paper.tex
@@
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
-%This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components. 
-Generative models of images studied previously, e.g. variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn},  
-often defined distributions that were restricted to being either unconditioned or conditioned on simple structured annotations, for example classification labels.
-%are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels. 
-Despite the recent advances in new methods of generative models, such as Variational Auto-Encoders (VAE) \cite{kingma_vae}\cite{gregor_draw} and Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, learning the highly structured natural image distribution in high-dimensional pixel space proves to be a difficult task.
-In real world circumstances, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
-%One domain presents substantial amount of relating information of the other domain. 
-The additional information from these descriptions could be used to simplify the image modelling task. %%
-Moreover, learning generative models conditioned on text allows a better understanding of the generalization performance of the model, as we can create textual descriptions of completely new scenes not seen at training time.
+This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components.
+Despite the advances in the variational generative models \cite{kingma_vae}\cite{gregor_draw} and the Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, image generative models studied previously, are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
+In real world, however, images rarely appear in isolation.
+They often accompanied by their unstructured textual descriptions on web pages and in books.
+The additional information from image and unstructured text description could be used to simplify the image modelling task.
+Also, learning conditional generative models from text also allows better inspection of the generalization performance of the model.
 
-\comm{There are two directions in learning a generative model of images conditioned text.  One approach is to learn a text generative model conditioned on the images.}
+\comm{
+Significant amount of recent works has been focused on generating captions from images \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
+By contrast, image understanding may also be studied by generating images correctly interpreting the text description.
+Generating high dimensional realistic images from their descriptions is a more difficult approach that combines two challenging components of language modelling and image generation.  
+}
 
-A significant amount of recent work has been focused on generating captions from images \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions}.
-A question that can be posed is whether the opposite is possible, i.e.
-generating an image given a caption.
-%By contrast, learning a generative model for images and text may also be studied by generating images correctly interpreting the text description. 
-Generating high dimensional realistic images from their descriptions is a more difficult problem as it requires both language modelling and image generation.
+In this paper, we address the problem of image generation from unstructured natural language captions.
+Extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw} architecture, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
+Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images corresponding to captions using soft attention mechanism.
+The images generated by our alignDRAW model are refined by post-processing deterministic Laplacian pyramid adversarial network \cite{denton_lapgan}.
+We then illustrate how our method, learnt on Microsoft COCO, generalizes to the captions describing novel scenarios that is previously unseen from the dataset.
 
+\comm{There are two directions in learning a generative model of image and text.  One approach is to learn a text generative model conditioned on the images.} 
 \comm{The models take an image descriptor and generate unstructured texts through a recurrent decoder.}
@@
 }
-
-In this paper, we address the problem of image generation from unstructured natural language captions.
-By extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw}, our model iteratively draws patches on a canvas, while attending to relevant words in the description.
-Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images using a soft attention mechanism attending to different input words while generating images.
-The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \cite{denton_lapgan}.
-We then illustrate how our method, learnt on the Microsoft COCO dataset, generalizes to captions describing novel scenarios that are not seen in the training set.
 
@@
 }
+
+\section{Model}
+
 \begin{figure}[!t]
@@
 \begin{center}
-\includegraphics[width=0.99\textwidth]{figures/alignDraw-cropped.pdf}\quad
+\includegraphics[width=0.8\textwidth]{figures/alignDraw-cropped.pdf}\quad
 %
@@
 \vspace{-0.1in}
-\section{Model}
-\vspace{-0.1in}
-Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
-Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $\oimage$ be the image corresponding to that caption.
+
+Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image. Our proposed model
+
+defines a generative process over multiple steps $\sum_\Lat\prior(\Lat)p(\oimage\given\icaption,\Lat)$ 
+can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+By using a set of stochastic latent variables $\Lat$, the conditional generative model learns a distribution over the output images $\oimage$ conditioned on the caption, $p(\oimage\given\icaption)$.
 
@@
 \label{sec:lang}
-The input caption sentences are fed into a deterministic Bidirectional LSTM\cite{} that encodes the variable size sentences into the vector representation $s$.
-Bidirectional LSTM consists of one Forward LSTM and Backward LSTM which combine information from past and future respectively.
-The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$.
-Then these hidden states are concatenated together into the sequence $[\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
+We obtain the caption sentence representation by first transform each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN\cite{}.
+In a Bidrectional RNN, the two LSTM process the input sequence from both forward and backward directions.
+The two LSTMs produces the hidden states sequences $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ and $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$ respectively.
+Then these hidden states are concatenated together into the final sentence representation $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
 
-\subsection{Image Generation: the Conditional DRAW Network}
+\subsection{Image Generation: the Conditional alginDRAW Network}
 \comm{
@@
 }
-We extended the architecture the DRAW network \citep{gregor_draw} generative process to include additional input caption from the language model described in Sec.
+To include the caption information, we extended the architecture the DRAW network \citep{gregor_draw} generative process to condition on the language model described in Sec.
 (\ref{sec:lang}).
-Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
-The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}:
+Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step.
+Unlike the original DRAW network where latent variables are independent unit Gaussian $\mathcal{N}(0, I)$, the align DRAW model uses latent variables whose mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
+In particular, the prior distribution $\prior(\Lat_t\given\Lat_{1:t-1})$ takes the form $\mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1}))$.
 
@@
 \begin{align}
+\vspace{-0.1in}
 \label{eq:x_hat}
@@
 \end{align}
-
 where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
+The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}.
 Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
 \begin{align}
+\vspace{-0.1in}
 s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.
@@
 
-The conditional generative model is trained to maximize the variational lower bound on the log-likelihood $\icaption$, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$ similar to \cite{kingma_vae}.
-The posterior inference is approximated by an inference network $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in red dash box in Figure (\ref{fig:figmodel}).
-Unlike the original DRAW network where latent variables are independent unit Gaussian $\mathcal{N}(0, I)$, the align DRAW model uses latent variables whose mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{}.
-In particular, the prior distribution $\prior(\Lat_t\given\Lat_{1:t-1})$ takes the form $\mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1}))$.
+Our conditional generative model is trained to maximize the variational lower bound on the log-likelihood $\icaption$, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$ similar to \cite{kingma_vae}.
+The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in red dash box in Figure (\ref{fig:figmodel}).
 
@@
 \begin{align}
-\loss_\theta =  &\expectation_{\post(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
+\vspace{-0.1in}
+\loss_\theta =  &\expectation_{\post(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{1:t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{1:t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
 \end{align}

### 2015-10-08 ram_workshop/ram-paper.tex

@@
 \section{Introduction}
-
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
-This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components.
 Despite the advances in the variational generative models \cite{kingma_vae}\cite{gregor_draw} and the Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, image generative models studied previously, are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
@@
 The additional information from image and unstructured text description could be used to simplify the image modelling task.
-Also, learning conditional generative models from text also allows better inspection of the generalization performance of the model.
+Moreover, learning generative models conditioned on text also allows better understand of the generalization performance of the model as we can create textual descriptions of completely new scenes not seen at training time.
 
@@
 In this paper, we address the problem of image generation from unstructured natural language captions.
-Extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw} architecture, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
-Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images corresponding to captions using soft attention mechanism.
-The images generated by our alignDRAW model are refined by post-processing deterministic Laplacian pyramid adversarial network \cite{denton_lapgan}.
+By extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw}, our model iteratively draws patches on canvas, while attending to the relevant words in the description.
+Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images from captions using a soft attention mechanism.
+The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \cite{denton_lapgan}.
 We then illustrate how our method, learnt on Microsoft COCO, generalizes to the captions describing novel scenarios that is previously unseen from the dataset.
@@
 defines a generative process over multiple steps $\sum_\Lat\prior(\Lat)p(\oimage\given\icaption,\Lat)$ 
-can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
 By using a set of stochastic latent variables $\Lat$, the conditional generative model learns a distribution over the output images $\oimage$ conditioned on the caption, $p(\oimage\given\icaption)$.
@@
 }
-To include the caption information, we extended the architecture the DRAW network \citep{gregor_draw} generative process to condition on the language model described in Sec.
+To include the caption information, we extended the architecture the DRAW network \citep{gregor_draw} to condition on the language model described in Sec.
 (\ref{sec:lang}).
@@
 The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in red dash box in Figure (\ref{fig:figmodel}).
-
 \comm{
@@
 We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model.
-As we expected, the quality of image retrieval using generative models is worse compared to the disriminative models that were specifically build for retrieval.
-To deal with large computational complexity of looping through each test image, we create a shortlist of hundred images including the correct one, based on the convolutional features of VGG like model trained on CIFAR dataset.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likehood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
-We found that the reconstruction error $\rloss$ increased for the sharpened images that considerably hurt the retrieval results.
+Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval.
+To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
+Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
+We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.
 Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.
 
-Instead of calculating error per each pixel, we turn to the smarter metric, such as Structural Similarity Index (SSI), which incorporates luminace and contrast masking into the error calculation.
-Due to strong inter-dependencies of closer pixels, the metric is calculated on the small windows of the image.
-To calculate SSI, we sampled fifty images from prior of each generative model per each caption in the test set.
+Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
+Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image.
+To calculate SSI, we sampled fifty images from the prior of each generative model for every caption in the test set.
 

### 2015-10-08 ram_workshop/ram-paper.tex
update

@@
 Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image. Our proposed model 
-
-defines a generative process over multiple steps $\sum_\Lat\prior(\Lat)p(\oimage\given\icaption,\Lat)$ 
-can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
-By using a set of stochastic latent variables $\Lat$, the conditional generative model learns a distribution over the output images $\oimage$ conditioned on the caption, $p(\oimage\given\icaption)$.
+defines a generative process of images conditioned the caption.
+In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+Our model can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}.
 
@@
 Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step.
-Unlike the original DRAW network where latent variables are independent unit Gaussian $\mathcal{N}(0, I)$, the align DRAW model uses latent variables whose mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
-In particular, the prior distribution $\prior(\Lat_t\given\Lat_{1:t-1})$ takes the form $\mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1}))$.
+Unlike the original DRAW network where latent variables are independent unit Gaussian $\mathcal{N}(0, I)$, the latent variables in the alignDRAW model have their mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
 
@@
 \label{eq:x_hat}
-\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1})\\
+\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1}))\\
 \hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1})\\

### 2015-10-08 ram_workshop/ram-paper.tex
updte

@@
 \label{sec:lang}
-We obtain the caption sentence representation by first transform each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN\cite{}.
+We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN\cite{}.
 In a Bidrectional RNN, the two LSTM process the input sequence from both forward and backward directions.
@@
 
-\subsection{Image Generation: the Conditional alginDRAW Network}
+\subsection{Image Modelling: the Conditional alginDRAW Network}
 \comm{
@@
 }
-To include the caption information, we extended the architecture the DRAW network \citep{gregor_draw} to condition on the language model described in Sec.
-(\ref{sec:lang}).
-Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step.
+To generate images conditioned on the caption information, we extended the DRAW network\citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure (\ref{fig:figmodel}).
+The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step.
 Unlike the original DRAW network where latent variables are independent unit Gaussian $\mathcal{N}(0, I)$, the latent variables in the alignDRAW model have their mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
-
 Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
@@
 
-Our conditional generative model is trained to maximize the variational lower bound on the log-likelihood $\icaption$, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$ similar to \cite{kingma_vae}.
+Our conditional alignDRAW model is trained to maximize the variational lower bound on the log-likelihood, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$.
 The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in red dash box in Figure (\ref{fig:figmodel}).
+Overall, the variational objective $\loss$ is defined with the model parameters vector $\theta_\theta$ as follows:
+\begin{align}
+\vspace{-0.1in}
+\loss_\theta =  &\expectation_{\post(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{1:t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{1:t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
+\end{align}
+
 \comm{
@@
 }
-Overall, the variational objective $\loss$ is defined with the model parameters vector $\theta$ as follows:
-\begin{align}
-\vspace{-0.1in}
-\loss_\theta =  &\expectation_{\post(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{1:t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{1:t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
-\end{align}
-
 \comm{
@@
 
-During the image generation step, we detach the inference network from the decoder and discard it, and instead sample from the prior distribution.
-Due to the blurriness of samples generated by the DRAW model, we do an additional post processing step, where we use the generator of an adversarial network trained on the residuals of a laplacian pyramid to sharpen the generated images,
-similar to \citep{denton_lapgan}.
-By fixing the prior of the adversarial generator to be the mean of the original uniform distribution, it gets treated as a deterministic neural network which allows us to calculate the lower bound of the likelihood.
-The reconstuction loss becomes the loss between sharpened image and correct image, whereas the latent loss stays the same.
+At the test time, images are generated by ancestral sampling the latent variables from the prior $\prior(\Lat_{1:T})$.
+Due to the blurriness of samples generated by the alignDRAW model, the blurry samples are fed to an additional post processing step.
+The generator of an adversarial network, trained independently on the residuals of a Laplacian pyramid, is used to sharpen the generated images from alignDRAW.
+The adversarial network is conditioned on the input caption using skip-thought vector\cite{skip-thought} and is trained similar to \citep{denton_lapgan}.
+Instead of sampling from its prior, we fix the adversarial generator to be the mean of the original uniform distribution.
+This post-processing step is a deterministic mapping which enable us to calculate the lower bound on the new log-likelihood defined at the output of the adversarial generator.
+Interestingly, we also found the deterministic procoess generate samples with much less noise than if we had sampled from the uniform distribution.
+
+\comm{.
+The reconstruction loss becomes the loss between sharpened image and correct image, whereas the latent loss stays the same.
 We also noticed that inputting the mean of the uniform distribution into the edge generator network allowed us to generate samples with much less noise than if we had sampled from the uniform distribution.
+}
 

### 2015-10-08 ram_workshop/ram-paper.tex
small errors

@@
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
-Despite the advances in the variational generative models \cite{kingma_vae}\cite{gregor_draw} and the Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, image generative models studied previously, are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
-In real world, however, images rarely appear in isolation.
-They often accompanied by their unstructured textual descriptions on web pages and in books.
-The additional information from image and unstructured text description could be used to simplify the image modelling task.
-Moreover, learning generative models conditioned on text also allows better understand of the generalization performance of the model as we can create textual descriptions of completely new scenes not seen at training time.
+Despite advances in variational generative models \cite{kingma_vae}\cite{gregor_draw} and Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, 
+generative models of images studied previously often defined distributions that were restricted to being either unconditioned or conditioned on simple structured annotations, for example classification labels.
+In real world applications, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books. 
+The additional information from these descriptions could be used to simplify the image modelling task.
+Moreover, learning generative models conditioned on text also allows a better understanding of the generalization performance of the model, as we can create textual descriptions of completely new scenes not seen at training time.
 
@@
 In this paper, we address the problem of image generation from unstructured natural language captions.
-By extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw}, our model iteratively draws patches on canvas, while attending to the relevant words in the description.
+By extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw}, our model iteratively draws patches on a canvas, while attending to the relevant words in the description.
 Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images from captions using a soft attention mechanism.
 The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \cite{denton_lapgan}.
-We then illustrate how our method, learnt on Microsoft COCO, generalizes to the captions describing novel scenarios that is previously unseen from the dataset.
+We then illustrate how our method, learnt on Microsoft COCO, generalizes to captions describing novel scenarios that are not seen in the dataset.
 

### 2015-10-08 ram_workshop/ram-paper.tex
my changes

--- 2015-10-08 ram_workshop/ram-paper.tex
@@
 \usepackage{xcolor}
+\usepackage{hyperref}
 %\usepackage{natbib}
@@
 
-\section{Experiments}
-% \subsection{MNIST With Captions}
-% As a toy experiment, we have trained our model on the two digit MNIST dataset with captions constructed on the fly.
-Two random digits were either placed horizontally or vertically, so that they don't overlap, on the black $60 \times 60$ background.
-The caption indicated the way digits were placed in the image and had the following template: \textit{the digit (1) is (2) the (3) of the digit (4) \textless eos\textgreater}.
-(1) and (4) were the digit numbers, (2) was either on or at, (3) was one of the adjectives top, bottom, left or right.
-For example, the generated images corresponding to the caption \textit{the digit seven is at the bottom of the digit five \textless eos\textgreater} as well as other captions are shown on~\ref{fig:fig1}.
-While most of the generative models were trained on the version of binarized MNIST dataset \citep{russ_binarizedmnist}, our model was trained directly on pixel intensities with binary cross entropy cost function.
-
-% \begin{figure}[t]
-% \label{fig1}
-% \captionsetup{labelformat=empty}
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-4.png}
-%   \caption{\textit{is digit 5}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-6.png}
-%   \caption{\textit{is digit 5}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-9.png}
-%   \caption{\textit{is digit 5}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-13.png}
-%   \caption{\textit{is digit 5}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-17.png}
-%   \caption{\mbox{\textit{eos 7 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-22.png}
-%   \caption{\mbox{\textit{eos 7 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-26.png}
-%   \caption{\mbox{\textit{eos 7 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-27.png}
-%   \caption{\mbox{\textit{eos 7 digit}}}
-% \endminipage\hfill
-
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-4.png}
-%   \caption{\mbox{\textit{eos digit 2}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-6.png}
-%   \caption{\mbox{\textit{eos digit 2}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-9.png}
-%   \caption{\mbox{\textit{eos digit 2}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-13.png}
-%   \caption{\mbox{\textit{eos digit 2}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-17.png}
-%   \caption{\mbox{\textit{of top digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-22.png}
-%   \caption{\mbox{\textit{of top digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-26.png}
-%   \caption{\mbox{\textit{of top digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-27.png}
-%   \caption{\mbox{\textit{of top digit}}}
-% \endminipage\hfill
-
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-4.png}
-%   \caption{\mbox{\textit{eos 0 of}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-6.png}
-%   \caption{\mbox{\textit{eos digit 0}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-9.png}
-%   \caption{\mbox{\textit{eos digit 0}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-12.png}
-%   \caption{\mbox{\textit{eos 0 of}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-17.png}
-%   \caption{\mbox{\textit{1 digit the}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-22.png}
-%   \caption{\mbox{\textit{1 digit the}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-25.png}
-%   \caption{\mbox{\textit{1 digit on}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-28.png}
-%   \caption{\mbox{\textit{digit the eos}}}
-% \endminipage\hfill
-
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-4.png}
-%   \caption{\textit{left 8 4}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-6.png}
-%   \caption{\textit{left 8 4}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-9.png}
-%   \caption{\mbox{\textit{left digit 4}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-14.png}
-%   \caption{\mbox{\textit{eos 8 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-18.png}
-%   \caption{\mbox{\textit{eos 8 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-22.png}
-%   \caption{\mbox{\textit{8 digit the}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-25.png}
-%   \caption{\mbox{\textit{8 digit digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-28.png}
-%   \caption{\mbox{\textit{digit digit the}}}
-% \endminipage\hfill
-
-% \captionsetup{labelformat=empty}
-% \caption{Figure 1: Four examples show the generated images unfolded over several timesteps as well as the top three words the model attends to while generating images.}
-% \end{figure}
-
-\subsection{COCO}
-
-Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
-The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on a caption very challenging.
-For consistency with related work on caption generation, we disregard four of the five captions when training our model.
-%Since some of the images have more than five captions attached to them,
+\vspace{-0.2cm}
+\section{Experiments on MS COCO dataset}
 
 In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models.
+To see a larger number of samples, we invite the reader to the following website \url{http://www.cs.toronto.edu/~emansim/cap2im.html}.
 
-\subsubsection{Analysis of Generated Images}
-The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
-and then use that knowledge to generate relevant images.
-To verify that, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
-
+\vspace{-0.2cm}
+\subsection{Analysis of Generated Images}
 First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
@@
 
-Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
-However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
+Despite some success with changing colors and backgrounds in descriptions, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
 In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
-This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
 
@@
 
-\subsubsection{Analysis of Attention}
+\vspace{-0.2cm}
+\subsection{Analysis of Attention}
 
-After flipping sets of words in the captions, we were curious to see which words the model attended to when generating images.
-It turned out that during the generation step, the model mostly focused on the specific words (or their neighbors) that carried the main semantic meaning expressed in the sentences.
+During the generation step, the model mostly focused on the specific words that carried the main semantic meaning expressed in the sentences.
 The attention values in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
-For example, in Figure~\ref{fig:attentimages1} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
-This suggests that, in their respective sentences, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
-In contrast, when we swap words ``beach'' and ``sun'', we can see a drastic change between sentences in the probability distribution over words.
-By noting that the model completely ignores the word ``sun'' in the second sentence, we can therefore gain a more thorough understanding of why we see no visual differences between the images generated by each caption.
-
+In Figure~\ref{fig:attentimages1} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
+Effectively, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
+In contrast, when we swap words ``beach'' and ``sun'', the model completely ignores the word ``sun'' in the second sentence, which gives us a more thorough understanding of why we see no visual differences between the images generated by each caption.
 We also tried to analyze the way the model generated images.
@@
 
-\subsubsection{Comparison With Other Models}
-Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffer from its own specific drawbacks.
-Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
+\vspace{-0.2cm}
+\subsection{Comparison With Other Models}
+Quantitatively evaluating generative models remains a challenging task in of itself as the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
 To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
 
-\begin{figure}[!t]
+\begin{figure}[!b]
 \captionsetup[subfigure]{labelformat=empty}
 \begin{center}
-\subfloat[Align DRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp.png}}\quad
+\subfloat[Align DRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
 %
-\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan.png}}\quad
+\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
 %
-\subfloat[Conv. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae.png}}\quad
+\subfloat[Conv. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
 %
-\subfloat[Fully-conn. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae.png}}\quad
+\subfloat[Fully-conn. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
 %
 \end{center}
-\caption{Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
+\caption{Four different models displaying results from sampling caption ``A group of people walk on a beach with surf boards"}
 \label{fig:diffmodels}
@@
 In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects,
-%don't have a clear structure in them
-whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
 
-As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods.
-We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model.
-Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval.
-To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
-We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.
+To compare performances of variational models, we rank the images conditioned on the captions in test set based on the lower bound of likelihood and then report the Precision-Recall metric.
+To avoid looping through each test image, we create a shortlist of 100 images including the correct one, based on the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
+To deal with ``easy'' images, we took a ratio of image likelihood conditioned on the caption to image likelihood conditioned on the mean train caption representation.
+We found that the reconstruction error $\rloss$ increased for sharpened images, which considerably hurt the retrieval results.
 Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.
@@
 Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
-Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image.
-To calculate SSI, we sampled fifty images from the prior of each generative model for every caption in the test set.
+The metric is also calculated on small windows of the image.
+Since all captions in the test set are independent from each other, we sampled 50 images from the prior of each generative model for every test caption and calculated SSI.
 
-\begin{table}[!b]
+\begin{table}[!t]
 \begin{center}

### 2015-10-08 ram_workshop/ram-paper.tex
More changes. Paper reduced to 4.5 pages. needs abstract

@@
 
+\nipsfinaltrue
 \captionsetup{font=small}
@@
 Elman Mansimov, Emilio Parisotto, Jimmy Lei Ba \& Ruslan Salakhutdinov\\
-Department of Computer Science\\
 University of Toronto\\
-Toronto, Ontario, Canada \\
-\texttt{\{emansim,eparisotto,rsalakhu\}@cs.toronto.edu}, \texttt{jimmy@psi.utoronto.ca}
 }
@@
 
+\vspace{-0.1in}
 \section{Introduction}
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
-Despite advances in variational generative models \cite{kingma_vae}\cite{gregor_draw} and Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, 
-generative models of images studied previously often defined distributions that were restricted to being either unconditioned or conditioned on simple structured annotations, for example classification labels.
+Generative models of images studied previously, such as Variational Generative Models \cite{kingma_vae}\cite{gregor_draw} and Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, often defined distributions that were restricted to being either unconditioned or conditioned on classification labels.
 In real world applications, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
@@
 In this paper, we address the problem of image generation from unstructured natural language captions.
-By extending the Deep Recurrent Attention Writer (DRAW)\cite{gregor_draw}, our model iteratively draws patches on a canvas, while attending to the relevant words in the description.
+By extending the Deep Recurrent Attention Writer (DRAW) \cite{gregor_draw}, our model iteratively draws patches on a canvas, while attending to the relevant words in the description.
 Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images from captions using a soft attention mechanism.
@@
 
+\vspace{-0.1in}
 \section{Model}
-
-\begin{figure}[!t]
-\vspace{-0.3in}
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/alignDraw-cropped.pdf}\quad
-%
-\end{center}
-\vspace{-0.2in}
-\caption{The image of the proposed model}
-\label{fig:figmodel}
-\vspace{-1cm}
-\end{figure}
-\vspace{-0.1in}
 
@@
 In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
-Our model can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}.
+Our model can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}.
 
+\vspace{-0.1in}
 \subsection{Language Representation: the Bidirectional Attention RNN}
 \label{sec:lang}
-We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN\cite{}.
-In a Bidrectional RNN, the two LSTM process the input sequence from both forward and backward directions.
-The two LSTMs produces the hidden states sequences $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ and $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$ respectively.
+We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN.
+In a Bidirectional RNN, the two LSTM process the input sequence from both forward and backward directions.
+They produce the hidden states sequences $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ and $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$ respectively.
 Then these hidden states are concatenated together into the final sentence representation $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
 
-\subsection{Image Modelling: the Conditional alginDRAW Network}
+\vspace{-0.3in}
+\subsection{Image Modelling: the Conditional alignDRAW Network}
 \comm{
@@
 }
-To generate images conditioned on the caption information, we extended the DRAW network\citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure (\ref{fig:figmodel}).
+To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}.
 The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step.
@@
 \subsection{Learning}
+
+\begin{figure}[!t]
+\vspace{-0.3in}
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/alignDraw-cropped.pdf}\quad
+%
+\end{center}
+\vspace{-0.2in}
+\caption{The image of the proposed model}
+\label{fig:figmodel}
+%\vspace{-1cm}
+\end{figure}
+%\vspace{-0.1in}
 
@@
 The generator of an adversarial network, trained independently on the residuals of a Laplacian pyramid, is used to sharpen the generated images from alignDRAW.
-The adversarial network is conditioned on the input caption using skip-thought vector\cite{skip-thought} and is trained similar to \citep{denton_lapgan}.
+The adversarial network is conditioned on the input caption using skip-thought vector \cite{skip-thought} and is trained similar to \citep{denton_lapgan}.
 Instead of sampling from its prior, we fix the adversarial generator to be the mean of the original uniform distribution.
@@
 
-\vspace{-0.2cm}
+\vspace{-0.1cm}
 \subsection{Analysis of Generated Images}

### 2015-10-08 ram_workshop/ram-paper.tex
now is 4 pages

@@
 
+\renewcommand{\baselinestretch}{0.95}
 \newcommand{\comm}[1]{}
@@
 
-\title{Generating Images From Captions\\ With Attention}
+\vspace*{-1.0cm}
+\title{Generating Images From Captions With Attention}
 
+\vspace*{-1.0cm}
 \author{
@@
 \maketitle
-
+\vspace*{-0.4cm}
 \begin{abstract}
-At the end.
+\vspace*{-0.15cm}
+Images rarely appear in isolation as they are often accompanied by textual description that could be used to simplify image generation.
+In this work, we introduce model that iteratively draws patches on a canvas, while attending to the relevant words in the description.
+We then illustrate how our method learnt on MS COCO, generalizes to captions describing novel scenarios that are not seen in dataset.
 \end{abstract}
@@
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
-Generative models of images studied previously, such as Variational Generative Models \cite{kingma_vae}\cite{gregor_draw} and Generative Adversarial Networks (GANs) \citep{goodfellow_gan}\cite{denton_lapgan}, often defined distributions that were restricted to being either unconditioned or conditioned on classification labels.
+Generative models of images studied previously often defined distributions that were restricted to being either unconditioned or conditioned on classification labels.
 In real world applications, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
@@
 We then illustrate how our method, learnt on Microsoft COCO, generalizes to captions describing novel scenarios that are not seen in the dataset.
+
+\comm{to the best of our knowledge this paper is the first to introduce the generative model of images conditioned on captions.}
 
@@
 
-\vspace{-0.3in}
+\vspace{-0.1in}
 \subsection{Image Modelling: the Conditional alignDRAW Network}
@@
 \hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1})\\
-s_{t} &= align(\hdec_{t-1}, \hlangall) \\
+s_{t} = align(\hdec_{t-1}, & \hlangall); \quad \quad
+canv_t = \canv_{t-1} + \writeop(\hdec_t)
 \label{eq:write}
-\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
+%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
 \end{align}
@@
 Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
-\begin{align}
-\vspace{-0.1in}
-s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.
-\end{align}
+%\begin{align}
+%\vspace{-0.1in}
+$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.$
+%\end{align}
 The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlang$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}
@@
 }
-\subsection{Learning}
 
-\begin{figure}[!t]
+\vspace{-0.2cm}
+\subsection{Learning and Generation}
+
+\begin{figure}[!h]
 \vspace{-0.3in}
@@
 \label{fig:figmodel}
-%\vspace{-1cm}
+\vspace{-0.2in}
 \end{figure}
@@
 \vspace{-0.1in}
+\small
 \loss_\theta =  &\expectation_{\post(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{1:t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{1:t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
@@
 }
-\subsection{Generating Images from Captions}
+%\subsection{Generating Images from Captions}
 
@@
 
-In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models.
-To see a larger number of samples, we invite the reader to the following website \url{http://www.cs.toronto.edu/~emansim/cap2im.html}.
-
-\vspace{-0.1cm}
-\subsection{Analysis of Generated Images}
+In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
+%\vspace{-0.1cm}
+%\subsection{Analysis of Generated Images}
 First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
@@
 
+% \vspace{-0.5cm}
 \begin{figure}[!h]
@@
 \label{fig:genimages1}
-\vspace{-0.3cm}
+\vspace{-0.6cm}
 \end{figure}
@@
 
+\vspace{-0.5cm}
 \begin{figure}[!h]
@@
 \label{fig:genimages2}
-\vspace{-0.3cm}
+\vspace{-0.4cm}
 \end{figure}
@@
 
-\vspace{-0.2cm}
-\subsection{Analysis of Attention}
+%\subsection{Analysis of Attention}
 
@@
 
+% \begin{table}[!t]
+% \small
+% \begin{center}
+% \begin{tabulary}{\linewidth}{c || c c c c c || c}
+% %\hline
+% %\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
+% %\hline
+% & \multicolumn{5}{c||}{Image Search} & Image Similarity \\
+% %\cline{2-6}
+% \textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
+% \hline
+% \hline
+% LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
+% \hline
+% F.-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+% Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+% Skipthought DRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+% Noalign DRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+% Align DRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+% \end{tabulary}
+% \end{center}
+% \vspace{-0.1cm}
+% \caption{Results of different models on COCO dataset (before sharpening).}
+% \vspace{-0.3cm}
+% \end{table}
+
+\vspace{-0.5cm}
 \begin{figure}[!h]
@@
 \label{fig:attentimages1}
-\vspace{-0.3cm}
+\vspace{-0.4cm}
 \end{figure}
@@
 \subsection{Comparison With Other Models}
-Quantitatively evaluating generative models remains a challenging task in of itself as the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
+To compare performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
 
-\begin{figure}[!b]
+\vspace{-0.5cm}
+\begin{figure}[!h]
 \captionsetup[subfigure]{labelformat=empty}
@@
 \end{center}
-\caption{Four different models displaying results from sampling caption ``A group of people walk on a beach with surf boards"}
+\caption{Four different models displaying results from sampling caption ``A group of people walk on a beach with surf boards".
+LAPGAN samples look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect.}
 \label{fig:diffmodels}
-\vspace{-0.2cm}
+\vspace{-0.4cm}
 \end{figure}
@@
 In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
 
@@
 To deal with ``easy'' images, we took a ratio of image likelihood conditioned on the caption to image likelihood conditioned on the mean train caption representation.
-We found that the reconstruction error $\rloss$ increased for sharpened images, which considerably hurt the retrieval results.
-Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.
+We found that the reconstruction error $\rloss$ increased for sharpened images, which considerably hurt the retrieval.
 
-Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
+In addition we calculate Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
 The metric is also calculated on small windows of the image.
@@
 
-\begin{table}[!t]
+\begin{table}[!h]
+\small
 \begin{center}
 \begin{tabulary}{\linewidth}{c || c c c c c || c}
-\hline
-\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
-\hline
+%\hline
+%\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
+%\hline
 & \multicolumn{5}{c||}{Image Search} & Image Similarity \\
@@
 \hline
-Fully-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+F.-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
 Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
@@
 \end{center}
+\vspace{-0.1cm}
+\caption{Results of different models on COCO dataset (before sharpening).}
+\vspace{-0.3cm}
 \end{table}

### 2015-10-08 ram_workshop/ram-paper.tex
fixed table reference

@@
 
-To compare performances of variational models, we rank the images conditioned on the captions in test set based on the lower bound of likelihood and then report the Precision-Recall metric.
+To compare performances of variational models, we rank the images conditioned on the captions in test set based on the lower bound of likelihood and then report the Precision-Recall metric (see Table~\ref{tab:results}).
 To avoid looping through each test image, we create a shortlist of 100 images including the correct one, based on the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
@@
 The metric is also calculated on small windows of the image.
-Since all captions in the test set are independent from each other, we sampled 50 images from the prior of each generative model for every test caption and calculated SSI.
+We sampled 50 images from the prior of each generative model for every test caption and calculated SSI, which is reported in Table~\ref{tab:results}.
 
@@
 \caption{Results of different models on COCO dataset (before sharpening).}
+\label{tab:results}
 \vspace{-0.3cm}

### 2015-10-08 ram_workshop/ram-paper.tex
spelling errors

@@
 Images rarely appear in isolation as they are often accompanied by textual description that could be used to simplify image generation.
-In this work, we introduce model that iteratively draws patches on a canvas, while attending to the relevant words in the description.
-We then illustrate how our method learnt on MS COCO, generalizes to captions describing novel scenarios that are not seen in dataset.
+In this work, we introduce a model that iteratively draws patches on a canvas, while attending to the relevant words in the description.
+After training on COCO, we then illustrate how our model generalizes to captions describing novel scenarios that are not seen in the dataset.
 \end{abstract}
@@
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
-Generative models of images studied previously often defined distributions that were restricted to being either unconditioned or conditioned on classification labels.
+Previously studied generative models of images often defined distributions that were restricted to being either unconditioned or conditioned on classification labels.
 In real world applications, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
@@
 Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image. Our proposed model 
-defines a generative process of images conditioned the caption.
+defines a generative process of images conditioned on the caption.
 In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
@@
 We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN.
-In a Bidirectional RNN, the two LSTM process the input sequence from both forward and backward directions.
+In a Bidirectional RNN, the two LSTMs process the input sequence from both forward and backward directions.
 They produce the hidden states sequences $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ and $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$ respectively.
-Then these hidden states are concatenated together into the final sentence representation $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
+These hidden states are then concatenated together into the final sentence representation $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
 
@@
 The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step.
-Unlike the original DRAW network where latent variables are independent unit Gaussian $\mathcal{N}(0, I)$, the latent variables in the alignDRAW model have their mean and variance depends on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
+Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
 Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
@@
 Our conditional alignDRAW model is trained to maximize the variational lower bound on the log-likelihood, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$.
-The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in red dash box in Figure (\ref{fig:figmodel}).
-Overall, the variational objective $\loss$ is defined with the model parameters vector $\theta_\theta$ as follows:
+The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in the red dashed box in Figure (\ref{fig:figmodel}).
+Overall, the variational objective $\loss$ is defined with the model parameters vector $\theta$ as follows:
 \begin{align}
@@
 At the test time, images are generated by ancestral sampling the latent variables from the prior $\prior(\Lat_{1:T})$. 
-Due to the blurriness of samples generated by the alignDRAW model, the blurry samples are fed to an additional post processing step.
-The generator of an adversarial network, trained independently on the residuals of a Laplacian pyramid, is used to sharpen the generated images from alignDRAW.
-The adversarial network is conditioned on the input caption using skip-thought vector \cite{skip-thought} and is trained similar to \citep{denton_lapgan}.
-Instead of sampling from its prior, we fix the adversarial generator to be the mean of the original uniform distribution.
-This post-processing step is a deterministic mapping which enable us to calculate the lower bound on the new log-likelihood defined at the output of the adversarial generator.
-Interestingly, we also found the deterministic procoess generate samples with much less noise than if we had sampled from the uniform distribution.
+%An additional post-processing step is performed to sharpen the often blurry images generated by the alignDRAW model. 
+The generator of an adversarial network, trained independently on the residuals of a Laplacian pyramid, is used to sharpen the generated images from alignDRAW, which are often blurry.
+The adversarial network is trained as in \citep{denton_lapgan} but is conditioned on the skip-thought vector representation \cite{skip-thought} of the input caption.
+Instead of sampling from its prior, we fix the input to the adversarial generator to be the mean of the original uniform distribution.
+This post-processing step is a deterministic mapping which enables us to calculate the lower bound on the new log-likelihood defined at the output of the adversarial generator.
+Interestingly, we found that the deterministic process generates samples with much less noise than if we had sampled from the uniform distribution.
 
@@
 The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
-Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
+Nevertheless, as shown in Figure~\ref{fig:genimages2}, changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
 The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.
@@
 We also tried to analyze the way the model generated images.
-Unfortunately, we found that there was no connection between the patches drawn on canvas and the most attended words at particular timesteps.
+Unfortunately, we found that there was no connection between the patches drawn on the canvas and the words with highest attention at particular timesteps.
 
@@
 
-To compare performances of variational models, we rank the images conditioned on the captions in test set based on the lower bound of likelihood and then report the Precision-Recall metric (see Table~\ref{tab:results}).
+To compare performances of variational models, we rank the images conditioned on the captions in the test set based on the lower bound of the likelihood and then report the Precision-Recall metric (see Table~\ref{tab:results}).
 To avoid looping through each test image, we create a shortlist of 100 images including the correct one, based on the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
-To deal with ``easy'' images, we took a ratio of image likelihood conditioned on the caption to image likelihood conditioned on the mean train caption representation.
+To deal with ``easy'' images, we took a ratio of image likelihood conditioned on the caption to image likelihood conditioned on the mean training caption representation.
 We found that the reconstruction error $\rloss$ increased for sharpened images, which considerably hurt the retrieval.

### 2015-10-09 ram_workshop/ram-paper.tex
a

@@
 \vspace*{-1.0cm}
-\title{Generating Images From Captions With Attention}
+\title{Generating Images From Captions}
 
@@
 The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ at each time step.
-Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
+Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
 Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
@@
 \end{align}
-where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
+where $\writeop$ and $\readop$ are the same attention operators as in \citep{gregor_draw}.
 The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}.
@@
 %\end{align}
-The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlang$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}
+The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}
 \comm{
@@
 \vspace{-0.2in}
-\caption{The image of the proposed model}
+\caption{alignDRAW: Generative model of images conditioned on captions.}
 \label{fig:figmodel}

### 2015-10-09 ram_workshop/ram-paper.tex
a

@@
 \graphicspath{{../}}
+
+\renewcommand{\baselinestretch}{0.97}
+
 \begin{document}
@@
 where $\writeop$ and $\readop$ are the same attention operators as in \citep{gregor_draw}.
-The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}.
+The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}.
 Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
@@
 %\end{align}
-The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}
+The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}.
 \comm{
@@
 \begin{figure}[!h]
+ \vspace{-0.5cm}
 \captionsetup[subfigure]{labelformat=empty}
@@
 
-Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
+Apart from changing the colors of objects, we experimented with changing the background of the scene described in a caption to see whether this would result in the appropriate changes in the generated samples.
 The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
@@
 \end{center}
-\caption{Bottom: Examples of changing the background while keeping the caption fixed. Top: The respective nearest training images based on pixel-wise L2 distance.}
+\caption{{\bf Bottom}: Examples of changing the background while keeping the caption fixed. {\bf Top}: The respective nearest training images based on pixel-wise L2 distance.}
 \label{fig:genimages2}
@@
 \begin{figure}[!h]
+\vspace{-0.5cm}
 \captionsetup[subfigure]{labelformat=empty}
@@
 
-To compare performances of variational models, we rank the images conditioned on the captions in the test set based on the lower bound of the likelihood and then report the Precision-Recall metric (see Table~\ref{tab:results}).
+To compare performances of variational models, we rank the images conditioned on the captions in the test set based on the lower bound of the log-probability and then report the Precision-Recall metric (see Table~\ref{tab:results}).
 To avoid looping through each test image, we create a shortlist of 100 images including the correct one, based on the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
-To deal with ``easy'' images, we took a ratio of image likelihood conditioned on the caption to image likelihood conditioned on the mean training caption representation.
-We found that the reconstruction error $\rloss$ increased for sharpened images, which considerably hurt the retrieval.
+To deal with ``easy'' images, we took a ratio of image
+likelihood conditioned on the caption to image likelihood conditioned on the mean training caption representation~\cite{kiros_captions}. 
+We found that the lower bound of the test log-probability decreased for sharpened images, which considerably hurt the retrieval.
 

### 2015-10-09 ram_workshop/ram-paper.tex
a
update

--- 2015-10-08 ram_workshop/ram-paper.tex
@@
 \vspace*{-0.15cm}
-Images rarely appear in isolation as they are often accompanied by textual description that could be used to simplify image generation.
-In this work, we introduce a model that iteratively draws patches on a canvas, while attending to the relevant words in the description.
-After training on COCO, we then illustrate how our model generalizes to captions describing novel scenarios that are not seen in the dataset.
+Motivated by the recent progress in generative models, we introduce a model that generates images from natural language descriptions.
+The proposed model iteratively draws patches on a canvas, while attending to the relevant words in the description.
+We empirically demonstrate our model generates images with novel scene composition corresponding to previously unseen captions.
+We then show our model outperforms other baseline generative models on image generation tasks.
+We further evaluated image retrieval performance of our proposed generative model on MSCOCO dataset.
 \end{abstract}

### 2015-10-09 ram_workshop/ram-paper.tex
fixed alignDraw figures

--- 2015-10-09 ram_workshop/ram-paper.tex
@@
 In this work, we introduce a model that iteratively draws patches on a canvas, while attending to the relevant words in the description.
-After training on COCO, we then illustrate how our model generalizes to captions describing novel scenarios that are not seen in the dataset.
+After training on COCO, we illustrate how our model generalizes to captions describing novel scenarios that are not seen in the dataset.
 \end{abstract}
@@
 \label{eq:x_hat}
-\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1}))\\
-\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1})\\
+\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
+\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1}), \\
 s_{t} = align(\hdec_{t-1}, & \hlangall); \quad \quad
-canv_t = \canv_{t-1} + \writeop(\hdec_t)
+canv_t = \canv_{t-1} + \writeop(\hdec_t0, 
 \label{eq:write}
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{figures/alignDraw-cropped.pdf}\quad
+\includegraphics[width=0.8\textwidth]{figures/alignDrawNew.pdf}\quad
 %
@@
 \begin{center}
-\subfloat[Align DRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
+\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
 %
@@
 \hline
-F.-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Fully-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
 Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-Skipthought DRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-Noalign DRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-Align DRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
 \end{tabulary}

### 2015-10-09 ram_workshop/ram-paper.tex

--- 2015-10-09 ram_workshop/ram-paper.tex
@@
 \vspace*{-0.15cm}
-Images rarely appear in isolation as they are often accompanied by textual description that could be used to simplify image generation.
-In this work, we introduce a model that iteratively draws patches on a canvas, while attending to the relevant words in the description.
-After training on COCO, we then illustrate how our model generalizes to captions describing novel scenarios that are not seen in the dataset.
+Motivated by the recent progress in generative models, we introduce a model that generates images from natural language descriptions.
+The proposed model iteratively draws patches on a canvas, while attending to the relevant words in the description.
+After training on MSCOCO, we compare our models with several baseline generative models on image generation and retrieval tasks.
+We then demonstrate our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.
 \end{abstract}

### 2015-10-09 ram_workshop/ram-paper.tex

--- 2015-10-09 ram_workshop/ram-paper.tex
@@
 \vspace*{-0.15cm}
-Images rarely appear in isolation as they are often accompanied by textual description that could be used to simplify image generation.
-In this work, we introduce a model that iteratively draws patches on a canvas, while attending to the relevant words in the description.
-After training on COCO, we illustrate how our model generalizes to captions describing novel scenarios that are not seen in the dataset.
+Motivated by the recent progress in generative models, we introduce a model that generates images from natural language descriptions.
+The proposed model iteratively draws patches on a canvas, while attending to the relevant words in the description.
+After training on MSCOCO, we compare our models with several baseline generative models on image generation and retrieval tasks.
+We then demonstrate our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.
 \end{abstract}

### 2015-10-09 ram_workshop/ram-paper.tex
update

@@
 \vspace*{-1.0cm}
-\title{Generating Images From Captions}
+\title{Generating Images from Captions with Attention}
 
@@
 s_{t} = align(\hdec_{t-1}, & \hlangall); \quad \quad
-canv_t = \canv_{t-1} + \writeop(\hdec_t0, 
+\canv_t = \canv_{t-1} + \writeop(\hdec_t), 
 \label{eq:write}
@@
 The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}.
-Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
+Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities $\alpha_{1...N}$:
 %\begin{align}
@@
 %An additional post-processing step is performed to sharpen the often blurry images generated by the alignDRAW model. 
-The generator of an adversarial network, trained independently on the residuals of a Laplacian pyramid, is used to sharpen the generated images from alignDRAW, which are often blurry.
-The adversarial network is trained as in \citep{denton_lapgan} but is conditioned on the skip-thought vector representation \cite{skip-thought} of the input caption.
+The generator of an adversarial network, trained independently as in \cite{denton_lapgan} on the residuals of a Laplacian pyramid, is used to sharpen the generated images from alignDRAW, which are often blurry.
+%The adversarial network is trained as in \citep{denton_lapgan} but is conditioned on the skip-thought vector representation \cite{skip-thought} of the input caption.
 Instead of sampling from its prior, we fix the input to the adversarial generator to be the mean of the original uniform distribution.

### 2015-10-09 ram_workshop/ram-paper.tex
update

@@
 
-\vspace*{-1.0cm}
+%\vspace*{-1.0cm}
 \title{Generating Images from Captions with Attention}
@@
 
-Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image. Our proposed model 
+Our proposed model 
 defines a generative process of images conditioned on the caption.
@@
 \label{sec:lang}
+Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image.
 We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN.
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{figures/alignDrawNew.pdf}\quad
+\includegraphics[width=0.8\textwidth]{figures/alignDrawAnnotated.pdf}\quad
 %
@@
 \end{center}
+\vspace{-0.3cm}
 \caption{Examples of changing the color while keeping the caption fixed.}
 \label{fig:genimages1}
-\vspace{-0.6cm}
+\vspace{-0.3cm}
 \end{figure}
-
 Apart from changing the colors of objects, we experimented with changing the background of the scene described in a caption to see whether this would result in the appropriate changes in the generated samples.
@@
 \end{center}
+\vspace{-0.3cm}
 \caption{{\bf Bottom}: Examples of changing the background while keeping the caption fixed. {\bf Top}: The respective nearest training images based on pixel-wise L2 distance.}
 \label{fig:genimages2}
-\vspace{-0.4cm}
+\vspace{-0.3cm}
 \end{figure}
@@
 \end{center}
+\vspace{-0.3cm}
 \caption{Examples of changing the object while keeping the caption fixed.}
@@
 
-\vspace{-0.5cm}
+%\vspace{-0.3cm}
 \begin{figure}[!h]
@@
 \end{center}
+\vspace{-0.3cm}
 \caption{Examples of most attended words while changing the background in the caption.}
@@
 \end{center}
-\caption{Four different models displaying results from sampling caption ``A group of people walk on a beach with surf boards".
+\vspace{-0.3cm}
+\caption{alignDRAW and three baseline models conditioned on skip-thoughts \cite{skip-thought} displaying results from sampling caption ``A group of people walk on a beach with surf boards".
 LAPGAN samples look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect.}
 \label{fig:diffmodels}
-\vspace{-0.4cm}
+\vspace{-0.3cm}
 \end{figure}
@@
 We sampled 50 images from the prior of each generative model for every test caption and calculated SSI, which is reported in Table~\ref{tab:results}.
-
+\vspace{-0.12in}
 \begin{table}[!h]
@@
 \end{table}
-
+\vspace{-0.3in}
 {\footnotesize

### 2015-10-09 ram_workshop/ram-paper.tex
update
a

@@
 \maketitle
-\vspace*{-0.4cm}
+\vspace*{-0.0cm}
 \begin{abstract}
@@
 The proposed model iteratively draws patches on a canvas, while attending to the relevant words in the description.
-After training on MSCOCO, we compare our models with several baseline generative models on image generation and retrieval tasks.
-We then demonstrate our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.
+After training on MS COCO, we compare our models with several baseline generative models on image generation and retrieval tasks.
+We demonstrate our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.
 \end{abstract}
@@
 defines a generative process of images conditioned on the caption.
-In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas $c_t$ over time $t=1,...,T$.
 Our model can be viewed as utilizing the  sequence-to-sequence framework \citep{ilya_mt}.
@@
 \subsection{Image Modelling: the Conditional alignDRAW Network}
+\vspace{-0.05in}
 \comm{
@@
 Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}.
-Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
+Formally, the image is generated by
+iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
 \begin{align}
@@
 
-\vspace{-0.2cm}
+\vspace{-0.05cm}
 \subsection{Learning and Generation}
+\vspace{-0.05in}
 
 \begin{figure}[!h]
-\vspace{-0.3in}
+\vspace{-0.45in}
 \captionsetup[subfigure]{labelformat=empty}
@@
 Our conditional alignDRAW model is trained to maximize the variational lower bound on the log-likelihood, $\log \sum_{\Lat_{1:T}} \prior(\Lat_{1:T})p(\oimage\given\icaption, \Lat_{1:T})$.
-The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in the red dashed box in Figure (\ref{fig:figmodel}).
+The posterior inference is approximated by an inference RNN $\post(\Lat_{1:T}\given\icaption,\oimage)$ shown in the red dashed box in Figure~\ref{fig:figmodel}.
 Overall, the variational objective $\loss$ is defined with the model parameters vector $\theta$ as follows:
@@
 \section{Experiments on MS COCO dataset}
-
+\vspace{-0.1in}
 In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
-%\vspace{-0.1cm}
-%\subsection{Analysis of Generated Images}
 First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
@@
 \begin{figure}[!h]
- \vspace{-0.5cm}
+ \vspace{-0.4cm}
 \captionsetup[subfigure]{labelformat=empty}
@@
 \end{center}
-\vspace{-0.3cm}
-\caption{Examples of changing the color while keeping the caption fixed.}
+\vspace{-0.2cm}
+\caption{Examples of changing the color while keeping the caption fixed. Best viewed in colour.}
 \label{fig:genimages1}
@@
 
-\vspace{-0.5cm}
+\vspace{-0.4cm}
 \begin{figure}[!h]
@@
 \vspace{-0.3cm}
-\caption{{\bf Bottom}: Examples of changing the background while keeping the caption fixed. {\bf Top}: The respective nearest training images based on pixel-wise L2 distance.}
+\caption{{\bf Bottom}: Examples of changing the background while keeping the caption fixed. {\bf Top}: The respective nearest training images based on pixel-wise L2 distance. Best viewed in colour.}
 \label{fig:genimages2}
@@
 \end{center}
-\vspace{-0.3cm}
+\vspace{-0.2cm}
 \caption{Examples of most attended words while changing the background in the caption.}
 \label{fig:attentimages1}
-\vspace{-0.4cm}
+\vspace{-0.3cm}
 \end{figure}
@@
 \subsection{Comparison With Other Models}
+\vspace{-0.1in}
 To compare performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
+In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
 
-\vspace{-0.5cm}
+\vspace{-0.3cm}
 \begin{figure}[!h]
@@
 \end{center}
-\vspace{-0.3cm}
+\vspace{-0.2cm}
 \caption{alignDRAW and three baseline models conditioned on skip-thoughts \cite{skip-thought} displaying results from sampling caption ``A group of people walk on a beach with surf boards".
@@
 \label{fig:diffmodels}
-\vspace{-0.3cm}
+\vspace{-0.2cm}
 \end{figure}
 
-In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
+%In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
 
@@
 We sampled 50 images from the prior of each generative model for every test caption and calculated SSI, which is reported in Table~\ref{tab:results}.
-\vspace{-0.12in}
+\vspace{-0.1in}
 \begin{table}[!h]

### 2015-10-16 iclr-paper-new.tex
New version is in iclr-paper-new.tex

--- 2015-10-07 iclr-paper.tex
@@
 
+\newcommand{\ssmall}[1]{{\scriptsize {#1}}}
+
 \newenvironment{myfont}{\fontfamily{[pcr]}\selectfont}{\par}
@@
 \begin{abstract}
-At the end.
+Motivated by the recent progress in generative models, we introduce a model that generates images from natural language descriptions.
+The proposed model iteratively draws patches on a canvas, while attending to the relevant words in the description.
+After training on Microsoft COCO, we compare our models with several baseline generative models on image generation and retrieval tasks.
+We demonstrate our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.
 \end{abstract}
@@
 Image generative models studied previously are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
-Despite the advances in generative models, learning the highly structured natural image distribution alone in the high dimensional pixel space alone proves to be a difficult task.
 In real world, however, images rarely appear in isolation.
-They often accompanied by their unstructured textual descriptions on web pages and in books.
-One domain presents substantial amount of relating information of the other domain.
+They are often accompanied by their unstructured textual descriptions on web pages and in books.
 The additional information from image and unstructured text description could be used to simplify the image modelling task.
+Moreover, learning generative models conditioned on text also allows a better understanding of the generalization performance of the model, as we can create textual descriptions of completely new scenes not seen at training time.
 
@@
 Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image.
-Although, the interesting high dimensional natural images lay on a small manifold that is difficult to capture, the additional text description cues of a target image may simplify the learning problem by focusing on the conditional distribution.
 
 In this paper, we illustrate how simple sequential deep learning techniques can be used to build a conditional probabilistic model over natural image space effectively.
-By using a sequence to sequence framework to approach the problem of image generation from unstructured natural language captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.
-Overall, the main contributions of this work are the following:
+By extending the Deep Recurrent Attention Writer (DRAW) \citep{gregor_draw}, our model iteratively draws patches on a canvas, while attending to the relevant words in the description.
+Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images from captions using a soft attention mechanism.
+The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \citep{denton_lapgan}.
+We then illustrate how our method, learnt on Microsoft COCO, generalizes to captions describing novel scenarios that are not seen in the dataset.
 
@@
 
-where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
+where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}, that apply a grid of Gaussian filters by specifying coordinates of the centre and focal length in order to yield a desirable image patch.
 Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
@@
 \section{Experiments}
-% \subsection{MNIST With Captions}
-% As a toy experiment, we have trained our model on the two digit MNIST dataset with captions constructed on the fly.
-Two random digits were either placed horizontally or vertically, so that they don't overlap, on the black $60 \times 60$ background.
-The caption indicated the way digits were placed in the image and had the following template: \textit{the digit (1) is (2) the (3) of the digit (4) \textless eos\textgreater}.
-(1) and (4) were the digit numbers, (2) was either on or at, (3) was one of the adjectives top, bottom, left or right.
-For example, the generated images corresponding to the caption \textit{the digit seven is at the bottom of the digit five \textless eos\textgreater} as well as other captions are shown on~\ref{fig:fig1}.
-While most of the generative models were trained on the version of binarized MNIST dataset \citep{russ_binarizedmnist}, our model was trained directly on pixel intensities with binary cross entropy cost function.
+\subsection{MNIST With Captions}
+As a toy experiment, we have trained our model on the MNIST dataset with captions constructed on the fly.
+The $60 \times 60$ image consisted of either two or one digits, where one digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image and two digits were either placed horizontally or vertically in nonoverlapping fashion.
+While most of the generative models were trained on the binarized version of MNIST dataset, our model was trained directly on pixel intensities with binary cross entropy cost function.
+The generated images and the attention alignment are displayed in Figure
 
-% \begin{figure}[t]
-% \label{fig1}
-% \captionsetup{labelformat=empty}
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-4.png}
-%   \caption{\textit{is digit 5}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-6.png}
-%   \caption{\textit{is digit 5}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-9.png}
-%   \caption{\textit{is digit 5}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-13.png}
-%   \caption{\textit{is digit 5}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-17.png}
-%   \caption{\mbox{\textit{eos 7 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-22.png}
-%   \caption{\mbox{\textit{eos 7 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-26.png}
-%   \caption{\mbox{\textit{eos 7 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/5-7-27.png}
-%   \caption{\mbox{\textit{eos 7 digit}}}
-% \endminipage\hfill
+\begin{figure}[!t]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\includegraphics[width=0.99\textwidth]{figures/new/mnist/prototype3.pdf}\quad
+%
+\end{center}
+\caption{Example MNIST}
+\label{fig:figmnist}
+\vspace{-0.3cm}
+\end{figure}
 
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-4.png}
-%   \caption{\mbox{\textit{eos digit 2}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-6.png}
-%   \caption{\mbox{\textit{eos digit 2}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-9.png}
-%   \caption{\mbox{\textit{eos digit 2}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-13.png}
-%   \caption{\mbox{\textit{eos digit 2}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-17.png}
-%   \caption{\mbox{\textit{of top digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-22.png}
-%   \caption{\mbox{\textit{of top digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-26.png}
-%   \caption{\mbox{\textit{of top digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/9-2-27.png}
-%   \caption{\mbox{\textit{of top digit}}}
-% \endminipage\hfill
+\subsection{Microsoft COCO}
 
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-4.png}
-%   \caption{\mbox{\textit{eos 0 of}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-6.png}
-%   \caption{\mbox{\textit{eos digit 0}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-9.png}
-%   \caption{\mbox{\textit{eos digit 0}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-12.png}
-%   \caption{\mbox{\textit{eos 0 of}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-17.png}
-%   \caption{\mbox{\textit{1 digit the}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-22.png}
-%   \caption{\mbox{\textit{1 digit the}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-25.png}
-%   \caption{\mbox{\textit{1 digit on}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/1-0-28.png}
-%   \caption{\mbox{\textit{digit the eos}}}
-% \endminipage\hfill
-
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-4.png}
-%   \caption{\textit{left 8 4}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-6.png}
-%   \caption{\textit{left 8 4}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-9.png}
-%   \caption{\mbox{\textit{left digit 4}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-14.png}
-%   \caption{\mbox{\textit{eos 8 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-18.png}
-%   \caption{\mbox{\textit{eos 8 digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-22.png}
-%   \caption{\mbox{\textit{8 digit the}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-25.png}
-%   \caption{\mbox{\textit{8 digit digit}}}
-% \endminipage\hfill
-% \minipage{0.1\textwidth}
-%   \includegraphics[width=\linewidth]{figures/4-8-28.png}
-%   \caption{\mbox{\textit{digit digit the}}}
-% \endminipage\hfill
-
-% \captionsetup{labelformat=empty}
-% \caption{Figure 1: Four examples show the generated images unfolded over several timesteps as well as the top three words the model attends to while generating images.}
-% \end{figure}
-
-\subsection{COCO}
-
-Microsoft COCO \citep{mscoco} is the largest dataset of images annotated with captions consisting of roughly 83k images.
-The rich collection of images with variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on caption very challenging.
-Since some of the images have more than five captions attached to them, for consistency with related work on caption generation we disregard extra captions.
-
-In the following subsections we make both qualitative and quantitative analysis of our model as well as compare its performance with the performances of other related generative models.
+Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
+The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on a caption very challenging.
+For consistency with related work on caption generation, we use only five captions when training our model.
+In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models.
 
 \subsubsection{Analysis of Generated Images}
-The main goal of this work is to learn a model that can understand the semantic meaning expressed in the descriptions of the images, such as the properties of objects, the relationships between them, etc.
+The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
 and then use that knowledge to generate relevant images.
-To verify that, we wrote a set of captions inspired by COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
+To verify that, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
 
 First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
-As shown in Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue.
-Although, there are images of buses with different colors in the training set, all school buses specifically are colored yellow.
-Despite that, the model managed to generate images of an object that is reminiscent of the bus pained with the relevant color.
+In Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue.
+Although, there are images of buses with different colors in the training set, all mentioned school buses are specifically colored yellow.
+Despite that, the model managed to generate images of an object that is visually reminiscent of a school bus that is painted with the specified color.
 
-\begin{figure}[!t]
+\begin{figure}[!h]
 \captionsetup[subfigure]{labelformat=empty}
@@
 \end{center}
-\caption{Examples of changing the background while keeping the caption fixed. The respective nearest training images based on pixelwise distance are displayed on top.}
+\caption{\textbf{Bottom}: Examples of changing the background while keeping the caption fixed. \textbf{Top}: The respective nearest training images based on pixel-wise L2 distance.}
 \label{fig:genimages2}
@@
 
-Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar patter was present in the training set.
-However, it wasn't always the case when changing an object itself in the description.
-As you can see in Figure~\ref{fig:genimages3}, when objects didn't have a clear fine grained differences, such as different shape or color, the relevant changes in the generated samples weren't very clearly seen.
-This highlighted the limitation of the model to grasp the detailed understanding of each object.
+Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
+However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
+In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
+This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
 
@@
 
+Aside from flipping words in the captions that describe the common situations, we were curious about trying nonsense captions that model has never seen in the training phase.
+These captions described some object, like ``elephants'', doing something in the completely new environment, like ``flying in the blue skies''.
+That test shows whether the model truly generalizes to completely new situations or overfits on the scenes that the model seen during training.
+The model managed to generalize to new situations, as shown in Figure~\ref{fig:genimages4}.
+
+\begin{figure}[!h]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\subfloat[A yellow school bus is flying in blue skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-is-flying-in-blue-skies-sharp.png}}\quad
+%
+\subfloat[A herd of elephants flying in the blue skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
+%
+\subfloat[A toilet seat sits open in the grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
+%
+\subfloat[A person skiing on sand clad vast desert.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\quad
+%
+\end{center}
+\caption{Examples of novel scene compositions.}
+\label{fig:genimages4}
+\vspace{-0.3cm}
+\end{figure}
+
 \subsubsection{Analysis of Attention}
-Unfortunately, we found that there was no connection between the patches drawn on canvas and most attended words at each timestep.
-During the image generation, the model mostly focused on several words that carried the semantic meaning of caption.
-The words which were mostly attended during all generation steps indicated the kind of scene the model would generate.
-For example, as shown in Figure~\ref{fig:attentimages1} by equally looking at words deset and forest allowed the model to make relevant changes in the scene.
-Whereas in the second example, the model completely ignored word sun and didn't make any changes.
+
+After flipping sets of words in the captions, we were curious to see which words the model attended to when generating images.
+It turned out that during the generation step, the model mostly focused on the specific words (or their neighbors) that carried the main semantic meaning expressed in the sentences.
+The attention values in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
+For example, in Figure~\ref{fig:attentimages1} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
+This suggests that, in their respective sentences, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
+In contrast, when we swap words ``beach'' and ``sun'', we can see a drastic change between sentences in the probability distribution over words.
+By noting that the model completely ignores the word ``sun'' in the second sentence, we can therefore gain a more thorough understanding of why we see no visual differences between the images generated by each caption.
+
+We also tried to analyze the way the model generated images.
+Unfortunately, we found that there was no connection between the patches drawn on canvas and the most attended words at particular timesteps.
 
@@
 \subsubsection{Comparison With Other Models}
-The quantitative evaluation of generative models has been a subject of ambiguity in a machine learning community.
+Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffer from its own specific drawbacks.
 Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To get a better comparisions between performances of generative models, we report results on two different metrics as well as do some qualitative comparison of different generative models.
+To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
 
@@
 
-As you can see on Figure~\ref{fig:diffmodels}, we generated several samples from prior of each of the current state-of-the-art generative models corresponding to the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and don't have a very clear structure in them, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
+In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects,
+%don't have a clear structure in them
+whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
 
@@
 We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model.
-As we expected, the quality of image retrieval using generative models is worse compared to the disriminative models that were specifically build for retrieval.
-To deal with large computational complexity of looping through each test image, we create a shortlist of hundred images including the correct one, based on the convolutional features of VGG like model trained on CIFAR dataset.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likehood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
-We found that the reconstruction error $\rloss$ increased for the sharpened images that considerably hurt the retrieval results.
+Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval.
+To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
+Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
+We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.
 Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.
 
-Instead of calculating error per each pixel, we turn to the smarter metric, such as Structural Similarity Index (SSI), which incorporates luminace and contrast masking into the error calculation.
-Due to strong inter-dependencies of closer pixels, the metric is calculated on the small windows of the image.
-To calculate SSI, we sampled fifty images from prior of each generative model per each caption in the test set.
+Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
+Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image.
+To calculate SSI, we sampled fifty images from the prior of each generative model for every caption in the test set.
 
-\begin{table}[!t]
+\begin{table}[!b]
 \begin{center}
@@
 Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-Skipthought DRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-Noalign DRAW & 2.8 & \textbf{14.1} & \textbf{23.1} & 68.0 & \textbf{31} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-Align DRAW & \textbf{3.0} & 14.0 & 22.9 & \textbf{68.5} & \textbf{31} & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+Skipthought DRAW & 2.6 & 12.6 & 21.3 & 66.9 & 33 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+Noalign DRAW & 3.2 & \textbf{16.3} & \textbf{26.0} & 72.5 & \textbf{27} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+Align DRAW & \textbf{3.6} & 16.1 & \textbf{26.0} & \textbf{72.8} & 28 & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
 \end{tabulary}
@@
 
+\newpage
+\appendix
+\input{supp.tex}
+
 \end{document}

### 2015-10-16 supp.tex
New version is in iclr-paper-new.tex

\section{Training Details}
\label{sec:training_details}

\subsection{Hyperparameters}

Each parameter in alignDRAW was initialized by sampling from the Gaussian distribution of mean $0$ and standard deviation of $0.01$. All models were trained with \textit{RMSProp} with initial learning rate of $0.001$. The norm of the gradients was rescaled to $10.0$ during training. For all tasks we used $128$ units in language model and $512$ units in alignment multi-layer perceptron.

% \begin{table}[!h]
% \begin{center}
% \begin{tabulary}{\linewidth}{c | c c c}
% \hline
% \multicolumn{4}{c}{\textbf{Language Model}} \\
% \hline
% Task & Vocab size & LSTM \#$h$ & Align \#$h$\\
% \hline
% MNIST With Captions & 22 & 128 & 512  \\
% Microsoft COCO & 25323 & 128 & 512
% \end{tabulary}
% \end{center}
% \end{table}

\begin{table}[!h]
\begin{center}
\begin{tabulary}{\linewidth}{c | c c c c c c}
\hline
\multicolumn{7}{c}{\textbf{Image Model}} \\
\hline
Task & \#glimpses & Inference \#$h$ & Generative \#$h$ & \#$z$ & Read Size & Write Size\\
\hline
MNIST Captions & 32 & 550 & 550 & 275 & 9 & 9\\
MS COCO & 32 & 300 & 300 & 150 & 8 & 8\\
\end{tabulary}
\end{center}
\end{table}

\subsection{Evaluation}
\begin{table}[!h]
\begin{center}
\begin{tabulary}{\linewidth}{c | c c c c}
\hline
Model & Train NLL & Valid NLL & Test NLL & Test NLL (after sharpening)\\
\hline
skipthoughtDRAW & 1794.29 & 1797.41 & 1791.37 & 2045.84 \\
noalignDRAW & 1792.14 & 1796.94 & 1791.15 & 2051.07 \\
alignDRAW & 1792.15 & 1797.24 & 1791.53 & 2042.31
\end{tabulary}
\end{center}
\end{table}

### 2015-10-19 iclr-paper-new.tex
New version is in iclr-paper-new.tex
Updated couple of figures. Added new sentences

@@
 \begin{center}
-\includegraphics[width=0.99\textwidth]{figures/alignDraw-cropped.pdf}\quad
+\includegraphics[width=0.99\textwidth]{figures/alignDrawAnnotated.pdf}\quad
 %
@@
 \begin{center}
-\includegraphics[width=0.99\textwidth]{figures/new/mnist/prototype3.pdf}\quad
+\includegraphics[width=0.99\textwidth]{figures/new/mnist/test3.pdf}\quad
 %
 \end{center}
-\caption{Example MNIST}
+\caption{Examples of generating $60 \times 60$ images corresponding to respective captions.
+Examples at the \textbf{left column} were present during training, whereas the \textbf{right column} examples were hidden during training.}
 \label{fig:figmnist}
@@
 \begin{center}
-\subfloat[A yellow school bus is flying in blue skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-is-flying-in-blue-skies-sharp.png}}\quad
+\subfloat[A stop sign is flying in blue skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
 %

### 2015-10-21 iclr-paper-new.tex
new changes

@@
 \subsection{MNIST With Captions}
-As a toy experiment, we have trained our model on the MNIST dataset with captions constructed on the fly.
-The $60 \times 60$ image consisted of either two or one digits, where one digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image and two digits were either placed horizontally or vertically in nonoverlapping fashion.
+As a first experiment, we trained our proposed model on the MNIST dataset with captions constructed on the fly.
+Either one or two digits from MNIST training dataset were placed on the $60 \times 60$ blank image.
+One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image.
+Two digits were either placed horizontally or vertically in nonoverlapping fashion.
 While most of the generative models were trained on the binarized version of MNIST dataset, our model was trained directly on pixel intensities with binary cross entropy cost function.
-The generated images and the attention alignment are displayed in Figure
+The generated images and the attention alignment are displayed in Figure~\ref{fig:figmnist}.
 
@@
 
-Aside from flipping words in the captions that describe the common situations, we were curious about trying nonsense captions that model has never seen in the training phase.
-These captions described some object, like ``elephants'', doing something in the completely new environment, like ``flying in the blue skies''.
-That test shows whether the model truly generalizes to completely new situations or overfits on the scenes that the model seen during training.
-The model managed to generalize to new situations, as shown in Figure~\ref{fig:genimages4}.
+As a real test of model generalization, we were tried generating images corresponding to captions that describe scenarios that never occur in real life.
+These captions describe some common object doing unusual things considering the physics of real world.
+Even though these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
+Nevertheless, as you can see in Figure~\ref{fig:genimages4}, the model managed to generate reasonable images.
 
@@
 
+We compared the performance of the proposed model to the DRAW model conditioned on captions without align function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors (skipthoughtDRAW).
+All of the conditional DRAW models were trained with binary cross entropy cost function.
+We also compared our model with Fully-connected (FC) and Convolutional-Deconvolutional (Conv.) Variational Autoencoders which were trained with the least squares cost function.
+The LAPGAN model was trained on two level Laplacian Pyramid and was conditioned on the skipthought vectors.
+
 In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects,
-%don't have a clear structure in them
-whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
 
 As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods.
-We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model.
+We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model (see Table~\ref{tab:results}.).
 Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval.
 To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set.
+Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set\cite{kiros_captions}.
 We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.
@@
 Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image.
-To calculate SSI, we sampled fifty images from the prior of each generative model for every caption in the test set.
+Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI.
+The results are reported in Table~\ref{tab:results}.
 
@@
 
+\section{Discussion}
+
+We have shown that our proposed model manages to generate reasonable images corresponding to respective captions as well as generalize to completely novel scenarios that never occur in real life.
+Sequential generation of images using attention played a crutial part in it.
+However the DRAW model still underfits on training data which leads to the blurry samples.
+We think that extending this process by generating very small sections of images like superpixels over potentially hundreds of timesteps might help the model fit the data and generate sharp images.
+
+\section{Acknowledgements}
+We would like to thank developers of Theano for a powerful software and authors of \cite{denton_lapgan} for open sourcing their code.
+We would also like to thank Ryan Kiros and Nitish Srivastava for helpful discussions.
+We will release our code with pretrained models as soon as we clean it.
+Acknowledging funding (Russ)
+
 \bibliography{iclr-paper}

### 2015-10-22 iclr-paper-new.tex
fixed errors

@@
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
-This has motivated recent approaches in generative modelling applied to natural images by employing deep neural networks for their inference and generative components.
-Image generative models studied previously are often restricted to learning unconditional models of images distribution or conditioned on simple structured annotations, for example classification labels.
-In real world, however, images rarely appear in isolation.
-They are often accompanied by their unstructured textual descriptions on web pages and in books.
-The additional information from image and unstructured text description could be used to simplify the image modelling task.
+The challenging nature of this task has motivated recent approaches to exploit the inference and generative capabilities of deep neural networks.
+Previously studied generative models of images often defined distributions that were restricted to being either unconditioned or conditioned on classification labels.
+In real world applications, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
+The additional information from these descriptions could be used to simplify the image modelling task.
 Moreover, learning generative models conditioned on text also allows a better understanding of the generalization performance of the model, as we can create textual descriptions of completely new scenes not seen at training time.
 
-There are two directions in learning a generative model of image and text.
-One approach is to learn a text generative model conditioned on the images.
-Significant amount of recent works has been focused on generating captions from images \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.
-The models take an image descriptor and generate unstructured texts through a recurrent decoder.
-By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description.
-Generating high dimensional realistic images from their descriptions is a more difficult approach that combines two challenging components of language modelling and image generation.
-Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image.
+%There are two primary directions in learning a generative model of image and text.  
+There are numerous ways to learn a generative model over both image and text modalities. One approach is to learn a generative model of text conditioned on images, known as caption generation
+%A significant amount of recent work has focused on generating captions from images
+\citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions}.
+These models take an image descriptor and generate unstructured texts using a recurrent decoder.
+In contrast, we will explore models that condition in the opposite direction, i.e.
+taking textual descriptions as input and using them to generate relevant images.
+%By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description. 
+Generating high dimensional realistic images from their descriptions 
+%is a more difficult approach that 
+combines the two challenging components of language modelling and image generation, and can be considered to be more difficult than caption generation.
+%Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image.
 
-In this paper, we illustrate how simple sequential deep learning techniques can be used to build a conditional probabilistic model over natural image space effectively.
+In this paper, we illustrate how sequential deep learning techniques can be used to build a conditional probabilistic model over natural image space effectively.
 By extending the Deep Recurrent Attention Writer (DRAW) \citep{gregor_draw}, our model iteratively draws patches on a canvas, while attending to the relevant words in the description.
@@
 
-Deep Neural Networks have achieved a remarkable performance in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech} and etc.
-While most of the recent success has been achieved by discriminative models, the generative models have not yet enjoyed the same level of success.
-Most of the previous work in generative models has been focused on variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}.
-While these models are very powerful, each iteration of training requires a computationally costly step of MCMC to approximate an intractable normalization constant that makes it hard to scale them to large datasets.
+Deep Neural Networks have achieved significant success in various tasks such as image recognition \citep{krizhevsky_imagenet} and speech transcription \citep{graves_speech}. 
+While most of the recent success has been achieved by discriminative models, generative models have not yet enjoyed the same level of success.
+Most of the previous work in generative models has focused on variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}.
+While these models are very powerful, each iteration of training requires a computationally costly step of MCMC to approximate an intractable normalization constant, making it difficult to scale them to large datasets.
 
 \cite{kingma_vae} have introduced the Variational Auto-Encoder (VAE) which can be seen as a neural network with continous latent variables.
-The encoder is used to approximate posterior distribution and the decoder is used to stochastically reconstruct the data from latent variables.
-The model performs an efficient inference and learning that allows it to scale to large datasets.
-\cite{gregor_draw} have introduced Deep Recurrent Attention Writer (DRAW), where they have incorporated a novel differentiable attention mechanism into the VAE which significantly improved its performance as well as quality of generated samples.
-While most of the samples of VAE and DRAW resemble a clear structure of objects, the generated images are blurry most of the time.
+The encoder is used to approximate a posterior distribution and the decoder is used to stochastically reconstruct the data from latent variables.
+The model's efficient inference and learning procedures allow it to scale to large datasets.
+\cite{gregor_draw} introduced the Deep Recurrent Attention Writer (DRAW), extending the VAE approach by incorporating a novel differentiable attention mechanism which significantly improved performance.% as well as quality of generated samples.
+%While most of the samples from VAE and DRAW models resemble a clear structure of objects, the generated images are blurry most of the time.
 
-Generative Adversarial Networks (GANs) \citep{goodfellow_gan} are another type of generative models that use noise-contrastive estimation \citep{gutmann_nce} to avoid calculating intractable normalization constant.
-The model consists of generator that generates samples from uniform distribution and discriminator that discriminates between real and generated images.
-Both networks are playing the game, where generator tries to produce samples that look real and discriminator tries not to be fooled by generator.
-Recently, \cite{denton_lapgan} have scaled those models, by training GANs at each level of Laplacian pyramid of images.
-While their model has generated sharp looking samples, the generated images have lacked a clear structure.
-Compared to other mentioned generative models, GANs are more unstable and are harder to train.
+Generative Adversarial Networks (GANs) \citep{goodfellow_gan} are another type of generative models that use noise-contrastive estimation \citep{gutmann_nce} to avoid calculating an intractable normalization constant.
+The model consists of a generator that generates samples using a uniform distribution and a discriminator that discriminates between real and generated images.
+Both networks can be seen as playing a game against each other, where the generator tries to produce samples that look real and the discriminator tries not to be fooled by the generator. 
+Recently, \cite{denton_lapgan} have scaled those models by training conditional GANs at each level of a Laplacian pyramid of images. 
+%While their model has generated sharp looking samples, the generated images have lacked a clear structure. 
+%Compared to other mentioned generative models, GANs are more unstable and are harder to train.
 
-While all of the previous work has been focused on unconditional models or models conditioned on labels, to the best of our knowledge this paper is the first to introduce the generative model of images conditioned on captions.
+While all of the previous work has been focused on unconditional models or models conditioned on labels, to the best of our knowledge this paper is the first to introduce a generative model of images conditioned on captions.
 
@@
 \section{Model}
-Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+Our proposed model 
+defines a generative process of images conditioned on a caption.
+In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on a canvas $c_t$ over time $t=1,...,T$.
+Our model can be viewed as utilizing the sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}.
+%Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
 Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $\oimage$ be the image corresponding to that caption.
@@
 The DRAW network \cite{gregor_draw} is a sequential probabilistic model generating images by accumulating the output at each iterative step.
-While the original DRAW network assumes the latent variables are independent, it has shown in \citep{bachman_sdm} the model performance is improved by including the dependencies of latent variables.
+While the original DRAW network assumes the latent variables are independent, it has been shown in \citep{bachman_sdm} that model performance is improved by including dependencies between latent variables.
 
-We extended the architecture the DRAW network generative process to include additional input caption from the language model described in Sec.
+We extended the architecture of the DRAW network to include an additional input caption from the language model described in Sec.
 (\ref{sec:lang}).
-Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$.
+Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of an Inference LSTM that infers the distribution of latent variables of image $x$ given caption $y$ and then a Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given caption $y$, where $\icaption$ is the input caption consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$.
 The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}:
@@
 
-where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}, that apply a grid of Gaussian filters by specifying coordinates of the centre and focal length in order to yield a desirable image patch.
+where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
+% which apply a grid of Gaussian filters by specifying coordinates of the location and scale in order to yield a desirable image patch.
 Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
@@
 The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}.
-The model is trained to maximize the lower bound of marginal likelihood $\loss$ of the correct image $x$ given the input caption $y$.
+The model is trained to maximize a lower bound of the marginal likelihood $\loss$ of the correct image $x$ given the input caption $y$.
 The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$.
@@
 
-During the image generation step, we throw away the inference network from the decoder and instead sample from the prior distribution.
-Due to the blurriness of samples generated by DRAW model, we do an additional post processing step, where we use the generator of the adversarial network trained on residuals of laplacian pyramid to sharpen the generated images, similar to \citep{denton_lapgan}.
-By fixing the prior of adversarial generator to the mean of uniform distribution, it gets treated as a deterministic neural network which allows us to calculate the lower bound of likelihood.
+During the image generation step, we discard the inference network and instead sample from the prior distribution. 
+Due to the blurriness of samples generated by the DRAW model, we perform an additional post processing step where we use a 
+adversarial network trained on residuals of a laplacian pyramid 
+to sharpen the generated images, similar to \citep{denton_lapgan}.
+By fixing the prior of the adversarial generator to its mean, it gets treated as a deterministic neural network which allows us to calculate the lower bound of likelihood.
 The reconstuction loss becomes the loss between sharpened image and correct image, whereas the latent loss stays the same.
-We also noticed that sampling from the mean of uniform distribution allowed us to generate much less noisy samples than by sampling from the uniform distribution itself.
+Additionally, we noticed that sampling from the mean of uniform distribution generated samples with less noise than if we had otherwise sampled from the uniform distribution itself.
 
@@
 \subsection{MNIST With Captions}
-As a first experiment, we trained our proposed model on the MNIST dataset with captions constructed on the fly.
-Either one or two digits from MNIST training dataset were placed on the $60 \times 60$ blank image.
+As a first experiment, we trained our proposed model on the MNIST dataset with artificial captions.
+Either one or two digits from the MNIST training dataset were placed on a $60 \times 60$ blank image.
 One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image.
 Two digits were either placed horizontally or vertically in nonoverlapping fashion.
-While most of the generative models were trained on the binarized version of MNIST dataset, our model was trained directly on pixel intensities with binary cross entropy cost function.
+While many generative models are trained on the binarized version of the MNIST dataset, we trained directly on pixel intensities with a binary cross entropy cost function. 
 The generated images and the attention alignment are displayed in Figure~\ref{fig:figmnist}.
@@
 Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
-The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model conditioned on a caption very challenging.
+The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model
+%conditioned on a caption 
+very challenging.
 For consistency with related work on caption generation, we use only five captions when training our model.
@@
 and then use that knowledge to generate relevant images.
-To verify that, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
+To examine the understanding of our model, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
 
@@
 
-Despite an infinite number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
+Despite a large number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
 However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
@@
 
-As a real test of model generalization, we were tried generating images corresponding to captions that describe scenarios that never occur in real life.
-These captions describe some common object doing unusual things considering the physics of real world.
+As a test of model generalization, we tried generating images corresponding to captions that describe scenarios that are highly unlikely to occur in real life.
+These captions describe some common object doing unusual things
+or in a strange location.
+%that are impossible in the real world. 
+%considering the physics of real world. 
 Even though these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
@@
 We also tried to analyze the way the model generated images.
-Unfortunately, we found that there was no connection between the patches drawn on canvas and the most attended words at particular timesteps.
+Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular timesteps.
 
@@
 \subsubsection{Comparison With Other Models}
-Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffer from its own specific drawbacks.
+Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffers from its own specific drawbacks.
 Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
@@
 As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods.
-We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model (see Table~\ref{tab:results}.).
+We rank the images in the test set conditioned on the captions based on the the variational lower bound of likelihood and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
 Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval.

### 2015-10-25 iclr-paper-new.tex
update learning

@@
 
+\section{Model}
+Our proposed model 
+defines a generative process of images conditioned on a caption.
+In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on a canvas $c_t$ over time $t=1,...,T$.
+Our model can be viewed as utilizing the sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}.
+%Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
+Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $\oimage$ be the image corresponding to that caption.
+
+\subsection{Language Model: the Bidirectional Attention RNN}
+\label{sec:lang}
+Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image.
+We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN.
+In a Bidirectional RNN, the two LSTMs process the input sequence from both forward and backward directions.
+The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$.
+These hidden states are then concatenated together into the sequence $[\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
+
 \begin{figure}[!t]
@@
 
-\section{Model}
-Our proposed model 
-defines a generative process of images conditioned on a caption.
-In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on a canvas $c_t$ over time $t=1,...,T$.
-Our model can be viewed as utilizing the sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}.
-%Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
-Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $\oimage$ be the image corresponding to that caption.
-
-\subsection{Language Model: the Bidirectional Attention RNN}
-\label{sec:lang}
-The input caption sentences are fed into a deterministic Bidirectional LSTM\cite{} that encodes the variable size sentences into the vector representation $s$.
-Bidirectional LSTM consists of one Forward LSTM and Backward LSTM which combine information from past and future respectively.
-The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$.
-Then these hidden states are concatenated together into the sequence $[\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
-
 \subsection{Image Model: the Conditional DRAW Network}
 
-The DRAW network \cite{gregor_draw} is a sequential probabilistic model generating images by accumulating the output at each iterative step.
-While the original DRAW network assumes the latent variables are independent, it has been shown in \citep{bachman_sdm} that model performance is improved by including dependencies between latent variables.
+To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}.
+The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ and accumulates the output at each iterative step..
 
+Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
+It has been shown in \citep{bachman_sdm} that model performance is improved by including dependencies between latent variables.
+Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
+\begin{align}
+\label{eq:x_hat}
+\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
+\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1}), \\
+s_{t} &= align(\hdec_{t-1}, \hlangall), \\
+\canv_t &= \canv_{t-1} + \writeop(\hdec_t), 
+\label{eq:write}
+%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
+\end{align}
+where $\writeop$ and $\readop$ are the same attention operators as in \citep{gregor_draw}.
+The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}.
+Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities $\alpha_{1...N}$:
+%\begin{align}
+%\vspace{-0.1in}
+$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.$
+%\end{align}
+The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}.
+The final canvas $\canv_T$ is transformed using a sigmoid function to produce an Bernoulli data likelihood over pixels given the latent variables and the input caption.
+
+\comm{
 We extended the architecture of the DRAW network to include an additional input caption from the language model described in Sec.
@@
 
-where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
+where $\writeop$ is the same attention operators as in \citep{gregor_draw}.
+%where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
 % which apply a grid of Gaussian filters by specifying coordinates of the location and scale in order to yield a desirable image patch.
@@
 Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. 
+}
 
@@
 
-The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}.
-The model is trained to maximize a lower bound of the marginal likelihood $\loss$ of the correct image $x$ given the input caption $y$.
-The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$.
+%The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. 
+The model is trained to maximize a variational lower bound $\loss$ on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$, where $\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat_{1:T}) + \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars P(\Lat)\right)\right) \le \log P(\oimage\given\icaption)$ The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using a inference recurrent neural net.
+Similarly to the DRAW network, the inference RNN produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator over the training image at each step.
+\begin{align}
+\label{eq:x_hat}
+\hat{\oimage}_t &= \oimage-\sigmoid(\canv_{t-1})\\
+\label{eq:read}
+r_t &= \readop(\oimage_t, \hat{\oimage}_t, \hdec_{t-1})\\
+\henc_t &= \encoder(\henc_{t-1}, [r_t, \hdec_{t-1}])\\
+\post(\Lat_t|\oimage,\icaption,\Lat_{1:t-1}) &= \mathcal{N}\left(\mu_t(\henc_t), \sigma_t(\henc_t)\right)
+\end{align}
 
-The reconstruction loss $\rloss$ equals to $\frac{1}{L}\sum_{l=1}^{L}(log\,p(x_{t}|y,z)$ where $L$ is the number of samples used during training, which was set to $1$ in our experiments.
+%The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$.
 
-The latent loss $\lloss$ is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term.
-Since the patches drawn on canvas over time are not independent of each other, naturally the sufficient statistics of the prior distribution at time $t$ should be dependent on the sufficient statistics of the prior distribution at time $t-1$.
-Therefore, instead of setting $\prior(\Lat_1), ..., \prior(\Lat_T)$ to be independent unit gaussian distributions, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$, which forms a Markov chain $\prior(\Lat_1), \prior(\Lat_2\given\Lat_1), ..., \prior(\Lat_T\given\Lat_{T-1})$ as in \citep{bachman_sdm}, where
+%The reconstruction loss $\rloss$ equals to $\frac{1}{L}\sum_{l=1}^{L}(log\,p(x_{t}|y,z)$ where $L$ is the number of samples used during training, which was set to $1$ in our experiments.
+
+%The latent loss $\lloss$ is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term.
+Since the prior distribution at time $t$ is dependent on the sufficient statistics of the prior distribution at time $t-1$.
+Namely, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$.
 \begin{align}
@@
 \end{align}
-Overall, the loss function $\loss$ is calculated as follows:
+
+The terms in the variational lower bound are rearranged using the law of total expectation. Therefore, the loss function $\loss$ is calculated as follows:
 \begin{align}
@@
 To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set\cite{kiros_captions}.
+Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set \cite{kiros_captions}.
 We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.

### 2015-10-26 iclr-paper-new.tex
More changes

@@
 \usepackage{xcolor}
+\usepackage[bottom]{footmisc}
 
@@
 \newcommand{\hlangall}{\boldsymbol{h^{lang}}}
-\newcommand{\hdec}{h^{dec}}
-\newcommand{\henc}{h^{enc}}
+\newcommand{\hdec}{h^{gen}}
+\newcommand{\henc}{h^{infer}}
 \newcommand{\readop}{\mathit{read}}
 \newcommand{\writeop}{\mathit{write}}
-\newcommand{\encoder}{\mathit{LSTM}^{enc}}
-\newcommand{\decoder}{\mathit{LSTM}^{dec}}
+\newcommand{\encoder}{\mathit{LSTM}^{infer}}
+\newcommand{\decoder}{\mathit{LSTM}^{gen}}
 \newcommand{\canv}{c}
@@
 University of Toronto\\
-Toronto, Ontario, Canada \\
+Toronto, Ontario, Canada\\
 \texttt{\{emansim,eparisotto,rsalakhu\}@cs.toronto.edu}, \texttt{jimmy@psi.utoronto.ca}
@@
 \begin{abstract}
+%\vspace{-0.3cm}
 Motivated by the recent progress in generative models, we introduce a model that generates images from natural language descriptions.
@@
 \section{Introduction}
-
+%\vspace{-0.2cm}
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding.
@@
 %A significant amount of recent work has focused on generating captions from images
-\citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions}.
+\citep{karpathy_captions}, \citep{vinyals_captions}, \citep{xu_captions}.
 These models take an image descriptor and generate unstructured texts using a recurrent decoder.
@@
 The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \citep{denton_lapgan}.
-We then illustrate how our method, learnt on Microsoft COCO, generalizes to captions describing novel scenarios that are not seen in the dataset.
+We then illustrate how our method, learnt on Microsoft COCO \citep{mscoco}, generalizes to captions describing novel scenarios that are not seen in the dataset.
 
@@
 
-Deep Neural Networks have achieved significant success in various tasks such as image recognition \citep{krizhevsky_imagenet} and speech transcription \citep{graves_speech}. 
+Deep Neural Networks have achieved significant success in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech} and etc. 
 While most of the recent success has been achieved by discriminative models, generative models have not yet enjoyed the same level of success.
@@
 Our proposed model 
-defines a generative process of images conditioned on a caption.
+defines a generative process of images conditioned on captions.
 In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on a canvas $c_t$ over time $t=1,...,T$.
-Our model can be viewed as utilizing the sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}.
+Our model can be viewed as a part of the sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}.
 %Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
@@
 
-Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \cite{bachman_sdm}, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
+Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \citep{bachman_sdm}, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
 It has been shown in \citep{bachman_sdm} that model performance is improved by including dependencies between latent variables.
@@
 \end{align}
-where $\writeop$ and $\readop$ are the same attention operators as in \citep{gregor_draw}.
+where $\writeop$ is the same attention operator as in \citep{gregor_draw}.
 The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}.
@@
 
+\begin{figure}[!t]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\includegraphics[width=0.99\textwidth]{figures/new/mnist/test3.pdf}\quad
+%
+\end{center}
+\caption{Examples of generating $60 \times 60$ MNIST images corresponding to respective captions.
+The captions on the \textbf{left column} were part of the training set.
+The digits described in the captions on the \textbf{right column} were hidden during training for the respective configurations.}
+\label{fig:figmnist}
+\vspace{-0.3cm}
+\end{figure}
+
 During the image generation step, we discard the inference network and instead sample from the prior distribution. 
 Due to the blurriness of samples generated by the DRAW model, we perform an additional post processing step where we use a 
-adversarial network trained on residuals of a laplacian pyramid 
+adversarial network trained on residuals of a Laplacian pyramid conditioned on the skipthought representation \citep{kiros_skipthought} of the captions 
 to sharpen the generated images, similar to \citep{denton_lapgan}.
@@
 Two digits were either placed horizontally or vertically in nonoverlapping fashion.
-While many generative models are trained on the binarized version of the MNIST dataset, we trained directly on pixel intensities with a binary cross entropy cost function. 
-The generated images and the attention alignment are displayed in Figure~\ref{fig:figmnist}.
-
-\begin{figure}[!t]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\includegraphics[width=0.99\textwidth]{figures/new/mnist/test3.pdf}\quad
-%
-\end{center}
-\caption{Examples of generating $60 \times 60$ images corresponding to respective captions.
-Examples at the \textbf{left column} were present during training, whereas the \textbf{right column} examples were hidden during training.}
-\label{fig:figmnist}
-\vspace{-0.3cm}
-\end{figure}
+While many generative models were trained on the binarized version of the MNIST dataset, we trained directly on pixel intensities with a binary cross-entropy cost function. 
+The generated images with the attention alignments are displayed in Figure~\ref{fig:figmnist}.
 
@@
 very challenging.
-For consistency with related work on caption generation, we use only five captions when training our model.
-In the following subsections, we analyze both the qualitative and quantitative aspects of our model as well as compare its performance with that of other, related generative models.
+For consistency with related work on caption generation, we used only first five captions when training and evaluating our model.
+The images were resized to $32 \times 32$ dimension for consistency with other tiny images datasets \citep{krizhevsky_cifar}.
+In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
 
@@
 \captionsetup[subfigure]{labelformat=empty}
+\vspace{-0.3cm}
 \begin{center}
@@
 
-Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
-The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
-Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
-The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.
-
-\begin{figure}[!h]
+\begin{figure}[!b]
 \captionsetup[subfigure]{labelformat=empty}
+\vspace{-0.5cm}
 \begin{center}
@@
 \end{center}
-\caption{\textbf{Bottom}: Examples of changing the background while keeping the caption fixed. \textbf{Top}: The respective nearest training images based on pixel-wise L2 distance.}
+\caption{\textbf{Bottom}: Examples of changing the background while keeping the caption fixed.
+\textbf{Top}: The respective nearest training images based on pixel-wise L2 distance.
+The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.}
 \label{fig:genimages2}
@@
 \end{figure}
+
+Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
+The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
+Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
 
@@
 
-We also tried to analyze the way the model generated images.
-Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular timesteps.
-
 \begin{figure}[!h]
 \captionsetup[subfigure]{labelformat=empty}
+\vspace{-0.3cm}
 \begin{center}
@@
 
+We also tried to analyze the way the model generated images.
+Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular timesteps.
+
 \subsubsection{Comparison With Other Models}
@@
 
-\begin{figure}[!t]
+We compared the performance of the proposed model to the DRAW model conditioned on captions without align function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors \citep{kiros_skipthought} (skipthoughtDRAW).
+All of the conditional DRAW models were trained with binary cross-entropy cost function.
+We also compared our model with Fully-Connected (Fully-Conn) and Convolutional-Deconvolutional (Conv-Deconv) Variational Autoencoders which were trained with the least squares cost function.
+The LAPGAN model was trained on two level Laplacian Pyramid and was conditioned on the skipthought vectors.
+
+In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with least squares cost function have a watercolor effect on the images.
+We also found that the quality of generated samples was very similar among different variants of conditional DRAW models.
+
+\begin{figure}[!h]
 \captionsetup[subfigure]{labelformat=empty}
+\vspace{-0.3cm}
 \begin{center}
-\subfloat[Align DRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp.png}}\quad
+\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
 %
-\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan.png}}\quad
+\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
 %
-\subfloat[Conv. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae.png}}\quad
+\subfloat[Conv-Deconv VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
 %
-\subfloat[Fully-conn. VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae.png}}\quad
+\subfloat[Fully-Conn VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
 %
@@
 \label{fig:diffmodels}
-\vspace{-0.2cm}
+\vspace{-0.3cm}
 \end{figure}
-
-We compared the performance of the proposed model to the DRAW model conditioned on captions without align function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors (skipthoughtDRAW).
-All of the conditional DRAW models were trained with binary cross entropy cost function.
-We also compared our model with Fully-connected (FC) and Convolutional-Deconvolutional (Conv.) Variational Autoencoders which were trained with the least squares cost function.
-The LAPGAN model was trained on two level Laplacian Pyramid and was conditioned on the skipthought vectors.
-
-In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.
 
@@
 We rank the images in the test set conditioned on the captions based on the the variational lower bound of likelihood and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
-Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval.
-To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set \cite{kiros_captions}.
+Unsurprisingly, generative models didn't perform well in the image retrieval task.
+To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
+The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/nns.pkl}.} \citep{krizhevsky_cifar}.
+Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set \citep{kiros_captions}.
 We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.
@@
 
-Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation.
-Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image.
+Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI) \citep{wang_ssi}, which incorporates luminance and contrast masking into the error calculation.
+Strong inter-dependencies of closer pixels is also taken into account and the metric is calculated on small windows of the images.
 Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI.
@@
 
-\begin{table}[!b]
+\begin{table}[!t]
 \begin{center}
@@
 \hline
-\multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
+\multicolumn{7}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
 \hline
@@
 \hline
-Fully-conn. VAE (L2 cost) & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv. VAE (L2 cost) & 1.0 & 5.9 & 10.9 & 50.8 & 50 & \textbf{0.164} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-Skipthought DRAW & 2.6 & 12.6 & 21.3 & 66.9 & 33 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-Noalign DRAW & 3.2 & \textbf{16.3} & \textbf{26.0} & 72.5 & \textbf{27} & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-Align DRAW & \textbf{3.6} & 16.1 & \textbf{26.0} & \textbf{72.8} & 28 & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+Fully-Conn VAE & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv-Deconv VAE & 1.0 & 5.9 & 10.9 & 50.8 & 50 & 0.164 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+skipthoughtDRAW & 2.6 & 12.6 & 21.3 & 66.9 & 33 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 3.2 & 16.3 & 26.0 & 72.5 & 27 & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & 3.6 & 16.1 & 26.0 & 72.8 & 28 & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
 \end{tabulary}
 \end{center}
+\caption{Results of different models.}
+\label{tab:results}
+\vspace{-0.3cm}
 \end{table}
@@
 
-We have shown that our proposed model manages to generate reasonable images corresponding to respective captions as well as generalize to completely novel scenarios that never occur in real life.
-Sequential generation of images using attention played a crutial part in it.
-However the DRAW model still underfits on training data which leads to the blurry samples.
-We think that extending this process by generating very small sections of images like superpixels over potentially hundreds of timesteps might help the model fit the data and generate sharp images.
+We showed that our proposed model managed to generate reasonable images corresponding to respective captions as well as managed to generalize to completely novel scenarios which are very unlikely to occur in real life.
+Sequential generation of images using attention played a crucial part in it.
+However the conditional DRAW model still underfits on training data which leads to the blurry samples.
+We think that extending this process by generating very small sections of images over potentially hundreds of timesteps might help the model fit the data and generate sharper images.
 
 \section{Acknowledgements}
-We would like to thank developers of Theano for a powerful software and authors of \cite{denton_lapgan} for open sourcing their code.
+We would like to thank developers of Theano \citep{theano} for a powerful software and authors of \citep{denton_lapgan} for open sourcing their code.
 We would also like to thank Ryan Kiros and Nitish Srivastava for helpful discussions.
-We will release our code with pretrained models as soon as we clean it.
-Acknowledging funding (Russ)
+We will be releasing our code with pretrained models as soon as we clean it.
 

### 2015-10-26 iclr-paper-new.tex
More changes made

@@
 \lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
-\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1}), \\
+\hdec_t &= \decoder(\hdec_{t-1}, [z_t, s_{t-1}]), \\
 s_{t} &= align(\hdec_{t-1}, \hlangall), \\
@@
 where $\writeop$ is the same attention operator as in \citep{gregor_draw}.
+The $\writeop$ operator applies an array of 2D Gaussian filters on the patch that places it on the canvas with smoothly varying location and zoom.
 The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}.
@@
 %\end{align}
-The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, as in \citep{bahdanau_mt}.
+The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies $e_{1...N}$ using $tanh$ perceptron and then normalizing them, as in \citep{bahdanau_mt}:
+%\begin{align}
+$\alpha_{k}=\frac{exp(v^{T}tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}exp(v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)}.$
+%\end{align}
 The final canvas $\canv_T$ is transformed using a sigmoid function to produce an Bernoulli data likelihood over pixels given the latent variables and the input caption.
@@
 %The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. 
-The model is trained to maximize a variational lower bound $\loss$ on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$, where $\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat_{1:T}) + \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars P(\Lat)\right)\right) \le \log P(\oimage\given\icaption)$ The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using a inference recurrent neural net.
-Similarly to the DRAW network, the inference RNN produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator over the training image at each step.
+The model is trained to maximize a variational lower bound $\loss$ on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$, where $\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat_{1:T}) + \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars P(\Lat)\right)\right) \le \log P(\oimage\given\icaption)$.
+The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using a inference recurrent neural net.
+Similarly to the DRAW network, the inference RNN produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using an array of Gaussian filters (inverse of $\writeop$), over the training image at each step.
 \begin{align}
@@
 
+\newpage
 \section{Experiments}
@@
 While many generative models were trained on the binarized version of the MNIST dataset, we trained directly on pixel intensities with a binary cross-entropy cost function.
-The generated images with the attention alignments are displayed in Figure~\ref{fig:figmnist}.
+
+The generated images together with the attention alignments are displayed in Figure~\ref{fig:figmnist}.
+The model correctly displayed defined digits at the described positions and even managed to generalize reasonably to the configurations that were never present during training.
+In case of generating two digits, the model attended to the digit in the caption that was drawn at the respective timestep while completely ignoring the other digit.
+As for generating one digit, the model attended to the digit in the caption during the whole generation process.
+However in both cases, the model placed small attention values on the words describing the position of digits in the images.
 
@@
 
-\begin{figure}[!b]
+\begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \begin{center}
@@
 Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI.
-The results are reported in Table~\ref{tab:results}.
+The results are reported on Table~\ref{tab:results}.
 
@@
 
-We showed that our proposed model managed to generate reasonable images corresponding to respective captions as well as managed to generalize to completely novel scenarios which are very unlikely to occur in real life.
-Sequential generation of images using attention played a crucial part in it.
-However the conditional DRAW model still underfits on training data which leads to the blurry samples.
-We think that extending this process by generating very small sections of images over potentially hundreds of timesteps might help the model fit the data and generate sharper images.
+We showed that our proposed model managed to generate reasonably looking images corresponding to respective captions as well as managed to generalize to completely novel scenarios which are very unlikely to occur in real life.
+Sequential generation of images over some number of steps has played an important part in this process.
+However, the tendency of the conditional DRAW models to underfit on the training data has resulted in the blurry samples.
+As a result, the samples were sharpened by the GAN trained on the residuals of the images.
+Unfortunately, this is not a quite satisfying solution due to the fact that the whole model wasn't trained in the end-to-end fashion.
+We think that further extention of the capabilities of the sequential generation of images might help the variational models better fit the data and escape the curse of blurriness.
 
@@
 
+\newpage
 \bibliography{iclr-paper}

### 2015-10-26 supp.tex
fixed small error in supplement
More changes made

@@
 Each parameter in alignDRAW was initialized by sampling from the Gaussian distribution of mean $0$ and standard deviation of $0.01$.
-All models were trained with \textit{RMSProp} with initial learning rate of $0.001$.
-The norm of the gradients was rescaled to $10.0$ during training.
-For all tasks we used $128$ units in language model and $512$ units in alignment multi-layer perceptron.
-
-% \begin{table}[!h]
-% \begin{center}
-% \begin{tabulary}{\linewidth}{c | c c c}
-% \hline
-% \multicolumn{4}{c}{\textbf{Language Model}} \\
-% \hline
-% Task & Vocab size & LSTM \#$h$ & Align \#$h$\\
-% \hline
-% MNIST With Captions & 22 & 128 & 512  \\
-% Microsoft COCO & 25323 & 128 & 512
-% \end{tabulary}
-% \end{center}
-% \end{table}
+The model was trained using \textit{RMSProp} with initial learning rate of $0.001$.
+The norm of the gradients was clipped to $10.0$ during training to deal with exploding gradients.
+For all tasks we used $128$ units in the language model and $512$ units in the $tanh$ perceptron in the alignment function.
+For the Microsoft COCO task, the alignDRAW model was trained for $18$ epochs (the model converges after $14$ epochs) and the learning rate was reduced by the factor of $0.1$ after $11$ epochs.
+The architectural configurations of alignDRAW models are shown on Table~\ref{tab:drawhyper}.
 
@@
 \hline
-\multicolumn{7}{c}{\textbf{Image Model}} \\
+\multicolumn{7}{c}{\textbf{alignDRAW Model}} \\
 \hline
@@
 \hline
-MNIST Captions & 32 & 550 & 550 & 275 & 9 & 9\\
-MS COCO & 32 & 300 & 300 & 150 & 8 & 8\\
+Microsoft COCO & 32 & 550 & 550 & 275 & 9 & 9\\
+MNIST Captions & 32 & 300 & 300 & 150 & 8 & 8\\
 \end{tabulary}
 \end{center}
+\label{tab:drawhyper}
+\caption{The architectural configurations of alignDRAW models.}
+%\vspace{-0.3cm}
 \end{table}
 
+The GAN model trained on residuals which was used for sharpening had the same configuration as the $28 \times 28$ model trained by \cite{denton_lapgan} on residuals of CIFAR dataset, that can be found at \url{https://gist.github.com/soumith/e3f722173ea16c1ea0d9}.
+The model was trained for $6$ epochs.
+
 \subsection{Evaluation}
+
+On Table~\ref{tab:nll}, you can see that the alignDRAW model didn't overfit on the training data as well as managed to achieve a better lower bound of log likelihood on the test data.
+Also notice that, the results worsened dramatically after sharpening test images.
+
 \begin{table}[!h]
@@
 \hline
-Model & Train NLL & Valid NLL & Test NLL & Test NLL (after sharpening)\\
+Model & Train LL & Valid LL & Test LL & Test LL (after sharpening)\\
 \hline
-skipthoughtDRAW & 1794.29 & 1797.41 & 1791.37 & 2045.84 \\
-noalignDRAW & 1792.14 & 1796.94 & 1791.15 & 2051.07 \\
-alignDRAW & 1792.15 & 1797.24 & 1791.53 & 2042.31
+skipthoughtDRAW & -1794.29 & -1797.41 & -1791.37 & -2045.84 \\
+noalignDRAW & -1792.14 & -1796.94 & -1791.15 & -2051.07 \\
+alignDRAW & -1792.15 & -1797.24 & -1791.53 & -2042.31
 \end{tabulary}
 \end{center}
+\label{tab:nll}
+\caption{The lower bound of log likelihood of conditional DRAW models on the Microsoft COCO.}
 \end{table}

### 2015-10-27 iclr-paper-new.tex
Small changes here and there

@@
 \hline
-Fully-Conn VAE & 1.0 & 5.6 & 10.4 & 51.1 & 51 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv-Deconv VAE & 1.0 & 5.9 & 10.9 & 50.8 & 50 & 0.164 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-skipthoughtDRAW & 2.6 & 12.6 & 21.3 & 66.9 & 33 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-noalignDRAW & 3.2 & 16.3 & 26.0 & 72.5 & 27 & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-alignDRAW & 3.6 & 16.1 & 26.0 & 72.8 & 28 & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
 \end{tabulary}
@@
 We showed that our proposed model managed to generate reasonably looking images corresponding to respective captions as well as managed to generalize to completely novel scenarios which are very unlikely to occur in real life.
-Sequential generation of images over some number of steps has played an important part in this process.
+Sequential generation of images over the number of steps using attention mechanism has played an important part in this process.
 However, the tendency of the conditional DRAW models to underfit on the training data has resulted in the blurry samples.

### 2015-10-27 supp.tex
Small changes here and there

@@
 For all tasks we used $128$ units in the language model and $512$ units in the $tanh$ perceptron in the alignment function.
-For the Microsoft COCO task, the alignDRAW model was trained for $18$ epochs (the model converges after $14$ epochs) and the learning rate was reduced by the factor of $0.1$ after $11$ epochs.
+For the Microsoft COCO task, the alignDRAW model was trained for $18$ epochs and the learning rate was reduced by the factor of $0.1$ after $11$ epochs.
 The architectural configurations of alignDRAW models are shown on Table~\ref{tab:drawhyper}.
@@
 \end{tabulary}
+\caption{The architectural configurations of alignDRAW models.}
+\label{tab:drawhyper}
 \end{center}
-\label{tab:drawhyper}
-\caption{The architectural configurations of alignDRAW models.}
 %\vspace{-0.3cm}
@@
 \end{tabulary}
+\caption{The lower bound of log likelihood of conditional DRAW models on the Microsoft COCO.}
+\label{tab:nll}
 \end{center}
-\label{tab:nll}
-\caption{The lower bound of log likelihood of conditional DRAW models on the Microsoft COCO.}
 \end{table}

### 2015-10-27 iclr-paper-new.tex
spelling

@@
 After training on Microsoft COCO, we compare our models with several baseline generative models on image generation and retrieval tasks.
-We demonstrate our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.
+We demonstrate that our model produces higher quality samples than other approaches and generates images with novel scene compositions corresponding to previously unseen captions in the dataset.
 \end{abstract}
@@
 The challenging nature of this task has motivated recent approaches to exploit the inference and generative capabilities of deep neural networks.
-Previously studied generative models of images often defined distributions that were restricted to being either unconditioned or conditioned on classification labels.
+Previously studied deep generative models of images often defined distributions that were restricted to being either unconditioned or conditioned on classification labels.
 In real world applications, however, images rarely appear in isolation as they are often accompanied by unstructured textual descriptions, such as on web pages and in books.
@@
 
-Deep Neural Networks have achieved significant success in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech} and etc. 
+Deep Neural Networks have achieved significant success in various tasks such as image recognition \citep{krizhevsky_imagenet} and speech transcription \citep{graves_speech}. 
 While most of the recent success has been achieved by discriminative models, generative models have not yet enjoyed the same level of success.
@@
 To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}.
-The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ and accumulates the output at each iterative step..
+The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ where the output is accumulated over all time steps.
 
@@
 %\end{align}
-The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies $e_{1...N}$ using $tanh$ perceptron and then normalizing them, as in \citep{bahdanau_mt}:
+The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies $e_{1...N}$ using a $tanh$ perceptron and then normalizing them, as in \citep{bahdanau_mt}:
 %\begin{align}
@@
 %\end{align}
-The final canvas $\canv_T$ is transformed using a sigmoid function to produce an Bernoulli data likelihood over pixels given the latent variables and the input caption.
+The final canvas $\canv_T$ is transformed using a sigmoid function to produce a conditional Bernoulli data likelihood over pixels given the latent variables and the input caption.
 
@@
 \end{align}
-The expectation can be approximate by $\numSamples$ Monte Carlo samples $\LatSample_{1:T}$ from $\post(\Lat_{1:T}\given\icaption, \oimage)$:
+The expectation can be approximated by $\numSamples$ Monte Carlo samples $\LatSample_{1:T}$ from $\post(\Lat_{1:T}\given\icaption, \oimage)$:
 \begin{align}
@@
 The reconstuction loss becomes the loss between sharpened image and correct image, whereas the latent loss stays the same.
-Additionally, we noticed that sampling from the mean of uniform distribution generated samples with less noise than if we had otherwise sampled from the uniform distribution itself.
+Additionally, we noticed that sampling from the mean of the uniform distribution generated samples with less noise than if we had otherwise sampled from the uniform distribution itself.
 
@@
 The generated images together with the attention alignments are displayed in Figure~\ref{fig:figmnist}.
-The model correctly displayed defined digits at the described positions and even managed to generalize reasonably to the configurations that were never present during training.
-In case of generating two digits, the model attended to the digit in the caption that was drawn at the respective timestep while completely ignoring the other digit.
-As for generating one digit, the model attended to the digit in the caption during the whole generation process.
+The model correctly displayed the specified digits at the described positions and even managed to generalize reasonably to the configurations that were never present during training.
+In the case of generating two digits, the model attended to the digit in the caption it was drawing at that particular timestep while completely ignoring the other digit. 
+Similarly, when generating one digit the model attended to the digit in the caption during the whole generation process.
 However in both cases, the model placed small attention values on the words describing the position of digits in the images.
@@
 %conditioned on a caption 
-very challenging.
-For consistency with related work on caption generation, we used only first five captions when training and evaluating our model.
-The images were resized to $32 \times 32$ dimension for consistency with other tiny images datasets \citep{krizhevsky_cifar}.
+very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
+The images were resized to $32 \times 32$ dimension for consistency with other tiny image datasets \citep{krizhevsky_cifar}.
 In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
@@
 As a test of model generalization, we tried generating images corresponding to captions that describe scenarios that are highly unlikely to occur in real life.
-These captions describe some common object doing unusual things
-or in a strange location.
+These captions describe a common object doing unusual things
+or set in a strange location.
 %that are impossible in the real world. 
 %considering the physics of real world. 
-Even though these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
+Even though some of these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
 Nevertheless, as you can see in Figure~\ref{fig:genimages4}, the model managed to generate reasonable images.
@@
 After flipping sets of words in the captions, we were curious to see which words the model attended to when generating images.
-It turned out that during the generation step, the model mostly focused on the specific words (or their neighbors) that carried the main semantic meaning expressed in the sentences.
-The attention values in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
+It turned out that during the generation step, the model mostly focused on the specific words (or nearby words) that carried the main semantic meaning expressed in the sentences.
+The attention values of words in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
 For example, in Figure~\ref{fig:attentimages1} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
@@
 We also compared our model with Fully-Connected (Fully-Conn) and Convolutional-Deconvolutional (Conv-Deconv) Variational Autoencoders which were trained with the least squares cost function.
-The LAPGAN model was trained on two level Laplacian Pyramid and was conditioned on the skipthought vectors.
+The LAPGAN model was trained on a two level Laplacian Pyramid with a GAN as a top layer generator and all stages were conditioned on the same skipthought vector.
 
@@
 As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods.
-We rank the images in the test set conditioned on the captions based on the the variational lower bound of likelihood and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
+We rank the images in the test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
 Unsurprisingly, generative models didn't perform well in the image retrieval task.
@@
 The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/nns.pkl}.} \citep{krizhevsky_cifar}.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set \citep{kiros_captions}.
+Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set \citep{kiros_captions}.
 We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.
@@
 Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI) \citep{wang_ssi}, which incorporates luminance and contrast masking into the error calculation.
-Strong inter-dependencies of closer pixels is also taken into account and the metric is calculated on small windows of the images.
+Strong inter-dependencies of closer pixels are also taken into account and the metric is calculated on small windows of the images.
 Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI.
@@
 We showed that our proposed model managed to generate reasonably looking images corresponding to respective captions as well as managed to generalize to completely novel scenarios which are very unlikely to occur in real life.
-Sequential generation of images over the number of steps using attention mechanism has played an important part in this process.
-However, the tendency of the conditional DRAW models to underfit on the training data has resulted in the blurry samples.
-As a result, the samples were sharpened by the GAN trained on the residuals of the images.
-Unfortunately, this is not a quite satisfying solution due to the fact that the whole model wasn't trained in the end-to-end fashion.
-We think that further extention of the capabilities of the sequential generation of images might help the variational models better fit the data and escape the curse of blurriness.
+Sequential generation of images over the number of steps using an attention mechanism has played an important part in this process.
+However, the tendency of the conditional DRAW models to underfit on the training data has resulted in blurry samples.
+As a result, the samples were sharpened by a LAPGAN trained on the residuals of the images.
+Unfortunately, this is not a quite satisfying solution due to the fact that the whole model wasn't trained in an end-to-end fashion.
+We think that a further extension of the capabilities of the sequential generation of images might help the variational models better fit the data and escape the curse of blurriness.
 

### 2015-10-27 supp.tex
spelling

@@
 
-The GAN model trained on residuals which was used for sharpening had the same configuration as the $28 \times 28$ model trained by \cite{denton_lapgan} on residuals of CIFAR dataset, that can be found at \url{https://gist.github.com/soumith/e3f722173ea16c1ea0d9}.
+The LAPGAN model used for sharpening had the same configuration as the $28 \times 28$ model trained by \cite{denton_lapgan} on the edge residuals of CIFAR dataset.
+that can be found at \url{https://gist.github.com/soumith/e3f722173ea16c1ea0d9}.
 The model was trained for $6$ epochs.

### 2015-10-27 iclr-paper-new.tex
change url

@@
 To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
-The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/nns.pkl}.} \citep{krizhevsky_cifar}.
+The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.
 Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set \citep{kiros_captions}.

### 2015-10-29 iclr-paper-new.tex
Added new results

@@
 Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI.
-The results are reported on Table~\ref{tab:results}.
+As you can see on Table~\ref{tab:results}, SSI scores achieved by variational models were higher compared to SSI score achieved by LAPGAN.
+As for another metric of image similarity, we calculated euclidian distance between generated images and real images in the feature space of VGG-like model trained on CIFAR dataset.
+This is not necessarily a good metric, since the statistics of generated images were vastly different from statistics of training images.
+Nevertheless, the results are reported on Table~\ref{tab:results}.
 
@@
 \begin{center}
-\begin{tabulary}{\linewidth}{c || c c c c c || c}
+\begin{tabulary}{\linewidth}{c || c c c c c || c c}
 \hline
-\multicolumn{7}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
+\multicolumn{8}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
 \hline
-& \multicolumn{5}{c||}{Image Search} & Image Similarity \\
+& \multicolumn{5}{c||}{Image Search} & \multicolumn{2}{c}{Image Similarity} \\
 %\cline{2-6}
-\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
+\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} & \textbf{Feature Distance} \\
 \hline
 \hline
-LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
+LAPGAN & - & - & - & - & - & 0.08 & 6.89 \\ % - & - & - & - & - & 0.08
 \hline
-Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 & 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 & 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 & 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 & 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 & 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
 \end{tabulary}
@@
 
-\newpage
 \bibliography{iclr-paper}

### 2015-10-29 iclr-paper-new.tex
discussion

@@
 
-We showed that our proposed model managed to generate reasonably looking images corresponding to respective captions as well as managed to generalize to completely novel scenarios which are very unlikely to occur in real life.
-Sequential generation of images over the number of steps using an attention mechanism has played an important part in this process.
-However, the tendency of the conditional DRAW models to underfit on the training data has resulted in blurry samples.
-As a result, the samples were sharpened by a LAPGAN trained on the residuals of the images.
-Unfortunately, this is not a quite satisfying solution due to the fact that the whole model wasn't trained in an end-to-end fashion.
-We think that a further extension of the capabilities of the sequential generation of images might help the variational models better fit the data and escape the curse of blurriness.
+In this paper, we have demonstrated that our alignDraw model, a combination of a sequential variational autoencoder and a alignment model over words, succeeds in generating images that correspond to a given input caption.
+By making extensive use of attentional mechanisms, our model gains several advantages.
+Namely, the use of the visual attention mechanism allows us to decompose the problem of image generation into a series of steps instead of a single forward pass, while the attention over words can provide us insight whenever our model fails to generate a relevant image.
+Additionally, because our model conditions on unstructured textual descriptions, we can generate images corresponding to captions that generalize beyond the training set such as sentences describing completely novel scenarios which are highly unlikely to occur in real life.
+Because we have observed that the alignDraw model tends to output slightly blurry samples, likely due to underfitting on the training data, we augmented the model with a sharpening post-processing step in which a LAPGAN generates edges that are added to the alignDraw samples.
+Unfortunately, this is not an ideal solution due to the fact that the whole model is not trained in an end-to-end fashion.
+Therefore a direction of future work would be to find methods that can the bypass the separate post-processing step and output sharp images directly in an end-to-end manner.
 

### 2015-10-30 iclr-paper-new.tex
Fixed Math

--- 2015-10-27 iclr-paper-new.tex
@@
 To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}.
-The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ where the output is accumulated over all time steps.
+The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ where the output is accumulated over all timesteps.
 
@@
 It has been shown in \citep{bachman_sdm} that model performance is improved by including dependencies between latent variables.
+Namely, the mean and variance of $\prior(\Lat_t)$ is:
+\begin{align}
+\mu_{t}^{prior} &= tanh(W_{\mu}\hdec_{t-1})\\
+\sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
+\end{align}
 Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
@@
 \lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
-\hdec_t &= \decoder(\hdec_{t-1}, [z_t, s_{t-1}]), \\
 s_{t} &= align(\hdec_{t-1}, \hlangall), \\
+\hdec_t &= \decoder(\hdec_{t-1}, [z_t, s_{t}]), \\
 \canv_t &= \canv_{t-1} + \writeop(\hdec_t), 
@@
 \end{align}
-where $\writeop$ is the same attention operator as in \citep{gregor_draw}.
-The $\writeop$ operator applies an array of 2D Gaussian filters on the patch that places it on the canvas with smoothly varying location and zoom.
+where $\mathit{LSTM}$ is a standard \textit{Long Short-Term Memory} architecture \citep{hochreiter_lstm} with \textit{forget} gates \citep{gers_forget}, $[v_1, v_2]$ is the concatenation of the vectors $v_1$ and $v_2$ into the single vector, and $\canv_0$, $\hdec_0$ are initialized to the the learned biases.
+The $\writeop$ operator applies an array of Gaussian filters on the patch that places it on the canvas with smoothly varying location and zoom (same as in \citep{gregor_draw}).
 The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}.
-Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities $\alpha_{1...N}$:
+Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities $\alpha_{1...N}^{t}$:
 %\begin{align}
 %\vspace{-0.1in}
-$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.$
+$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.$
 %\end{align}
-The corresponding alignment probabilities $\alpha_{1...N}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies $e_{1...N}$ using a $tanh$ perceptron and then normalizing them, as in \citep{bahdanau_mt}:
+The corresponding alignment probabilities $\alpha_{1...N}^{t}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies using a $tanh$ perceptron scaled by vector $v$ and then normalizing them, as in \citep{bahdanau_mt}:
 %\begin{align}
-$\alpha_{k}=\frac{exp(v^{T}tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}exp(v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)}.$
+%\vspace{-0.1in}
+$\alpha_{k}^{t}=\frac{exp(v^{T}tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}exp(v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)}.$
 %\end{align}
@@
 The model is trained to maximize a variational lower bound $\loss$ on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$, where $\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat_{1:T}) + \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars P(\Lat)\right)\right) \le \log P(\oimage\given\icaption)$.
-The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using a inference recurrent neural net.
+%The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using the inference RNN.
 Similarly to the DRAW network, the inference RNN produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using an array of Gaussian filters (inverse of $\writeop$), over the training image at each step.
@@
 %The latent loss $\lloss$ is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term.
-Since the prior distribution at time $t$ is dependent on the sufficient statistics of the prior distribution at time $t-1$.
+%Since the prior distribution at time $t$ is dependent on the sufficient statistics of the prior distribution at time $t-1$.
 Namely, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$.
-\begin{align}
-\mu_{t}^{prior} &= tanh(W_{\mu}\hdec_{t-1})\\
-\sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
-\end{align}
+%\begin{align}
+%\mu_{t}^{prior} &= tanh(W_{\mu}\hdec_{t-1})\\
+%\sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
+%\end{align}
+
+$\sigmoid$ operator is a logistic sigmoid function $\sigmoid=\frac{1}{1+exp(-x)}$ and $\henc_0$ is initialized to the learned bias.
 
@@
 \begin{align}
-\loss =  &\expectation_{Q(\Lat_{1:t}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
+\loss =  &\expectation_{Q(\Lat_{1:T}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
 \end{align}
@@
 \end{align}
+We used only one sample during training for all our experiments.
 \comm{
@@
 Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI.
-The results are reported on Table~\ref{tab:results}.
+As you can see on Table~\ref{tab:results}, SSI scores achieved by variational models were higher compared to SSI score achieved by LAPGAN.
+%As for another metric of image similarity, we calculated euclidian distance between generated images and real images in the feature space of VGG-like model trained on CIFAR dataset.
+This is not necessarily a good metric, since the statistics of generated images were vastly different from statistics of training images.
+Nevertheless, the results are reported on Table~\ref{tab:results}.
 
@@
 \begin{center}
-\begin{tabulary}{\linewidth}{c || c c c c c || c}
+\begin{tabulary}{\linewidth}{c || c c c c c || c c}
 \hline
@@
 \hline
-& \multicolumn{5}{c||}{Image Search} & Image Similarity \\
+& \multicolumn{5}{c||}{Image Search} & \multicolumn{1}{c}{Image Similarity} \\
 %\cline{2-6}
-\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
+\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
 \hline
 \hline
-LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
+LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
 \hline
-Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
 \end{tabulary}
@@
 
-We showed that our proposed model managed to generate reasonably looking images corresponding to respective captions as well as managed to generalize to completely novel scenarios which are very unlikely to occur in real life.
-Sequential generation of images over the number of steps using an attention mechanism has played an important part in this process.
-However, the tendency of the conditional DRAW models to underfit on the training data has resulted in blurry samples.
-As a result, the samples were sharpened by a LAPGAN trained on the residuals of the images.
-Unfortunately, this is not a quite satisfying solution due to the fact that the whole model wasn't trained in an end-to-end fashion.
-We think that a further extension of the capabilities of the sequential generation of images might help the variational models better fit the data and escape the curse of blurriness.
+In this paper, we demonstrated that the alignDRAW model, a combination of a recurrent variational autoencoder with an alignment model over words, succeeded in generating images that correspond to a given input caption.
+By extensively using attentional mechanisms, our model gained several advantages.
+Namely, the use of the visual attention mechanism allowed us to decompose the problem of image generation into a series of steps instead of a single forward pass, while the attention over words provided us an insight whenever our model failed to generate a relevant image.
+Additionally, our model generated images corresponding to captions which generalized beyond the training set, such as sentences describing novel scenarios which are highly unlikely to occur in real life.
 
+Because the alignDRAW model tends to output slightly blurry samples, likely due to underfitting on the training data, we augmented the model with a sharpening post-processing step in which GAN generated edges which were added to the alignDRAW samples.
+Unfortunately, this is not an ideal solution due to the fact that the whole model was not trained in an end-to-end fashion.
+Therefore a direction of future work would be to find methods that can bypass the separate post-processing step and output sharp images directly in an end-to-end manner.
+
+\newpage
 \section{Acknowledgements}

### 2015-10-30 iclr-paper-new.tex
Small change
Small changes again

@@
 \begin{align}
-\mu_{t}^{prior} &= tanh(W_{\mu}\hdec_{t-1})\\
-\sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
+\mu_{t}(\hdec_{t-1}) &= tanh(W_{\mu}\hdec_{t-1})\\
+\sigma_{t}(\hdec_{t-1}) &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
 \end{align}
@@
 
-$\sigmoid$ operator is a logistic sigmoid function $\sigmoid=\frac{1}{1+exp(-x)}$ and $\henc_0$ is initialized to the learned bias.
+where $\hat{\oimage}$ is the \textit{error} image, $\sigmoid$ operator is a logistic sigmoid function $\sigmoid=\frac{1}{1+exp(-x)}$ and $\henc_0$ is initialized to the learned bias.
 
@@
 \begin{align}
-\loss =  &\expectation_{Q(\Lat_{1:T}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
+\loss =  &\expectation_{Q(\Lat_{1:T}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
 \end{align}
@@
 \begin{align}
-\loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ - \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\icaption,\oimage)\klBars\prior(\Lat_1\given\icaption)\right).
+\loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ - \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
 \end{align}

### 2015-10-30 iclr-paper-new.tex
ugly version of the paper

@@
 The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \citep{denton_lapgan}.
-We then illustrate how our method, learnt on Microsoft COCO \citep{mscoco}, generalizes to captions describing novel scenarios that are not seen in the dataset.
+We then illustrate how our method, learnt on Microsoft COCO \citep{mscoco}, generalizes to captions describing novel scenarios that are not seen in the dataset (Figure~\ref{fig:genimages4}).
 
 \section{Related Work}
+
+\begin{figure}[!t]
+\captionsetup[subfigure]{labelformat=empty}
+\vspace{-0.3cm}
+\begin{center}
+\subfloat[A stop sign is flying in blue skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
+%
+\subfloat[A herd of elephants flying in the blue skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
+%
+\subfloat[A toilet seat sits open in the grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
+%
+\subfloat[A person skiing on sand clad vast desert.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\quad
+%
+\end{center}
+\caption{Examples of novel scene compositions.}
+\label{fig:genimages4}
+\vspace{-0.3cm}
+\end{figure}
 
@@
 
-\newpage
+%\newpage
 \section{Experiments}
@@
 \subsubsection{Analysis of Generated Images}
+
 The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
@@
 
-\begin{figure}[!h]
+\begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
+\vspace{-0.1cm}
 \begin{center}
@@
 \label{fig:genimages1}
-\vspace{-0.3cm}
+\vspace{-0.1cm}
 \end{figure}
@@
 \captionsetup[subfigure]{labelformat=empty}
-%\vspace{-0.5cm}
+\vspace{-0.1cm}
 \begin{center}
@@
 \label{fig:genimages2}
-\vspace{-0.3cm}
+\vspace{-0.1cm}
 \end{figure}
 
-Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
-The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
-Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
-
-Despite a large number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
-However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
-In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
-This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
-
-\begin{figure}[!h]
+\begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
+\vspace{-0.1cm}
 \begin{center}
@@
 \label{fig:genimages3}
-\vspace{-0.3cm}
+\vspace{-0.1cm}
 \end{figure}
 
-As a test of model generalization, we tried generating images corresponding to captions that describe scenarios that are highly unlikely to occur in real life.
-These captions describe a common object doing unusual things
-or set in a strange location.
-%that are impossible in the real world. 
-%considering the physics of real world. 
-Even though some of these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
-Nevertheless, as you can see in Figure~\ref{fig:genimages4}, the model managed to generate reasonable images.
-
-\begin{figure}[!h]
+\begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
-\begin{center}
-\subfloat[A stop sign is flying in blue skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
-%
-\subfloat[A herd of elephants flying in the blue skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
-%
-\subfloat[A toilet seat sits open in the grass field.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
-%
-\subfloat[A person skiing on sand clad vast desert.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\quad
-%
-\end{center}
-\caption{Examples of novel scene compositions.}
-\label{fig:genimages4}
-\vspace{-0.3cm}
-\end{figure}
-
-\subsubsection{Analysis of Attention}
-
-After flipping sets of words in the captions, we were curious to see which words the model attended to when generating images.
-It turned out that during the generation step, the model mostly focused on the specific words (or nearby words) that carried the main semantic meaning expressed in the sentences.
-The attention values of words in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
-For example, in Figure~\ref{fig:attentimages1} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
-This suggests that, in their respective sentences, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
-In contrast, when we swap words ``beach'' and ``sun'', we can see a drastic change between sentences in the probability distribution over words.
-By noting that the model completely ignores the word ``sun'' in the second sentence, we can therefore gain a more thorough understanding of why we see no visual differences between the images generated by each caption.
-
-\begin{figure}[!h]
-\captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
+\vspace{-0.1cm}
 \begin{center}
@@
 \label{fig:attentimages1}
-\vspace{-0.3cm}
+\vspace{-0.1cm}
 \end{figure}
+
+\begin{figure}[!h]
+\captionsetup[subfigure]{labelformat=empty}
+\vspace{-0.1cm}
+\begin{center}
+\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
+%
+\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
+%
+\subfloat[Conv-Deconv VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
+%
+\subfloat[Fully-Conn VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
+%
+\end{center}
+\caption{Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
+\label{fig:diffmodels}
+\vspace{-0.1cm}
+\end{figure}
+
+Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
+The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
+Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
+
+Despite a large number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
+However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
+In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
+This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
+
+As a test of model generalization, we tried generating images corresponding to captions that describe scenarios that are highly unlikely to occur in real life.
+These captions describe a common object doing unusual things
+or set in a strange location.
+%that are impossible in the real world. 
+%considering the physics of real world. 
+Even though some of these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
+Nevertheless, as you can see in Figure~\ref{fig:genimages4}, the model managed to generate reasonable images.
+
+\subsubsection{Analysis of Attention}
+
+After flipping sets of words in the captions, we were curious to see which words the model attended to when generating images.
+It turned out that during the generation step, the model mostly focused on the specific words (or nearby words) that carried the main semantic meaning expressed in the sentences.
+The attention values of words in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
+For example, in Figure~\ref{fig:attentimages1} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
+This suggests that, in their respective sentences, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
+In contrast, when we swap words ``beach'' and ``sun'', we can see a drastic change between sentences in the probability distribution over words.
+By noting that the model completely ignores the word ``sun'' in the second sentence, we can therefore gain a more thorough understanding of why we see no visual differences between the images generated by each caption.
 
@@
 We also found that the quality of generated samples was very similar among different variants of conditional DRAW models.
-
-\begin{figure}[!h]
-\captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
-\begin{center}
-\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
-%
-\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
-%
-\subfloat[Conv-Deconv VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
-%
-\subfloat[Fully-Conn VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
-%
-\end{center}
-\caption{Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
-\label{fig:diffmodels}
-\vspace{-0.3cm}
-\end{figure}
 
@@
 
-\newpage
+%\newpage
 \section{Acknowledgements}

### 2015-10-30 iclr-paper-new.tex
less ugly version of paper

--- 2015-10-30 iclr-paper-new.tex
@@
 
-Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
-The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model
-%conditioned on a caption 
-very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
-The images were resized to $32 \times 32$ dimension for consistency with other tiny image datasets \citep{krizhevsky_cifar}.
-In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
-
-\subsubsection{Analysis of Generated Images}
-The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
-and then use that knowledge to generate relevant images.
-To examine the understanding of our model, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
-
-First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
-In Figure~\ref{fig:genimages1}, we generated images of school buses with four different colors: yellow, red, green and blue.
-Although, there are images of buses with different colors in the training set, all mentioned school buses are specifically colored yellow.
-Despite that, the model managed to generate images of an object that is visually reminiscent of a school bus that is painted with the specified color.
-
-\begin{figure}[!h]
+\begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
@@
 ]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+\\
+%
+\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
+]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
+%
+\subfloat[\underline{A bowl of bananas} is on the table.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
+%
+\subfloat[A vintage photo of a \underline{cat}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
+%
+\subfloat[A vintage photo of a \underline{dog}.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
 %
 \end{center}
-\caption{Examples of changing the color while keeping the caption fixed.}
-\label{fig:genimages1}
+\caption{\textbf{Bottom}: Examples of changing the object while keeping the caption fixed. \textbf{Top}: Examples of changing the color while keeping the caption fixed.}
+\label{fig:genimages3}
 \vspace{-0.3cm}
 \end{figure}
+
+Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
+The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model
+%conditioned on a caption 
+very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
+The images were resized to $32 \times 32$ dimension for consistency with other tiny image datasets \citep{krizhevsky_cifar}.
+In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
+
+\subsubsection{Analysis of Generated Images}
+The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
+and then use that knowledge to generate relevant images.
+To examine the understanding of our model, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
+
+First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
+In Figure~\ref{fig:genimages3}, we generated images of school buses with four different colors: yellow, red, green and blue.
+Although, there are images of buses with different colors in the training set, all mentioned school buses are specifically colored yellow.
+Despite that, the model managed to generate images of an object that is visually reminiscent of a school bus that is painted with the specified color.
 
@@
 
-\begin{figure}[!h]
-\captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
-\begin{center}
-\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
-]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
-%
-\subfloat[\underline{A bowl of bananas} is on the table.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
-%
-\subfloat[A vintage photo of a \underline{cat}.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
-%
-\subfloat[A vintage photo of a \underline{dog}.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
-%
-\end{center}
-\caption{Examples of changing the object while keeping the caption fixed.}
-\label{fig:genimages3}
-\vspace{-0.3cm}
-\end{figure}
-
 As a test of model generalization, we tried generating images corresponding to captions that describe scenarios that are highly unlikely to occur in real life.
@@
 
-\begin{figure}[!h]
-\captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
-\begin{center}
-\subfloat[A stop sign is flying in blue skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
-%
-\subfloat[A herd of elephants flying in the blue skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
-%
-\subfloat[A toilet seat sits open in the grass field.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
-%
-\subfloat[A person skiing on sand clad vast desert.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\quad
-%
-\end{center}
-\caption{Examples of novel scene compositions.}
-\label{fig:genimages4}
-\vspace{-0.3cm}
-\end{figure}
-
 \subsubsection{Analysis of Attention}
@@
 The attention values of words in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
-For example, in Figure~\ref{fig:attentimages1} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
+For example, in Figure~\ref{fig:diffmodels} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
 This suggests that, in their respective sentences, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
@@
 ]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
+\\
+%
+\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
+%
+\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
+%
+\subfloat[Conv-Deconv VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
+%
+\subfloat[Fully-Conn VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
 %
 \end{center}
-\caption{Examples of most attended words while changing the background in the caption.}
-\label{fig:attentimages1}
+\caption{\textbf{Bottom}: Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.} \textbf{Top}: Examples of most attended words while changing the background in the caption.}
+\label{fig:diffmodels}
 \vspace{-0.3cm}

### 2015-10-30 iclr-paper-new.tex
update

@@
 
-\cite{kingma_vae} have introduced the Variational Auto-Encoder (VAE) which can be seen as a neural network with continous latent variables.
+\cite{kingma_vae} have introduced the Variational Auto-Encoder (VAE) which can be seen as a neural network with continuous latent variables.
 The encoder is used to approximate a posterior distribution and the decoder is used to stochastically reconstruct the data from latent variables.
@@
 
-Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$ as in \citep{bachman_sdm}, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
+Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
+Namely, the mean and variance of prior distribution $\prior(\Lat_t)$ is parameterized by:
+\begin{align}
+\mu_{t}(\hdec_{t-1}) &= tanh(W_{\mu}\hdec_{t-1}),\\
+\sigma_{t}(\hdec_{t-1}) &= {exp}\big(tanh(W_{\sigma}\hdec_{t-1})\big),
+\end{align}
+where $W_{\mu}, W_{\sigma}$ are learnt model parameters.
 It has been shown in \citep{bachman_sdm} that model performance is improved by including dependencies between latent variables.
-Namely, the mean and variance of $\prior(\Lat_t)$ is:
-\begin{align}
-\mu_{t}(\hdec_{t-1}) &= tanh(W_{\mu}\hdec_{t-1})\\
-\sigma_{t}(\hdec_{t-1}) &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
-\end{align}
 Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
@@
 \label{eq:x_hat}
-\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})),\\
+\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\big(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\big),\\
 s_{t} &= align(\hdec_{t-1}, \hlangall), \\
@@
 %\end{align}
-The final canvas $\canv_T$ is transformed using a sigmoid function to produce a conditional Bernoulli data likelihood over pixels given the latent variables and the input caption.
-
+The final canvas $\canv_T$ is transformed using a sigmoid function to produce a conditional Bernoulli data likelihood over the image pixels $\oimage$ given the latent variables $\Lat$ and the input caption $\icaption$,
+\begin{align}
+P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \sigmoid(\canv_{T,i}),
+\end{align}
+where, $\canv_{T,i}$ is the $i^{th}$ pixel on the final canvas computed at step T and $\sigmoid$ operator is a logistic sigmoid function $\sigmoid(x)=\frac{1}{1+exp(-x)}$. 
 \comm{
@@
 %The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. 
-The model is trained to maximize a variational lower bound $\loss$ on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$, where $\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat_{1:T}) + \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars P(\Lat)\right)\right) \le \log P(\oimage\given\icaption)$.
+The model is trained to maximize a variational lower bound $\loss$ on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$, where $\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat_{1:T}) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars P(\Lat)\right)\right) \le \log P(\oimage\given\icaption)$.
 %The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using the inference RNN.
@@
 \label{eq:x_hat}
-\hat{\oimage}_t &= \oimage-\sigmoid(\canv_{t-1})\\
+\hat{\oimage}_t &= \oimage-\sigmoid(\canv_{t-1}),\\
 \label{eq:read}
-r_t &= \readop(\oimage_t, \hat{\oimage}_t, \hdec_{t-1})\\
-\henc_t &= \encoder(\henc_{t-1}, [r_t, \hdec_{t-1}])\\
-\post(\Lat_t|\oimage,\icaption,\Lat_{1:t-1}) &= \mathcal{N}\left(\mu_t(\henc_t), \sigma_t(\henc_t)\right)
+r_t &= \readop(\oimage_t, \hat{\oimage}_t, \hdec_{t-1}),\\
+\henc_t &= \encoder(\henc_{t-1}, [r_t, \hdec_{t-1}]),\\
+\post(\Lat_t|\oimage,\icaption,\Lat_{1:t-1}) &= \mathcal{N}\left(\mu_t(\henc_t), \sigma_t(\henc_t)\right),
 \end{align}
@@
 
-where $\hat{\oimage}$ is the \textit{error} image, $\sigmoid$ operator is a logistic sigmoid function $\sigmoid=\frac{1}{1+exp(-x)}$ and $\henc_0$ is initialized to the learned bias.
+where $\hat{\oimage}$ is the \textit{error} image and $\henc_0$ is initialized to the learned bias.
 
@@
 \begin{align}
-\loss =  &\expectation_{Q(\Lat_{1:T}\given\icaption,\oimage)}\left[ - \log p(\oimage\given\icaption,\Lat_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
+\loss =  &\expectation_{Q(\Lat_{1:T}\given\icaption,\oimage)}\left[ \log p(\oimage\given\icaption,\Lat_{1:T}) - \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& - \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
 \end{align}
@@
 \begin{align}
-\loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ - \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) + \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption)\right) \right] \nonumber \\& + \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
+\loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) - \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption)\right) \right] \nonumber \\& - \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
 \end{align}

### 2015-10-30 iclr-paper-new.tex
u
update

@@
 Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
-Namely, the mean and variance of prior distribution $\prior(\Lat_t)$ is parameterized by:
+Namely, the mean and variance of prior distribution $\prior(\Lat_t)$ are parameterized by:
 \begin{align}
@@
 It has been shown in \citep{bachman_sdm} that model performance is improved by including dependencies between latent variables.
-Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
+An image is generated by iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
 \begin{align}
 \label{eq:x_hat}
-\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\big(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\big),\\
+\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
 s_{t} &= align(\hdec_{t-1}, \hlangall), \\
 \hdec_t &= \decoder(\hdec_{t-1}, [z_t, s_{t}]), \\
-\canv_t &= \canv_{t-1} + \writeop(\hdec_t), 
+\canv_t &= \canv_{t-1} + \writeop(\hdec_t), \\
+\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \sigmoid(\canv_{T,i}),
 \label{eq:write}
@@
 %\end{align}
-The final canvas $\canv_T$ is transformed using a sigmoid function to produce a conditional Bernoulli data likelihood over the image pixels $\oimage$ given the latent variables $\Lat$ and the input caption $\icaption$,
-\begin{align}
-P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \sigmoid(\canv_{T,i}),
-\end{align}
-where, $\canv_{T,i}$ is the $i^{th}$ pixel on the final canvas computed at step T and $\sigmoid$ operator is a logistic sigmoid function $\sigmoid(x)=\frac{1}{1+exp(-x)}$. 
-\comm{
-We extended the architecture of the DRAW network to include an additional input caption from the language model described in Sec.
-(\ref{sec:lang}).
-Similarly to the original DRAW network, the conditional DRAW network is a stochastic recurrent neural network that consists of an Inference LSTM that infers the distribution of latent variables of image $x$ given caption $y$ and then a Generative LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given caption $y$, where $\icaption$ is the input caption consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$.
-The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps as in \cite{bahdanau_mt}:
-
-Formally, the image is generated by iteratively computing the following equations for $t=1,...,T$
-\begin{align}
-\label{eq:x_hat}
-\hat{x}_t &= x-\sigmoid(\canv_{t-1})\\
-\label{eq:read}
-r_t &= \readop(x_t, \hat{x}_t, \hdec_{t-1})\\
-\henc_t &= \encoder(\henc_{t-1}, [r_t, \hdec_{t-1}])\\
-\lat_t &\sim \post(\Lat_t|\henc_t)\\
-\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1})\\
-s_{t} &= align(\hdec_{t-1}, \hlangall) \\
-\label{eq:write}
-\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
-\end{align}
-
-where $\writeop$ is the same attention operators as in \citep{gregor_draw}.
-%where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}.
-% which apply a grid of Gaussian filters by specifying coordinates of the location and scale in order to yield a desirable image patch.
-Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator computes the final sentence representation $s_t$ through a weighted sum using alignment probabilities $\alpha_{1...N}$:
-\begin{align}
-s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}.
-\end{align}
-
-The corresponding alignment probabilities $\alpha_{1...n}$ at each step are obtained by:
-\begin{align}
-e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)\\
-\alpha_{j} &= \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}.
-\end{align}
-
-Here $\hlang_{0}$ is initialized to the learned bias.
-Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. 
-}
+The $i^{th}$ pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli data likelihood over the image pixels $\oimage$ given the latent variables $\Lat$ and the input caption $\icaption$,
 
@@
 
-The terms in the variational lower bound are rearranged using the law of total expectation. Therefore, the loss function $\loss$ is calculated as follows:
+The terms in the variational lower bound are rearranged using the law of total expectation. Therefore, the objective function $\loss$ is calculated as follows:
 \begin{align}
+\label{eq:loss}
 \loss =  &\expectation_{Q(\Lat_{1:T}\given\icaption,\oimage)}\left[ \log p(\oimage\given\icaption,\Lat_{1:T}) - \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& - \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
@@
 \begin{align}
+\label{eq:loss_mc}
 \loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) - \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption)\right) \right] \nonumber \\& - \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
@@
 to sharpen the generated images, similar to \citep{denton_lapgan}.
-By fixing the prior of the adversarial generator to its mean, it gets treated as a deterministic neural network which allows us to calculate the lower bound of likelihood.
-The reconstuction loss becomes the loss between sharpened image and correct image, whereas the latent loss stays the same.
+By fixing the prior of the adversarial generator to its mean, it gets treated as a deterministic neural network that allows us to define the conditional data term in Eq.
+(\ref{eq:loss}) on the sharpened images and calculate the lower bound accordingly.
 Additionally, we noticed that sampling from the mean of the uniform distribution generated samples with less noise than if we had otherwise sampled from the uniform distribution itself.

### 2015-10-30 iclr-paper-new.tex
update

@@
 \mu_{t}(\hdec_{t-1}) &= tanh(W_{\mu}\hdec_{t-1}),\\
-\sigma_{t}(\hdec_{t-1}) &= {exp}\big(tanh(W_{\sigma}\hdec_{t-1})\big),
+\sigma_{t}(\hdec_{t-1}) &= \exp\big(tanh(W_{\sigma}\hdec_{t-1})\big),
 \end{align}
@@
 %\vspace{-0.1in}
-$\alpha_{k}^{t}=\frac{exp(v^{T}tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}exp(v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)}.$
+$\alpha_{k}^{t}=\frac{\exp(v^{T}tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}\exp(v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)}.$
 %\end{align}
@@
 &=
-\frac{1}{2}\sum_{t=1}^{T}(1 - 2\,log\,\sigma_{t}^{prior} + 2\,log\,\sigma_{t} - \frac{exp(2\,log\,\sigma_{t}) + (\mu_{t} - \mu_{t}^{prior})^{2}}{exp(2\,log\,\sigma_{t}^{prior})}) + \frac{1}{L}\sum_{l=1}^{L}log\,p(x_{t}|y,z)
+\frac{1}{2}\sum_{t=1}^{T}(1 - 2\,log\,\sigma_{t}^{prior} + 2\,log\,\sigma_{t} - \frac{\exp(2\,log\,\sigma_{t}) + (\mu_{t} - \mu_{t}^{prior})^{2}}{\exp(2\,log\,\sigma_{t}^{prior})}) + \frac{1}{L}\sum_{l=1}^{L}log\,p(x_{t}|y,z)
 \end{align}

### 2015-10-30 iclr-paper-new.tex
changes

@@
 Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set \citep{kiros_captions}.
-We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results.
+We found that the lower bound of log likelihood decreased for sharpened images, and that sharpening considerably hurt the retrieval results.
 Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.
@@
 We would also like to thank Ryan Kiros and Nitish Srivastava for helpful discussions.
-We will be releasing our code with pretrained models as soon as we clean it.
+We will be releasing our code with pretrained models
 

### 2015-10-30 supp.tex
changes

@@
 Each parameter in alignDRAW was initialized by sampling from the Gaussian distribution of mean $0$ and standard deviation of $0.01$.
-The model was trained using \textit{RMSProp} with initial learning rate of $0.001$.
+The model was trained using \textit{RMSprop} with initial learning rate of $0.001$.
 The norm of the gradients was clipped to $10.0$ during training to deal with exploding gradients.
-For all tasks we used $128$ units in the language model and $512$ units in the $tanh$ perceptron in the alignment function.
+For all tasks, $\overrightarrow{h}^{lang}_{n}$ and $\overleftarrow{h}^{lang}_{n}$ in the language model had $128$ units, and vector $v$ in \textit{align} perceptron had $512$ units.
 For the Microsoft COCO task, the alignDRAW model was trained for $18$ epochs and the learning rate was reduced by the factor of $0.1$ after $11$ epochs.
@@
 
-The LAPGAN model used for sharpening had the same configuration as the $28 \times 28$ model trained by \cite{denton_lapgan} on the edge residuals of CIFAR dataset.
-that can be found at \url{https://gist.github.com/soumith/e3f722173ea16c1ea0d9}.
+The GAN model used for sharpening had the same configuration as the $28 \times 28$ model trained by \cite{denton_lapgan} on the edge residuals of CIFAR dataset.
+The configuration can be found at \url{https://gist.github.com/soumith/e3f722173ea16c1ea0d9}.
 The model was trained for $6$ epochs.

### 2015-10-31 iclr-paper-new.tex
a

@@
 \captionsetup{font=small}
+\newcommand{\Figref}[1]{Fig.~\ref{#1}}
+\newcommand{\Secref}[1]{Sec.~\ref{#1}}
+\newcommand{\Eqref}[1]{Eq.~\ref{#1}} % The '~' stops breaking and gives correct spacing
+\newcommand{\Eqrefs}[2]{Eqs.~\ref{#1},~\ref{#2}}
+\def\beqa#1\eeqa{\begin{eqnarray}#1\end{eqnarray}}
 
@@
 In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on a canvas $c_t$ over time $t=1,...,T$.
-Our model can be viewed as a part of the sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video}.
+Our model can be viewed as a part of the sequence-to-sequence framework \citep{ilya_mt,cho_mt,nitish_video}.
 %Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
@@
 \label{sec:lang}
-Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{N}$, and $\oimage$ be the output image.
-We obtain the caption sentence representation by first transforming each word $y_{1},...,y_{N}$ to a vector representation using the Bidirectional RNN.
+Let $\icaption$ be the input caption, represented as a sequence
+of 1-of-K encoded words 
+${\bf y} = \{y_{1}, y_{2}, ..., y_{N}\}$, 
+where $K$ is the size of he vocabulary and $N$ is the length of the sequence.
+%and $\oimage$ be the output image. 
+We obtain the caption sentence representation by first 
+transforming each word $y_{n}$ to an $m$-dimensional 
+vector representation $\hlang_{n}$, $n=1,..,N$ using the Bidirectional RNN.
 In a Bidirectional RNN, the two LSTMs process the input sequence from both forward and backward directions.
 The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$.
-These hidden states are then concatenated together into the sequence $[\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
+These hidden states are then concatenated together 
+into the sequence $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, 
+with $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
 
-\begin{figure}[!t]
+\begin{figure}[t!]
 \captionsetup[subfigure]{labelformat=empty}
@@
 
-To generate images conditioned on the caption information, we extended the DRAW network \citep{gregor_draw} to include caption representation $\hlangall$ at each step, shown in Figure~\ref{fig:figmodel}.
-The conditional DRAW network is a stochastic recurrent neural network that consists of a set of latent variables $\Lat_t$ where the output is accumulated over all timesteps.
+To generate an image $\oimage$ conditioned on the caption information $\bf{y}$, 
+we extended the DRAW network~\citep{gregor_draw} to include caption representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}. 
+The conditional DRAW network is a stochastic recurrent neural network that consists of a 
+sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all timesteps.
 
-Unlike the original DRAW network where latent variables are independent unit Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
-Namely, the mean and variance of prior distribution $\prior(\Lat_t)$ are parameterized by:
-\begin{align}
-\mu_{t}(\hdec_{t-1}) &= tanh(W_{\mu}\hdec_{t-1}),\\
-\sigma_{t}(\hdec_{t-1}) &= \exp\big(tanh(W_{\sigma}\hdec_{t-1})\big),
-\end{align}
-where $W_{\mu}, W_{\sigma}$ are learnt model parameters.
-It has been shown in \citep{bachman_sdm} that model performance is improved by including dependencies between latent variables.
+Unlike the original DRAW network where latent variables are independent 
+spherical Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
+%Namely, the mean and variance of prior distribution $\prior(\Lat_t)$ are parameterized by: 
+%\begin{align}
+%\mu_{t}(\hdec_{t-1}) &= tanh(W_{\mu}\hdec_{t-1}),\\
+%\sigma_{t}(\hdec_{t-1}) &= \exp\big(tanh(W_{\sigma}\hdec_{t-1})\big),
+%\end{align}
+%where $W_{\mu}, W_{\sigma}$ are learnt model parameters. 
+Similar to \cite{bachman_sdm}, we have observed that 
+the model performance is improved by including dependencies between latent variables.
+
 An image is generated by iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):

### 2015-10-31 iclr-paper-new.tex
a

@@
 \newcommand{\hdec}{h^{gen}}
+\newcommand{\aalign}{{\textrm{align}}}
 \newcommand{\henc}{h^{infer}}
@@
 \newcommand{\lat}{z}
+\newcommand{\vv}{{\bf v}}
 \newcommand{\Lat}{Z}
@@
 We obtain the caption sentence representation by first 
-transforming each word $y_{n}$ to an $m$-dimensional 
-vector representation $\hlang_{n}$, $n=1,..,N$ using the Bidirectional RNN.
-In a Bidirectional RNN, the two LSTMs process the input sequence from both forward and backward directions.
+transforming each word $y_{i}$ to an $m$-dimensional 
+vector representation $\hlang_{i}$, $i=1,..,N$ using the Bidirectional RNN.
+In a Bidirectional RNN, the two LSTMs~\citep{hochreiter_lstm} process the input sequence from both forward and backward directions.
 The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$.
@@
 into the sequence $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, 
-with $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.
+with $\hlang_{i} = [\overrightarrow{h}^{lang}_{i}, \overleftarrow{h}^{lang}_{i}], 1\leq i\leq N$.
 
@@
 we extended the DRAW network~\citep{gregor_draw} to include caption representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}. 
-The conditional DRAW network is a stochastic recurrent neural network that consists of a 
-sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all timesteps.
+The conditional DRAW network is a stochastic recurrent neural network (LSTM) that consists of a 
+sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all $T$ timesteps.
 
@@
 spherical Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
-%Namely, the mean and variance of prior distribution $\prior(\Lat_t)$ are parameterized by: 
-%\begin{align}
-%\mu_{t}(\hdec_{t-1}) &= tanh(W_{\mu}\hdec_{t-1}),\\
-%\sigma_{t}(\hdec_{t-1}) &= \exp\big(tanh(W_{\sigma}\hdec_{t-1})\big),
-%\end{align}
-%where $W_{\mu}, W_{\sigma}$ are learnt model parameters. 
+Namely, the mean and variance of the prior distribution over $\Lat_t$ are parameterized by: 
+\begin{align}
+\lat_t \sim \prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
+\mu_{t}(\hdec_{t-1}) &= \tanh(W_{\mu}\hdec_{t-1}),\\
+\sigma_{t}(\hdec_{t-1}) &= \exp\big(\tanh(W_{\sigma}\hdec_{t-1})\big),
+\end{align}
+where $W_{\mu} \in \mathbb{R}^{D \times n}$, $W_{\sigma} \in \mathbb{R}^{D \times n}$ are 
+the learnt model parameters, where $n$ is the dimensionality of $\hdec_{t}$,
+the hidden state of the generative LSTM. 
 Similar to \cite{bachman_sdm}, we have observed that 
@@
 
-An image is generated by iteratively computing the following equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
+Formally, an image is generated by iteratively computing the following set of 
+equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
 \begin{align}
@@
 \lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
-s_{t} &= align(\hdec_{t-1}, \hlangall), \\
+s_{t} &= \aalign(\hdec_{t-1}, \hlangall), \\
 \hdec_t &= \decoder(\hdec_{t-1}, [z_t, s_{t}]), \\
@@
 \end{align}
-where $\mathit{LSTM}$ is a standard \textit{Long Short-Term Memory} architecture \citep{hochreiter_lstm} with \textit{forget} gates \citep{gers_forget}, $[v_1, v_2]$ is the concatenation of the vectors $v_1$ and $v_2$ into the single vector, and $\canv_0$, $\hdec_0$ are initialized to the the learned biases.
-The $\writeop$ operator applies an array of Gaussian filters on the patch that places it on the canvas with smoothly varying location and zoom (same as in \citep{gregor_draw}).
+where $\decoder$
+is the function defined by the LSTM network with \textit{forget} gates 
+\citep{gers_forget} at a single time-step. To generate the next 
+hidden state $\hdec_t$, 
+the decoder $\decoder$ takes its previous hidden state $\hdec_{t-1}$ and
+combines it with the input from both the sampled latent variable $z_t$ and 
+caption representation $s_t$ at time-step $t$, where
+$[v_1, v_2]$ denotes the concatenation of the vectors $v_1$ and $v_2$ into the single vector.
+The initial state of the generative LSTM is set to $\hdec_0 = \tanh(b)$, with learnt bias $b$.
+
 The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}. 
-Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities $\alpha_{1...N}^{t}$:
-%\begin{align}
-%\vspace{-0.1in}
-$s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.$
-%\end{align}
-The corresponding alignment probabilities $\alpha_{1...N}^{t}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies using a $tanh$ perceptron scaled by vector $v$ and then normalizing them, as in \citep{bahdanau_mt}:
-%\begin{align}
-%\vspace{-0.1in}
-$\alpha_{k}^{t}=\frac{\exp(v^{T}tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}\exp(v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)}.$
-%\end{align}
-The $i^{th}$ pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli data likelihood over the image pixels $\oimage$ given the latent variables $\Lat$ and the input caption $\icaption$,
+Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities
+$\alpha_{1...N}^{t}$:
+\beqa
+s_t=\aalign(\hdec_{t-1}, \hlangall) = \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.
+\eeqa
+The corresponding alignment probabilities $\alpha_{1...N}^{t}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies using a $\tanh$ perceptron scaled by vector $\vv$
+and then normalizing them: %, as in \citep{bahdanau_mt}:
+\beqa
+\alpha_{k}^{t}=\frac{\exp(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}
+\exp(\vv^{\top}\tanh(U\hlang_{j} + W\hdec_{t} + b)}, 
+\eeqa
+where $U,W,b$ are the learned model parameters.
+
+%is a standard \textit{Long Short-Term Memory} architecture \citep{hochreiter_lstm} with \textit{forget} gates \citep{gers_forget}, $[v_1, v_2]$ is the concatenation of the vectors $v_1$ and $v_2$ into the single vector, and $\canv_0$, $\hdec_0$ are initialized to the the learned biases.
+
+The $\writeop$ operator applies an array of Gaussian filters on the patch that places it on the canvas $c_t$ with smoothly varying location and zoom (same as in \citep{gregor_draw}), with 
+$c_0=0$. 
+Finally, the $i^{th}$ pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli data likelihood over the image pixels $\oimage$ given the latent variables $\Lat$ and the input caption $\icaption$,
 

### 2015-10-31 iclr-paper-new.tex
update model description

@@
 \section{Model}
+\label{sec:model}
 Our proposed model 
@@
 \end{center}
-\caption{The image of the proposed model}
+\caption{The diagram shows our model learns to draw images by learning an alignment of input captions and generating canvas. The details are explained in Section \ref{sec:model}.}
 \label{fig:figmodel}

### 2015-11-01 iclr-paper-new.tex
moved figures fixed spelling

@@
 \newcommand{\hdec}{h^{gen}}
-\newcommand{\aalign}{{\textrm{align}}}
+\newcommand{\aalign}{{\mathit{align}}}
 \newcommand{\henc}{h^{infer}}
@@
 ${\bf y} = \{y_{1}, y_{2}, ..., y_{N}\}$, 
-where $K$ is the size of he vocabulary and $N$ is the length of the sequence.
+where $K$ is the size of the vocabulary and $N$ is the length of the sequence.
 %and $\oimage$ be the output image. 
@@
 transforming each word $y_{i}$ to an $m$-dimensional 
-vector representation $\hlang_{i}$, $i=1,..,N$ using the Bidirectional RNN.
-In a Bidirectional RNN, the two LSTMs~\citep{hochreiter_lstm} process the input sequence from both forward and backward directions.
+vector representation $\hlang_{i}$, $i=1,..,N$ using the Bidirectional RNN. In a Bidirectional RNN, the two LSTMs~\citep{hochreiter_lstm} with \textit{forget} gates 
+\citep{gers_forget} process the input sequence from both forward and backward directions.
 The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$.
@@
 where $\decoder$
-is the function defined by the LSTM network with \textit{forget} gates 
-\citep{gers_forget} at a single time-step. To generate the next 
+is the function defined by the LSTM network at a single time-step. To generate the next 
 hidden state $\hdec_t$, 
@@
 \label{fig:genimages3}
-\vspace{-0.3cm}
+%\vspace{-0.3cm}
 \end{figure}
@@
 
+Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
+The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
+Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
+
+Despite a large number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
+However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
+In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
+This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
+
+As a test of model generalization, we tried generating images corresponding to captions that describe scenarios that are highly unlikely to occur in real life.
+These captions describe a common object doing unusual things
+or set in a strange location.
+%that are impossible in the real world. 
+%considering the physics of real world. 
+Even though some of these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
+Nevertheless, as you can see in Figure~\ref{fig:genimages4}, the model managed to generate reasonable images.
+
+\subsubsection{Analysis of Attention}
 \begin{figure}[!t]
@@
 \label{fig:genimages2}
-\vspace{-0.3cm}
+%\vspace{-0.3cm}
 \end{figure}
-
-Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
-The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
-Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
-
-Despite a large number of ways of changing colors and backgrounds in descriptions, in general we found that the model made appropriate changes as long as some similar pattern was present in the training set.
-However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
-In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
-This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
-
-As a test of model generalization, we tried generating images corresponding to captions that describe scenarios that are highly unlikely to occur in real life.
-These captions describe a common object doing unusual things
-or set in a strange location.
-%that are impossible in the real world. 
-%considering the physics of real world. 
-Even though some of these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
-Nevertheless, as you can see in Figure~\ref{fig:genimages4}, the model managed to generate reasonable images.
-
-\subsubsection{Analysis of Attention}
 
@@
 
+We also tried to analyze the way the model generated images.
+Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular timesteps.
+
+\subsubsection{Comparison With Other Models}
+
+Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffers from its own specific drawbacks.
+Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
+To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
 \begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
+%\vspace{-0.3cm}
 \begin{center}
@@
 \label{fig:diffmodels}
-\vspace{-0.3cm}
+%\vspace{-0.3cm}
 \end{figure}
-
-We also tried to analyze the way the model generated images.
-Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular timesteps.
-
-\subsubsection{Comparison With Other Models}
-Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffers from its own specific drawbacks.
-Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
 
@@
 Nevertheless, the results are reported on Table~\ref{tab:results}.
-
-\begin{table}[!t]
-\begin{center}
-\begin{tabulary}{\linewidth}{c || c c c c c || c c}
-\hline
-\multicolumn{7}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
-\hline
-& \multicolumn{5}{c||}{Image Search} & \multicolumn{1}{c}{Image Similarity} \\
-%\cline{2-6}
-\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
-\hline
-\hline
-LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
-\hline
-Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
-\end{tabulary}
-\end{center}
-\caption{Results of different models.}
-\label{tab:results}
-\vspace{-0.3cm}
-\end{table}
 
@@
 \section{Acknowledgements}
+\begin{table}[!t]
+\begin{center}
+\begin{tabulary}{\linewidth}{c || c c c c c || c c}
+\hline
+\multicolumn{7}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
+\hline
+& \multicolumn{5}{c||}{Image Search} & \multicolumn{1}{c}{Image Similarity} \\
+%\cline{2-6}
+\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
+\hline
+\hline
+LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
+\hline
+Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+\end{tabulary}
+\end{center}
+\caption{Results of different models.}
+\label{tab:results}
+%\vspace{-0.3cm}
+\end{table}
 We would like to thank developers of Theano \citep{theano} for a powerful software and authors of \citep{denton_lapgan} for open sourcing their code.
 We would also like to thank Ryan Kiros and Nitish Srivastava for helpful discussions.
-We will be releasing our code with pretrained models
+We will be releasing our code with pretrained models.
 

### 2015-11-01 iclr-paper-new.tex
a

--- 2015-10-31 iclr-paper-new.tex
@@
 The conditional DRAW network is a stochastic recurrent neural network (LSTM) that consists of a 
-sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all $T$ timesteps.
+sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all $T$ time-steps.
 
@@
 where $W_{\mu} \in \mathbb{R}^{D \times n}$, $W_{\sigma} \in \mathbb{R}^{D \times n}$ are 
-the learnt model parameters, where $n$ is the dimensionality of $\hdec_{t}$,
+the learned model parameters, and $n$ is the dimensionality of $\hdec_{t}$,
 the hidden state of the generative LSTM. 
@@
 \lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
-s_{t} &= \aalign(\hdec_{t-1}, \hlangall), \\
+s_{t} &= align(\hdec_{t-1}, \hlangall), \\
+\label{eq:decoder}
 \hdec_t &= \decoder(\hdec_{t-1}, [z_t, s_{t}]), \\
+\label{eq:writeop}
 \canv_t &= \canv_{t-1} + \writeop(\hdec_t), \\
-\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \sigmoid(\canv_{T,i}),
+\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \textrm{Bern}(\sigmoid(\canv_{T,i})). 
 \label{eq:write}
@@
 \end{align}
-where $\decoder$
-is the function defined by the LSTM network with \textit{forget} gates 
-\citep{gers_forget} at a single time-step. To generate the next 
-hidden state $\hdec_t$, 
-the decoder $\decoder$ takes its previous hidden state $\hdec_{t-1}$ and
-combines it with the input from both the sampled latent variable $z_t$ and 
-caption representation $s_t$ at time-step $t$, where
-$[v_1, v_2]$ denotes the concatenation of the vectors $v_1$ and $v_2$ into the single vector.
-The initial state of the generative LSTM is set to $\hdec_0 = \tanh(b)$, with learnt bias $b$.
-
-The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}. 
+The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps~\citep{bahdanau_mt}.
 Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities
@@
 \beqa
-s_t=\aalign(\hdec_{t-1}, \hlangall) = \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.
+s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.
 \eeqa
-The corresponding alignment probabilities $\alpha_{1...N}^{t}$ at each step are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t}$, by first calculating energies using a $\tanh$ perceptron scaled by vector $\vv$
+The corresponding alignment probabilities $\alpha_{1...N}^{t}$ are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t-1}$, by first calculating energies using a $\tanh$ perceptron scaled by vector $\vv$
 and then normalizing them: %, as in \citep{bahdanau_mt}:
 \beqa
-\alpha_{k}^{t}=\frac{\exp(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t} + b)}{\sum_{j=1}^{N}
-\exp(\vv^{\top}\tanh(U\hlang_{j} + W\hdec_{t} + b)}, 
+\alpha_{k}^{t}=\frac{\exp(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)}{\sum_{j=1}^{N}
+\exp(\vv^{\top}\tanh(U\hlang_{j} + W\hdec_{t-1} + b)},
 \eeqa
-where $U,W,b$ are the learned model parameters.
+where $U \in \mathbb{R}^{d \times m}$, $W \in \mathbb{R}^{d \times n}$ and 
+$b \in \mathbb{R}^{d \times 1} $ are the learned model parameters of the alignment model.
 
-%is a standard \textit{Long Short-Term Memory} architecture \citep{hochreiter_lstm} with \textit{forget} gates \citep{gers_forget}, $[v_1, v_2]$ is the concatenation of the vectors $v_1$ and $v_2$ into the single vector, and $\canv_0$, $\hdec_0$ are initialized to the the learned biases.
+The $\decoder$ function of \Eqref{eq:decoder} 
+is defined by the LSTM network with \textit{forget} gates 
+\citep{gers_forget} at a single time-step. To generate the next 
+hidden state $\hdec_t$, 
+the decoder $\decoder$ takes the previous hidden state $\hdec_{t-1}$ and
+combines it with the input from both the latent sample $z_t$ and 
+sentence representation $s_t$, where
+$[v_1, v_2]$ denotes the concatenation of $v_1$ and $v_2$ into the single vector.
+The initial state of the generative LSTM is set to $\hdec_0 = \tanh(b)$, 
+with learned bias $b$.
 
-The $\writeop$ operator applies an array of Gaussian filters on the patch that places it on the canvas $c_t$ with smoothly varying location and zoom (same as in \citep{gregor_draw}), with 
-$c_0=0$. 
-Finally, the $i^{th}$ pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli data likelihood over the image pixels $\oimage$ given the latent variables $\Lat$ and the input caption $\icaption$,
+The output of the decoder $\hdec_t$ is then passed through the $\writeop$ operator 
+which is added to a cumulative canvas matrix $c_t$, with $c_0=0$ (\Eqref{eq:writeop}). 
+The $\writeop$ operator
+applies an array of Gaussian filters on the patch, placing 
+it on the canvas $c_t$ with smoothly varying location and zoom (same as 
+defined in \citep{gregor_draw}).
+
+Finally, each pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli 
+distribution with the mean vector $\sigmoid(\canv_{T})$ over the image pixels 
+$\oimage$ given the latent variables $\Lat_{1:T}$ and the input caption $\icaption$,
 
@@
 The model correctly displayed the specified digits at the described positions and even managed to generalize reasonably to the configurations that were never present during training.
-In the case of generating two digits, the model attended to the digit in the caption it was drawing at that particular timestep while completely ignoring the other digit. 
+In the case of generating two digits, the model attended to the digit in the caption it was drawing at that particular time-step while completely ignoring the other digit. 
 Similarly, when generating one digit the model attended to the digit in the caption during the whole generation process.
@@
 We also tried to analyze the way the model generated images.
-Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular timesteps.
+Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular time-steps.
 

### 2015-11-01 iclr-paper-new.tex
update write

--- 2015-10-31 iclr-paper-new.tex
@@
 \newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
+\newcommand{\gFilterx}{F_x}
+\newcommand{\gFiltery}{F_y}
+\newcommand{\WriteFunc}{W}
+\newcommand{\real}{\mathbb{R}}
 
@@
 
-The $\writeop$ operator applies an array of Gaussian filters on the patch that places it on the canvas $c_t$ with smoothly varying location and zoom (same as in \citep{gregor_draw}), with 
+The $\writeop$ operator predicts two arrays of Gaussian filters using a parametric mapping $\gFilterx:\real$ applies an array of Gaussian filters on the patch that places it on the canvas $c_t$ with smoothly varying location and zoom (same as in \citep{gregor_draw}), with
 $c_0=0$. 
+\beqa
+\label{eq:write}
+\Delta \canv_{t-1} = \canv_{t} - \canv_{t-1} = \writeop(\hdec_t) = \gFilterx(\hdec_t) \WriteFunc(\hdec_t) \gFiltery(\hdec_t)^\top, 
+\eeqa
+
 Finally, the $i^{th}$ pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli data likelihood over the image pixels $\oimage$ given the latent variables $\Lat$ and the input caption $\icaption$,

### 2015-11-01 iclr-paper-new.tex
a

--- 2015-11-01 iclr-paper-new.tex
@@
 \def\beqa#1\eeqa{\begin{eqnarray}#1\end{eqnarray}}
+\newcommand{\FIXME}[1]{\textcolor{red}{[#1]}}
 
@@
 \begin{align}
-\lat_t \sim \prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
-\mu_{t}(\hdec_{t-1}) &= \tanh(W_{\mu}\hdec_{t-1}),\\
-\sigma_{t}(\hdec_{t-1}) &= \exp\big(\tanh(W_{\sigma}\hdec_{t-1})\big),
+\lat_t \sim \prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu(\hdec_{t-1}), \sigma(\hdec_{t-1})\bigg), \nonumber \\
+\mu(\hdec_{t-1}) &= \tanh(W_{\mu}\hdec_{t-1}),\nonumber \\
+\sigma(\hdec_{t-1}) &= \exp\big(\tanh(W_{\sigma}\hdec_{t-1})\big), \nonumber 
 \end{align}
@@
 Formally, an image is generated by iteratively computing the following set of 
-equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}):
+equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}), with $\hdec_0$ and $c_0$
+initialized to learned biases:
 \begin{align}
 \label{eq:x_hat}
-\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
+\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu(\hdec_{t-1}), \sigma(\hdec_{t-1})\bigg),\\
 s_{t} &= align(\hdec_{t-1}, \hlangall), \\
@@
 \label{eq:write}
-%\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
 \end{align}
-The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps~\citep{bahdanau_mt}.
+The $align$ function is used to compute the alignment between the input caption and intermediate image generative steps \citep{bahdanau_mt}. 
 Given the caption representation from the language model, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, the $align$ operator outputs a dynamic sentence representation $s_t$ at each step through a weighted sum using alignment probabilities
@@
 \beqa
-s_t=align(\hdec_{t-1}, \hlangall) = \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.
+s_t=align(\hdec_{t-1}, \hlangall) = 
+\alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.
+\nonumber 
 \eeqa
@@
 \exp(\vv^{\top}\tanh(U\hlang_{j} + W\hdec_{t-1} + b)},
+\nonumber 
 \eeqa
@@
 $[v_1, v_2]$ denotes the concatenation of $v_1$ and $v_2$ into the single vector.
-The initial state of the generative LSTM is set to $\hdec_0 = \tanh(b)$, 
-with learned bias $b$.
 
 The output of the decoder $\hdec_t$ is then passed through the $\writeop$ operator 
-which is added to a cumulative canvas matrix $c_t$, with $c_0=0$ (\Eqref{eq:writeop}). 
+which is added to a cumulative canvas matrix $c_t$ (\Eqref{eq:writeop}). 
+\FIXME{Jimmy, could you put more about the write operator}. 
 The $\writeop$ operator
@@
 %The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. 
-The model is trained to maximize a variational lower bound $\loss$ on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$, where $\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat_{1:T}) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars P(\Lat)\right)\right) \le \log P(\oimage\given\icaption)$.
+The model is trained to maximize a variational lower bound $\loss$ 
+on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$:
+\beqa
+\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars 
+  P(\Lat)\right)\right) \le \log P(\oimage\given\icaption).
+\eeqa
 %The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using the inference RNN. 
-Similarly to the DRAW network, the inference RNN produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using an array of Gaussian filters (inverse of $\writeop$), over the training image at each step.
+Similar to the DRAW model, the inference recurrent network 
+produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using an array of Gaussian filters (inverse of $\writeop$), over the training image at each step.
+Specifically,
 \begin{align}
@@
 \henc_t &= \encoder(\henc_{t-1}, [r_t, \hdec_{t-1}]),\\
-\post(\Lat_t|\oimage,\icaption,\Lat_{1:t-1}) &= \mathcal{N}\left(\mu_t(\henc_t), \sigma_t(\henc_t)\right),
+\post(\Lat_t|\oimage,\icaption,\Lat_{1:t-1}) &= \mathcal{N}\left(\mu(\henc_t), \sigma(\henc_t)\right),
 \end{align}
-
 %The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$. 
-
 %The reconstruction loss $\rloss$ equals to $\frac{1}{L}\sum_{l=1}^{L}(log\,p(x_{t}|y,z)$ where $L$ is the number of samples used during training, which was set to $1$ in our experiments.
-
 %The latent loss $\lloss$ is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term.
@@
 %\end{align}
-
-where $\hat{\oimage}$ is the \textit{error} image and $\henc_0$ is initialized to the learned bias.
+where $\hat{\oimage}$ is the \textit{error} image and $\henc_0$ is initialized to the learned bias $b$. 
+Note that the inference encoder $\encoder$ 
+takes as its input both the output
+of the $\readop$ operator $r_t$, which depends on the original input image $\oimage$,
+and the previous state of the 
+generative decoder $\hdec_{t-1}$, which depends on the latent sample history $z_{1:t-1}$ and 
+dynamic sentence representation $s_t$ (see \Eqref{eq:decoder}). 
+Hence, the approximate posterior $Q$ will depend on the input image $\oimage$,
+the corresponding caption $\icaption$, and the latent history $\Lat_{1:t-1}$, except  
+for the first step $Q(Z_1 | \oimage)$, which depends only on $\oimage$.
 
@@
 \end{align}
-We used only one sample during training for all our experiments.
+In all of our experiments, we used only a single sample 
+from $\post(\Lat_{1:T}\given\icaption, \oimage)$ during training.
+
 \comm{
@@
 
-\begin{figure}[!t]
-\captionsetup[subfigure]{labelformat=empty}
-\begin{center}
-\includegraphics[width=0.99\textwidth]{figures/new/mnist/test3.pdf}\quad
-%
-\end{center}
-\caption{Examples of generating $60 \times 60$ MNIST images corresponding to respective captions.
-The captions on the \textbf{left column} were part of the training set.
-The digits described in the captions on the \textbf{right column} were hidden during training for the respective configurations.}
-\label{fig:figmnist}
-\vspace{-0.3cm}
-\end{figure}
-
 During the image generation step, we discard the inference network and instead sample from the prior distribution. 
-Due to the blurriness of samples generated by the DRAW model, we perform an additional post processing step where we use a 
+Due to the blurriness of samples generated by the DRAW model, we perform an additional post processing step where we use an 
 adversarial network trained on residuals of a Laplacian pyramid conditioned on the skipthought representation \citep{kiros_skipthought} of the captions 
 to sharpen the generated images, similar to \citep{denton_lapgan}.
-By fixing the prior of the adversarial generator to its mean, it gets treated as a deterministic neural network that allows us to define the conditional data term in Eq.
-(\ref{eq:loss}) on the sharpened images and calculate the lower bound accordingly.
-Additionally, we noticed that sampling from the mean of the uniform distribution generated samples with less noise than if we had otherwise sampled from the uniform distribution itself.
+By fixing the prior of the adversarial generator to its mean, it gets treated as a deterministic neural network that allows us to define the conditional data term in \Eqref{eq:loss} on the sharpened images and
+estimate the variational lower bound accordingly. 
+%Additionally, we noticed that sampling from the mean of the uniform distribution generated samples with less noise than if we had otherwise sampled from the uniform distribution itself.
 
-%\newpage
 \section{Experiments}
-\subsection{MNIST With Captions}
-As a first experiment, we trained our proposed model on the MNIST dataset with artificial captions.
-Either one or two digits from the MNIST training dataset were placed on a $60 \times 60$ blank image.
-One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image.
-Two digits were either placed horizontally or vertically in nonoverlapping fashion.
-While many generative models were trained on the binarized version of the MNIST dataset, we trained directly on pixel intensities with a binary cross-entropy cost function.
-
-The generated images together with the attention alignments are displayed in Figure~\ref{fig:figmnist}.
-The model correctly displayed the specified digits at the described positions and even managed to generalize reasonably to the configurations that were never present during training.
-In the case of generating two digits, the model attended to the digit in the caption it was drawing at that particular time-step while completely ignoring the other digit. 
-Similarly, when generating one digit the model attended to the digit in the caption during the whole generation process.
-However in both cases, the model placed small attention values on the words describing the position of digits in the images.
-
 \subsection{Microsoft COCO}
@@
 \label{fig:diffmodels}
-\vspace{-0.3cm}
+%\vspace{-0.3cm}
 \end{figure}
@@
 
-\begin{table}[!t]
-\begin{center}
-\begin{tabulary}{\linewidth}{c || c c c c c || c c}
-\hline
-\multicolumn{7}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
-\hline
-& \multicolumn{5}{c||}{Image Search} & \multicolumn{1}{c}{Image Similarity} \\
-%\cline{2-6}
-\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
-\hline
-\hline
-LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
-\hline
-Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
-\end{tabulary}
-\end{center}
-\caption{Results of different models.}
-\label{tab:results}
-\vspace{-0.3cm}
-\end{table}
-
 % \begin{table}[!t]
@@
 
-%\newpage
-\section{Acknowledgements}
-We would like to thank developers of Theano \citep{theano} for a powerful software and authors of \citep{denton_lapgan} for open sourcing their code.
-We would also like to thank Ryan Kiros and Nitish Srivastava for helpful discussions.
-We will be releasing our code with pretrained models
+\begin{table}[!t]
+\begin{center}
+\begin{tabulary}{\linewidth}{c || c c c c c || c c}
+\hline
+\multicolumn{7}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
+\hline
+& \multicolumn{5}{c||}{Image Search} & \multicolumn{1}{c}{Image Similarity} \\
+%\cline{2-6}
+\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
+\hline
+\hline
+LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
+\hline
+Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+\end{tabulary}
+\end{center}
+\caption{Results of different models.}
+\label{tab:results}
+%\vspace{-0.3cm}
+\end{table}
+
+\subsection*{Acknowledgements}
+This work was supported by Samsung and IARPA, Raytheon BBN Contract No. D11PC20071.
+We would like to thank developers of Theano \citep{theano}, the authors of \citep{denton_lapgan} for open sourcing their code, and Ryan Kiros and Nitish Srivastava for helpful discussions. 
+%We will be releasing our code with pretrained models.
 

### 2015-11-01 supp.tex
a

@@
+\section{MNIST With Captions}
+
+\begin{figure}[!t]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\includegraphics[width=0.99\textwidth]{figures/new/mnist/test3.pdf}\quad
+%
+\end{center}
+\caption{Examples of generating $60 \times 60$ MNIST images corresponding to respective captions.
+The captions on the \textbf{left column} were part of the training set.
+The digits described in the captions on the \textbf{right column} were hidden during training for the respective configurations.}
+\label{fig:figmnist}
+\vspace{-0.3cm}
+\end{figure}
+
+As an additional experiment, we trained our proposed model on the MNIST dataset with artificial captions.
+Either one or two digits from the MNIST training dataset were placed on a $60 \times 60$ blank image.
+One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image.
+Two digits were either placed horizontally or vertically in nonoverlapping fashion.
+While many generative models were trained on the binarized version of the MNIST dataset, we trained directly on pixel intensities with a binary cross-entropy cost function.
+
+The generated images together with the attention alignments are displayed in Figure~\ref{fig:figmnist}.
+The model correctly displayed the specified digits at the described positions and even managed to generalize reasonably to the configurations that were never present during training.
+In the case of generating two digits, the model attended to the digit in the caption it was drawing at that particular time-step while completely ignoring the other digit.
+Similarly, when generating one digit the model attended to the digit in the caption during the whole generation process.
+However in both cases, the model placed small attention values on the words describing the position of digits in the images.
+
 \section{Training Details}

### 2015-11-01 iclr-paper-new.tex
update

--- 2015-11-01 iclr-paper-new.tex
@@
 \newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
+\newcommand{\gFilterx}{F_x}
+\newcommand{\gFiltery}{F_y}
+\newcommand{\WriteFunc}{K}
+\newcommand{\real}{\mathbb{R}}
 
@@
 \subsection{Image Model: the Conditional DRAW Network}
-
+\label{sec:image_model}
 To generate an image $\oimage$ conditioned on the caption information $\bf{y}$,
-we extended the DRAW network~\citep{gregor_draw} to include caption representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}. 
-The conditional DRAW network is a stochastic recurrent neural network (LSTM) that consists of a 
-sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all $T$ time-steps.
+we extended the DRAW network~\citep{gregor_draw} to include caption
+representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}.
+The conditional DRAW network is a stochastic recurrent neural network (LSTM)
+that consists of a sequence of latent variables $\Lat_t \in \mathbb{R}^D$,
+$t=1,..,T$, where the output is accumulated over all $T$ time-steps. For
+simplicity in notation, the images $\oimage\in\real^{h\times w}$ are assumed to have size $h$-by-$w$ and only one color channel.
 
 Unlike the original DRAW network where latent variables are independent 
-spherical Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
+spherical Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous hidden states of the generative LSTM $\hdec_{t-1}$, except for $\prior(\Lat_1) = \mathcal{N}(0, I)$.
 Namely, the mean and variance of the prior distribution over $\Lat_t$ are parameterized by: 
 \begin{align}
-\lat_t \sim \prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
-\mu_{t}(\hdec_{t-1}) &= \tanh(W_{\mu}\hdec_{t-1}),\\
-\sigma_{t}(\hdec_{t-1}) &= \exp\big(\tanh(W_{\sigma}\hdec_{t-1})\big),
+\prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu(\hdec_{t-1}), \sigma(\hdec_{t-1})\bigg), \nonumber \\
+\mu(\hdec_{t-1}) &= \tanh(W_{\mu}\hdec_{t-1}),\nonumber \\
+\sigma(\hdec_{t-1}) &= \exp\big(\tanh(W_{\sigma}\hdec_{t-1})\big), \nonumber 
 \end{align}
@@
 \label{eq:x_hat}
-\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu_{t}(\hdec_{t-1}), \sigma_{t}(\hdec_{t-1})\bigg),\\
+\lat_t &\sim \prior(\Lat_t|\Lat_{1:t-1}) = \mathcal{N}\bigg(\mu(\hdec_{t-1}), \sigma(\hdec_{t-1})\bigg),\\
 s_{t} &= align(\hdec_{t-1}, \hlangall), \\
@@
 hidden state $\hdec_t$, 
-the decoder $\decoder$ takes the previous hidden state $\hdec_{t-1}$ and
+the $\decoder$ takes the previous hidden state $\hdec_{t-1}$ and
 combines it with the input from both the latent sample $z_t$ and 
@@
 
-The output of the decoder $\hdec_t$ is then passed through the $\writeop$ operator 
-which is added to a cumulative canvas matrix $c_t$, with $c_0=0$ (\Eqref{eq:writeop}). 
-The $\writeop$ operator
-applies an array of Gaussian filters on the patch, placing 
-it on the canvas $c_t$ with smoothly varying location and zoom (same as 
-defined in \citep{gregor_draw}).
+The output of the $\decoder$ function $\hdec_t$ is then passed through the $\writeop$ operator 
+which is added to a cumulative canvas matrix $c_t \in \real^{h \times w}$, with $c_0=0$ (\Eqref{eq:writeop}). 
+%\FIXME{Jimmy, could you put more about the write operator}. 
+%The $\writeop$ operator
+%applies an array of Gaussian filters on the patch, placing 
+%it on the canvas $c_t$ with smoothly varying location and zoom (same as 
+%defined in \citep{gregor_draw}). 
+The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ with smoothly varying locations and scales predicted from the generative LSTM hidden state $\hdec_t$ (same as to \citep{gregor_draw}).
+The Gaussian filter banks apply to the generated $p$-by-$p$ image patch $\WriteFunc(\hdec_t) \in \real^{p\times p}$, placing it onto the canvas $c_t$.
+\beqa
+\label{eq:write}
+\Delta \canv_{t-1} = \canv_{t} - \canv_{t-1} = \writeop(\hdec_t) = \gFilterx(\hdec_t) \WriteFunc(\hdec_t) \gFiltery(\hdec_t)^\top, 
+\eeqa
 
@@
 distribution with the mean vector $\sigmoid(\canv_{T})$ over the image pixels 
-$\oimage$ given the latent variables $\Lat_{1:T}$ and the input caption $\icaption$,
+$\oimage$ given the latent variables $\Lat_{1:T}$ and the input caption $\icaption$.
 
 \subsection{Learning}
-
+\label{sec:learning}
 %The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. 
-The model is trained to maximize a variational lower bound $\loss$ on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$, where $\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat_{1:T}) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars P(\Lat)\right)\right) \le \log P(\oimage\given\icaption)$.
+The model is trained to maximize a variational lower bound $\loss$ 
+on the marginal likelihood of the correct image $\oimage$ given the input caption $\icaption$:
+\beqa
+\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\bigg(\log P(\oimage\given\icaption, \Lat) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars 
+  P(\Lat)\right)\bigg) \le \log P(\oimage\given\icaption).
+\eeqa
 %The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using the inference RNN. 
-Similarly to the DRAW network, the inference RNN produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using an array of Gaussian filters (inverse of $\writeop$), over the training image at each step.
+Similar to the DRAW model, the inference recurrent network 
+produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using two array of 1D Gaussian filters (inverse of $\writeop$ from section \ref{sec:image_model}), over the training image at each step.
+Specifically,
 \begin{align}
@@
 \end{align}
-
 %The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$. 
-
 %The reconstruction loss $\rloss$ equals to $\frac{1}{L}\sum_{l=1}^{L}(log\,p(x_{t}|y,z)$ where $L$ is the number of samples used during training, which was set to $1$ in our experiments.
-
 %The latent loss $\lloss$ is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term.
@@
 %\end{align}
-
 where $\hat{\oimage}$ is the \textit{error} image and $\henc_0$ is initialized to the learned bias.

### 2015-11-01 iclr-paper-new.tex

--- 2015-11-01 iclr-paper-new.tex
@@
 \subsection{Image Model: the Conditional DRAW Network}
-
+\label{sec:image_model}
 To generate an image $\oimage$ conditioned on the caption information $\bf{y}$,
-we extended the DRAW network~\citep{gregor_draw} to include caption representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}. 
-The conditional DRAW network is a stochastic recurrent neural network (LSTM) that consists of a 
-sequence of latent variables $\Lat_t \in \mathbb{R}^D$, $t=1,..,T$, where the output is accumulated over all $T$ time-steps.
+we extended the DRAW network~\citep{gregor_draw} to include caption
+representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}.
+The conditional DRAW network is a stochastic recurrent neural network (LSTM)
+that consists of a sequence of latent variables $\Lat_t \in \mathbb{R}^D$,
+$t=1,..,T$, where the output is accumulated over all $T$ time-steps. For
+simplicity in notation, the images $\oimage\in\real^{h\times w}$ are assumed to have size $h$-by-$w$ and only one color channel.
 
 Unlike the original DRAW network where latent variables are independent 
-spherical Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous recurrent hidden states $\hdec_{t-1}$, except for $\lat_1 \sim \prior(\Lat_1) = \mathcal{N}(0, I)$.
+spherical Gaussians $\mathcal{N}(0, I)$, the latent variables in the proposed alignDRAW model have their mean and variance depend on the previous hidden states of the generative LSTM $\hdec_{t-1}$, except for $\prior(\Lat_1) = \mathcal{N}(0, I)$.
 Namely, the mean and variance of the prior distribution over $\Lat_t$ are parameterized by: 
 \begin{align}
-\lat_t \sim \prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu(\hdec_{t-1}), \sigma(\hdec_{t-1})\bigg), \nonumber \\
+\prior(\Lat_t|\Lat_{1:t-1}) &= \mathcal{N}\bigg(\mu(\hdec_{t-1}), \sigma(\hdec_{t-1})\bigg), \nonumber \\
 \mu(\hdec_{t-1}) &= \tanh(W_{\mu}\hdec_{t-1}),\nonumber \\
@@
 
-The output of the decoder $\hdec_t$ is then passed through the $\writeop$ operator 
-which is added to a cumulative canvas matrix $c_t$ (\Eqref{eq:writeop}). 
-\FIXME{Jimmy, could you put more about the write operator}. 
-The $\writeop$ operator
-applies an array of Gaussian filters on the patch, placing 
-it on the canvas $c_t$ with smoothly varying location and zoom (same as 
-defined in \citep{gregor_draw}).
+The output of the $\decoder$ function $\hdec_t$ is then passed through the $\writeop$ operator 
+which is added to a cumulative canvas matrix $c_t \in \real^{h \times w}$, with $c_0=0$ (\Eqref{eq:writeop}). 
+%\FIXME{Jimmy, could you put more about the write operator}. 
+%The $\writeop$ operator
+%applies an array of Gaussian filters on the patch, placing 
+%it on the canvas $c_t$ with smoothly varying location and zoom (same as 
+%defined in \citep{gregor_draw}). 
+The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ with smoothly varying locations and scales predicted from the generative LSTM hidden state $\hdec_t$ (same as defined in \citep{gregor_draw}).
+The Gaussian filter banks apply to the generated $p$-by-$p$ image patch $\WriteFunc(\hdec_t) \in \real^{p\times p}$, placing it onto the canvas $c_t$:
+\beqa
+\label{eq:write}
+\Delta \canv_{t-1} = \canv_{t} - \canv_{t-1} = \writeop(\hdec_t) = \gFilterx(\hdec_t) \WriteFunc(\hdec_t) \gFiltery(\hdec_t)^\top. 
+\eeqa
 
@@
 distribution with the mean vector $\sigmoid(\canv_{T})$ over the image pixels 
-$\oimage$ given the latent variables $\Lat_{1:T}$ and the input caption $\icaption$,
+$\oimage$ given the latent variables $\Lat_{1:T}$ and the input caption $\icaption$.
 
 \subsection{Learning}
-
+\label{sec:learning}
 %The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. 
@@
 \beqa
-\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\left(\log P(\oimage\given\icaption, \Lat) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars 
-  P(\Lat)\right)\right) \le \log P(\oimage\given\icaption).
+\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\bigg(\log P(\oimage\given\icaption, \Lat) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars 
+  P(\Lat)\right)\bigg) \le \log P(\oimage\given\icaption).
 \eeqa
@@
 Similar to the DRAW model, the inference recurrent network 
-produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using an array of Gaussian filters (inverse of $\writeop$), over the training image at each step.
+produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using two array of 1D Gaussian filters (inverse of $\writeop$ from section \ref{sec:image_model}), over the training image at each step.
 Specifically,

### 2015-11-01 iclr-paper-new.tex
u

@@
 %defined in \citep{gregor_draw}). 
-The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ with smoothly varying locations and scales predicted from the generative LSTM hidden state $\hdec_t$ (same as defined in \citep{gregor_draw}).
+The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ with smoothly varying locations and scales predicted from the generative LSTM hidden state $\hdec_t$ (same as defined in \cite{gregor_draw}).
 The Gaussian filter banks apply to the generated $p$-by-$p$ image patch $\WriteFunc(\hdec_t) \in \real^{p\times p}$, placing it onto the canvas $c_t$:

### 2015-11-01 iclr-paper-new.tex
u

@@
 %defined in \citep{gregor_draw}). 
-The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ with smoothly varying locations and scales predicted from the generative LSTM hidden state $\hdec_t$ (same as defined in \cite{gregor_draw}).
+The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ with locations and scales predicted from the generative LSTM hidden state $\hdec_t$ (same as defined in \cite{gregor_draw}).
 The Gaussian filter banks apply to the generated $p$-by-$p$ image patch $\WriteFunc(\hdec_t) \in \real^{p\times p}$, placing it onto the canvas $c_t$:
@@
 where $\hat{\oimage}$ is the \textit{error} image and $\henc_0$ is initialized to the learned bias $b$. 
-Note that the inference encoder $\encoder$ 
+Note that the inference $\encoder$ 
 takes as its input both the output
-of the $\readop$ operator $r_t$, which depends on the original input image $\oimage$,
+of the $\readop$ operator $r_t\in \real^{p\times p}$, which depends on the original input image $\oimage$,
 and the previous state of the 

### 2015-11-01 iclr-paper-new.tex
u

@@
 Similar to the DRAW model, the inference recurrent network 
-produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using two array of 1D Gaussian filters (inverse of $\writeop$ from section \ref{sec:image_model}), over the training image at each step.
+produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using two arrays of 1D Gaussian filters (inverse of $\writeop$ from section \ref{sec:image_model}), over the training image at each step.
 Specifically,
@@
 \label{eq:loss}
-\loss =  &\expectation_{Q(\Lat_{1:T}\given\icaption,\oimage)}\left[ \log p(\oimage\given\icaption,\Lat_{1:T}) - \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{t-1},\icaption)\right) \right] \nonumber \\& - \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
+\loss =  &\expectation_{Q(\Lat_{1:T}\given\icaption,\oimage)}\left[ \log p(\oimage\given\icaption,\Lat_{1:T}) - \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\Lat_{1:t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\Lat_{1:t-1},\icaption)\right) \right] \nonumber \\& - \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
 \end{align}
@@
 \label{eq:loss_mc}
-\loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) - \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{t-1},\icaption)\right) \right] \nonumber \\& - \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
+\loss \approx  &\frac{1}{\numSamples}\sum_{\sampleIdx=1}^\numSamples\left[ \log p(\oimage\given\icaption,\LatSample^\sampleIdx_{1:T}) - \sum_{t=2}^T\kldiv\left(\post(\Lat_t\given\LatSample^\sampleIdx_{1:t-1},\icaption,\oimage)\klBars\prior(\Lat_t\given\LatSample^\sampleIdx_{1:t-1},\icaption)\right) \right] \nonumber \\& - \kldiv\left(\post(\Lat_1\given\oimage)\klBars\prior(\Lat_1)\right).
 \end{align}

### 2015-11-01 iclr-paper-new.tex
u
a

@@
 \label{eq:write}
-\Delta \canv_{t-1} = \canv_{t} - \canv_{t-1} = \writeop(\hdec_t) = \gFilterx(\hdec_t) \WriteFunc(\hdec_t) \gFiltery(\hdec_t)^\top. 
+\Delta \canv_{t} = \canv_{t} - \canv_{t-1} = \writeop(\hdec_t) = \gFilterx(\hdec_t) \WriteFunc(\hdec_t) \gFiltery(\hdec_t)^\top. 
 \eeqa
 
-Finally, each pixel $\canv_{T,i}$ from the final canvas $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli 
-distribution with the mean vector $\sigmoid(\canv_{T})$ over the image pixels 
-$\oimage$ given the latent variables $\Lat_{1:T}$ and the input caption $\icaption$.
+Finally, each entry $\canv_{T,i}$ from the final canvas matrix $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli 
+distribution with the mean vector $\sigmoid(\canv_{T})$ over the 
+$h\times w$ image pixels~$\oimage$ given the latent variables 
+$\Lat_{1:T}$ and the input caption $\icaption$.\footnote{We also experimented with a conditional Gaussian observation model, but
+it worked much worse compared to the Benoulli model.}
 
@@
 
-\subsection*{Acknowledgements}
+{\small
+{\bf Acknowledgements:}
 This work was supported by Samsung and IARPA, Raytheon BBN Contract No. D11PC20071.
@@
 %We will be releasing our code with pretrained models.
+}
 

### 2015-11-01 iclr-paper-new.tex
a

@@
 %A significant amount of recent work has focused on generating captions from images
-\citep{karpathy_captions}, \citep{vinyals_captions}, \citep{xu_captions}.
+\citep{karpathy_captions,vinyals_captions,xu_captions}.
 These models take an image descriptor and generate unstructured texts using a recurrent decoder.
-In contrast, we will explore models that condition in the opposite direction, i.e.
+In contrast, in this paper we explore models that condition in the opposite direction, i.e.
 taking textual descriptions as input and using them to generate relevant images.
@@
 The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \citep{denton_lapgan}.
-We then illustrate how our method, learnt on Microsoft COCO \citep{mscoco}, generalizes to captions describing novel scenarios that are not seen in the dataset (Figure~\ref{fig:genimages4}).
+We further
+illustrate how our method, learned on Microsoft COCO~\citep{mscoco}, generalizes to captions describing novel scenes that are not 
+seen in the dataset, such as ``A stop sign is flying in blue skies'' (see \Figref{fig:genimages4}).
 
@@
 
-Deep Neural Networks have achieved significant success in various tasks such as image recognition \citep{krizhevsky_imagenet} and speech transcription \citep{graves_speech}. 
+Deep Neural Networks have achieved significant success in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech}, and machine 
+translation~\citep{bahdanau_mt}. 
 While most of the recent success has been achieved by discriminative models, generative models have not yet enjoyed the same level of success.
@@
 The encoder is used to approximate a posterior distribution and the decoder is used to stochastically reconstruct the data from latent variables.
-The model's efficient inference and learning procedures allow it to scale to large datasets.
-\cite{gregor_draw} introduced the Deep Recurrent Attention Writer (DRAW), extending the VAE approach by incorporating a novel differentiable attention mechanism which significantly improved performance.% as well as quality of generated samples.
-%While most of the samples from VAE and DRAW models resemble a clear structure of objects, the generated images are blurry most of the time.
-
+%The model's efficient inference and learning procedures allow it to scale to large datasets. 
+\cite{gregor_draw} introduced the Deep Recurrent Attention Writer (DRAW), extending the VAE approach by incorporating a novel differentiable attention mechanism.
 Generative Adversarial Networks (GANs) \citep{goodfellow_gan} are another type of generative models that use noise-contrastive estimation \citep{gutmann_nce} to avoid calculating an intractable normalization constant.
 The model consists of a generator that generates samples using a uniform distribution and a discriminator that discriminates between real and generated images.
-Both networks can be seen as playing a game against each other, where the generator tries to produce samples that look real and the discriminator tries not to be fooled by the generator. 
+%Both networks can be seen as playing a game against each other, where the generator tries to produce samples that look real and the discriminator tries not to be fooled by the generator. 
 Recently, \cite{denton_lapgan} have scaled those models by training conditional GANs at each level of a Laplacian pyramid of images. 
@@
 
-While all of the previous work has been focused on unconditional models or models conditioned on labels, to the best of our knowledge this paper is the first to introduce a generative model of images conditioned on captions.
+While all of the previous work has been focused on unconditional models or models conditioned on labels, %to the best of our knowledge 
+in this paper we develop a generative model of images conditioned on captions.
 
@@
 
+{\small
 \bibliography{iclr-paper}
 \bibliographystyle{iclr-style/iclr2016_conference}
+}
 

### 2015-11-01 iclr-paper-new.tex
a

@@
 Formally, an image is generated by iteratively computing the following set of 
-equations for $t=1,...,T$ (see Figure~\ref{fig:figmodel}), with $\hdec_0$ and $c_0$
+equations for $t=1,...,T$ (see \Figref{fig:figmodel}), with $\hdec_0$ and $c_0$
 initialized to learned biases:
@@
 
-First, we wanted to see whether the model understood one of the most basic properties of any object, the color.
-In Figure~\ref{fig:genimages3}, we generated images of school buses with four different colors: yellow, red, green and blue.
+First, we explored whether the model understood one of the most basic properties of any object, the color.
+In \Figref{fig:genimages3}, we generated images of school buses with four different colors: yellow, red, green and blue.
 Although, there are images of buses with different colors in the training set, all mentioned school buses are specifically colored yellow.
@@
 
-Apart from changing the colors of objects, we were curious whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
+Apart from changing the colors of objects, we next examined whether changing the background of the scene described in a caption would result in the appropriate changes in the generated samples.
 The task of changing the background of an image is somewhat harder than just changing the color of an object because the model will have to make alterations over a wider visual area.
-Nevertheless, as shown in Figure~\ref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
+Nevertheless, as shown in \Figref{fig:genimages2} changing the skies from blue to rainy in a caption as well as changing the grass type from dry to green in another caption resulted in the appropriate changes in the generated image.
 
@@
 However, the model struggled when the visual difference between objects was very small, such as when the objects have the same general shape and color.
-In Figure~\ref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
+In \Figref{fig:genimages3}, we demonstrate that when we swap two objects that are both visually similar, for example cats and dogs, it is difficult to discriminate solely from the generated samples whether it is an image of a cat or dog, even though we might notice an animal-like shape.
 This highlights a limitation of the model in that it has difficulty modelling the fine-grained details of objects.
@@
 These captions describe a common object doing unusual things
-or set in a strange location.
-%that are impossible in the real world. 
-%considering the physics of real world. 
-Even though some of these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene.
-Nevertheless, as you can see in Figure~\ref{fig:genimages4}, the model managed to generate reasonable images.
+or set in a strange location, for example ``A toilet seat sits open in the grass field''.
+Even though some of these scenarios may never occur in real life, it is very easy for humans to imagine the corresponding scene.
+Nevertheless, as you can see in \Figref{fig:genimages4}, the model managed to generate reasonable images.
 
-\subsubsection{Analysis of Attention}
 \begin{figure}[!t]
@@
 
-After flipping sets of words in the captions, we were curious to see which words the model attended to when generating images.
+\subsubsection{Analysis of Attention}
+After flipping sets of words in the captions, we further explored which words the model attended to when generating images.
 It turned out that during the generation step, the model mostly focused on the specific words (or nearby words) that carried the main semantic meaning expressed in the sentences.
 The attention values of words in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
-For example, in Figure~\ref{fig:diffmodels} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
+For example, in \Figref{fig:diffmodels} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
 This suggests that, in their respective sentences, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
@@
 
-We compared the performance of the proposed model to the DRAW model conditioned on captions without align function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors \citep{kiros_skipthought} (skipthoughtDRAW).
+We compared the performance of the proposed model to the DRAW model conditioned on captions without $align$ function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors \citep{kiros_skipthought} (skipthoughtDRAW).
 All of the conditional DRAW models were trained with binary cross-entropy cost function.
 We also compared our model with Fully-Connected (Fully-Conn) and Convolutional-Deconvolutional (Conv-Deconv) Variational Autoencoders which were trained with the least squares cost function.
-The LAPGAN model was trained on a two level Laplacian Pyramid with a GAN as a top layer generator and all stages were conditioned on the same skipthought vector.
+The LAPGAN model of~\cite{denton_lapgan} was trained on a two level Laplacian Pyramid with a GAN as a top layer generator and all stages were conditioned on the same skipthought vector.
 
-In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
+In \Figref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
 While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with least squares cost function have a watercolor effect on the images.
@@
 As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods.
-We rank the images in the test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
-Unsurprisingly, generative models didn't perform well in the image retrieval task.
+We rank the images in the test set conditioned on the captions based on the variational lower bound of the log-probabilities
+and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
+Unsurprisingly, generative models did not perform well on the image retrieval task.
 To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
 The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.
-Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set \citep{kiros_captions}.
-We found that the lower bound of log likelihood decreased for sharpened images, and that sharpening considerably hurt the retrieval results.
-Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.
+Since there are ``easy'' images for which the model assigns high log-probabilities independent of the query caption, 
+we instead look at the ratio of 
+the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set, following the retrieval protocol of~\citep{kiros_captions}.
+We also found that the lower bound on the test log-probabilities decreased for sharpened images, and that sharpening considerably hurt the retrieval results.
+Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.
 

### 2015-11-01 iclr-paper-new.tex
a

@@
 
-\section{Model}
+\section{AlignDRAW Model}
 \label{sec:model}
@@
 In particular, captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on a canvas $c_t$ over time $t=1,...,T$.
-Our model can be viewed as a part of the sequence-to-sequence framework \citep{ilya_mt,cho_mt,nitish_video}.
+The model can be viewed as a part of the sequence-to-sequence framework \citep{ilya_mt,cho_mt,nitish_video}.
 %Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
@@
 
-The terms in the variational lower bound are rearranged using the law of total expectation. Therefore, the objective function $\loss$ is calculated as follows:
+The terms in the variational lower bound can be rearranged using the law of total expectation. Therefore, the variational bound $\loss$ is calculated as follows:
 \begin{align}
@@
 \end{align}
+The model can be trained using stochastic gradient descent. 
 In all of our experiments, we used only a single sample 
-from $\post(\Lat_{1:T}\given\icaption, \oimage)$ during training.
+from $\post(\Lat_{1:T}\given\icaption, \oimage)$ for parameter learning.
+Training details, hyperparameter settings,
+and the overall model architecture are specified in Appendix. 
+We will also release our code with pretrained models.
 

### 2015-11-01 iclr-paper-new.tex
a
u

--- 2015-11-01 iclr-paper-new.tex
@@
 \eeqa
-The corresponding alignment probabilities $\alpha_{1...N}^{t}$ are obtained using the caption representation $\hlangall$ and the hidden state of the generative model $\hdec_{t-1}$, by first calculating energies using a $\tanh$ perceptron scaled by vector $\vv$
-and then normalizing them: %, as in \citep{bahdanau_mt}:
+The corresponding alignment probability $\alpha_{k}^{t}$ for the $k^{th}$ word in the caption is obtained using the caption representation $\hlangall$ and the current hidden state of the generative model $\hdec_{t-1}$:
+%first calculating energies using a $\tanh$ perceptron scaled by vector $\vv$
+%and then normalizing them: %, as in \citep{bahdanau_mt}:
 \beqa
-\alpha_{k}^{t}=\frac{\exp(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)}{\sum_{j=1}^{N}
-\exp(\vv^{\top}\tanh(U\hlang_{j} + W\hdec_{t-1} + b)},
+%\alpha_{k}^{t}=\frac{\exp({\hlang_k}^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b))}{\sum_{j=1}^{N}
+\alpha_{k}^{t} \propto \exp\left({\hlang_k}^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)\right),
 \nonumber 
 \eeqa
-where $U \in \mathbb{R}^{d \times m}$, $W \in \mathbb{R}^{d \times n}$ and 
-$b \in \mathbb{R}^{d \times 1} $ are the learned model parameters of the alignment model.
+where $U \in \mathbb{R}^{m \times m}$, $W \in \mathbb{R}^{m \times n}$ and 
+$b \in \mathbb{R}^{m} $ are the learned model parameters of the alignment model.
 
@@
 which is added to a cumulative canvas matrix $c_t \in \real^{h \times w}$, with $c_0=0$ (\Eqref{eq:writeop}). 
-%\FIXME{Jimmy, could you put more about the write operator}. 
-%The $\writeop$ operator
-%applies an array of Gaussian filters on the patch, placing 
-%it on the canvas $c_t$ with smoothly varying location and zoom (same as 
-%defined in \citep{gregor_draw}). 
 The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ with locations and scales predicted from the generative LSTM hidden state $\hdec_t$ (same as defined in \cite{gregor_draw}).

### 2015-11-01 iclr-paper-new.tex

@@
 
-\section{Model}
+\section{The AlignDRAW Model}
 \label{sec:model}
@@
 \beqa
-\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption)\bigg(\log P(\oimage\given\icaption, \Lat) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars 
-  P(\Lat)\right)\bigg) \le \log P(\oimage\given\icaption).
+\loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption) \log P(\oimage\given\icaption, \Lat) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars 
+  P(\Lat \given\icaption)\right) \le \log P(\oimage\given\icaption).
 \eeqa
@@
 
-In \Figref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
-While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with least squares cost function have a watercolor effect on the images.
-We also found that the quality of generated samples was very similar among different variants of conditional DRAW models.
-
-As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods.
-We rank the images in the test set conditioned on the captions based on the variational lower bound of the log-probabilities
-and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
-Unsurprisingly, generative models did not perform well on the image retrieval task.
-To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
-The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.
-Since there are ``easy'' images for which the model assigns high log-probabilities independent of the query caption, 
-we instead look at the ratio of 
-the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set, following the retrieval protocol of~\citep{kiros_captions}.
-We also found that the lower bound on the test log-probabilities decreased for sharpened images, and that sharpening considerably hurt the retrieval results.
-Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.
-
-Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI) \citep{wang_ssi}, which incorporates luminance and contrast masking into the error calculation.
-Strong inter-dependencies of closer pixels are also taken into account and the metric is calculated on small windows of the images.
-Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI.
-As you can see on Table~\ref{tab:results}, SSI scores achieved by variational models were higher compared to SSI score achieved by LAPGAN.
-%As for another metric of image similarity, we calculated euclidian distance between generated images and real images in the feature space of VGG-like model trained on CIFAR dataset.
-This is not necessarily a good metric, since the statistics of generated images were vastly different from statistics of training images.
-Nevertheless, the results are reported on Table~\ref{tab:results}.
-
-% \begin{table}[!t]
-% \begin{center}
-% \begin{tabulary}{\linewidth}{c || c c c c c || c}
-% \hline
-% \multicolumn{7}{c}{\textbf{COCO (before sharpening)}} \\
-% \hline
-% & \multicolumn{5}{c||}{Image Search} & Image Similarity \\
-% %\cline{2-6}
-% \textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI} \\
-% \hline
-% \hline
-% LAPGAN & - & - & - & - & - & 0.08 \\ % - & - & - & - & - & 0.08
-% \hline
-% Fully-conn. VAE (L2 cost) & 0.7 & 4.6 & 8.7 & 46.8 & 54 & 0.122 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-% Conv. VAE (L2 cost) & 0.7 & 4.4 & 8.5 & 46.6 & 54 & \textbf{0.131} \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-% Noalign DRAW & \textbf{1.0} & 6.1 & 11.2 & 52.1 & \textbf{48} & 0.114 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-% Skipthought DRAW & \textbf{1.0} & \textbf{6.2} & \textbf{11.5} & \textbf{52.5} & \textbf{48} & 0.118 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-% Align DRAW & 0.9 & \textbf{6.2} & 11.2 & \textbf{52.5} & \textbf{48} & 0.115 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
-% \end{tabulary}
-% \end{center}
-% \end{table}
-
-\section{Discussion}
-
-In this paper, we demonstrated that the alignDRAW model, a combination of a recurrent variational autoencoder with an alignment model over words, succeeded in generating images that correspond to a given input caption.
-By extensively using attentional mechanisms, our model gained several advantages.
-Namely, the use of the visual attention mechanism allowed us to decompose the problem of image generation into a series of steps instead of a single forward pass, while the attention over words provided us an insight whenever our model failed to generate a relevant image.
-Additionally, our model generated images corresponding to captions which generalized beyond the training set, such as sentences describing novel scenarios which are highly unlikely to occur in real life.
-
-Because the alignDRAW model tends to output slightly blurry samples, likely due to underfitting on the training data, we augmented the model with a sharpening post-processing step in which GAN generated edges which were added to the alignDRAW samples.
-Unfortunately, this is not an ideal solution due to the fact that the whole model was not trained in an end-to-end fashion.
-Therefore a direction of future work would be to find methods that can bypass the separate post-processing step and output sharp images directly in an end-to-end manner.
-
 \begin{table}[!t]
@@
 
+In \Figref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
+While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with least squares cost function have a watercolor effect on the images.
+We also found that the quality of generated samples was very similar among different variants of conditional DRAW models.
+
+As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods.
+We rank the images in the test set conditioned on the captions based on the variational lower bound of the log-probabilities
+and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
+Unsurprisingly, generative models did not perform well on the image retrieval task.
+To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
+The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.
+Since there are ``easy'' images for which the model assigns high log-probabilities independent of the query caption, 
+we instead look at the ratio of 
+the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set, following the retrieval protocol of~\citep{kiros_captions}.
+We also found that the lower bound on the test log-probabilities decreased for sharpened images, and that sharpening considerably hurt the retrieval results.
+Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.
+
+Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI) \citep{wang_ssi}, which incorporates luminance and contrast masking into the error calculation.
+Strong inter-dependencies of closer pixels are also taken into account and the metric is calculated on small windows of the images.
+Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI.
+As you can see on Table~\ref{tab:results}, SSI scores achieved by variational models were higher compared to SSI score achieved by LAPGAN.
+%As for another metric of image similarity, we calculated euclidian distance between generated images and real images in the feature space of VGG-like model trained on CIFAR dataset.
+This is not necessarily a good metric, since the statistics of generated images were vastly different from statistics of training images.
+Nevertheless, the results are reported on Table~\ref{tab:results}.
+
+\section{Discussion}
+
+In this paper, we demonstrated that the alignDRAW model, a combination of a recurrent variational autoencoder with an alignment model over words, succeeded in generating images that correspond to a given input caption.
+By extensively using attentional mechanisms, our model gained several advantages.
+Namely, the use of the visual attention mechanism allowed us to decompose the problem of image generation into a series of steps instead of a single forward pass, while the attention over words provided us an insight whenever our model failed to generate a relevant image.
+Additionally, our model generated images corresponding to captions which generalized beyond the training set, such as sentences describing novel scenarios which are highly unlikely to occur in real life.
+
+Because the alignDRAW model tends to output slightly blurry samples, likely due to underfitting on the training data, we augmented the model with a sharpening post-processing step in which GAN generated edges which were added to the alignDRAW samples.
+Unfortunately, this is not an ideal solution due to the fact that the whole model was not trained in an end-to-end fashion.
+Therefore a direction of future work would be to find methods that can bypass the separate post-processing step and output sharp images directly in an end-to-end manner.
+
 {\small

### 2015-11-01 iclr-paper-new.tex
u
a

--- 2015-11-01 iclr-paper-new.tex
@@
 
-\section{AlignDRAW Model}
+\section{The AlignDRAW Model}
 \label{sec:model}
@@
 \end{center}
-\caption{The diagram shows our model learns to draw images by learning an alignment of input captions and generating canvas. The details are explained in Section \ref{sec:model}.}
+\caption{AlignDRAW model for generating images by learning an alignment between the input captions
+and generating canvas. The caption is encoded using 
+the Bidirectional RNN (left). The generative RNN takes a latent sequence $z_{1:T}$ sampled from the 
+prior along with dynamic caption representation $s_{1:T}$ to generate the canvas matrix $c_T$, which is then
+used to generate the final image $\oimage$ (right).
+The inference RNN is used to compute approximate posterior $Q$ over the latent sequence.} 
 \label{fig:figmodel}

### 2015-11-01 iclr-paper-new.tex

--- 2015-11-01 iclr-paper-new.tex
@@
 \beqa
+\label{eq:varbound}
 \loss = \sum_{\Lat}Q(\Lat\given\oimage,\icaption) \log P(\oimage\given\icaption, \Lat) - \kldiv\left(Q(\Lat\given\oimage,\icaption)\klBars 
@@
 
-The terms in the variational lower bound can be rearranged using the law of total expectation. Therefore, the variational bound $\loss$ is calculated as follows:
+The terms in the variational lower bound Eq. \ref{eq:varbound} can be rearranged using the law of total expectation. Therefore, the variational bound $\loss$ is calculated as follows:
 \begin{align}

### 2015-11-01 iclr-paper-new.tex
u

--- 2015-11-01 iclr-paper-new.tex
@@
 the $\decoder$ takes the previous hidden state $\hdec_{t-1}$ and
-combines it with the input from both the latent sample $z_t$ and 
-sentence representation $s_t$, where
-$[v_1, v_2]$ denotes the concatenation of $v_1$ and $v_2$ into the single vector.
+the concatenation of the latent sample $z_t$ with sentence representation $s_t$ 
+as the input to the LSTM network.
+%combines it with the input from both the latent sample $z_t$ and 
+%sentence representation $s_t$, where
+%$[v_1, v_2]$ denotes the concatenation of $v_1$ and $v_2$ into the single vector.
 

### 2015-11-01 iclr-paper-new.tex
fixed some notation

@@
 the $\decoder$ takes the previous hidden state $\hdec_{t-1}$ and
-the concatenation of the latent sample $z_t$ with sentence representation $s_t$ 
-as the input to the LSTM network.
+combines it with the input from both the latent sample $z_t$ and the sentence representation $s_t$.
 %combines it with the input from both the latent sample $z_t$ and 

### 2015-11-01 iclr-paper-new.tex
fixed the merge error

@@
 
-We also tried to analyze the way the model generated images.
-Unfortunately, we found that there was no significant connection between the patches drawn on canvas and the most attended words at particular timesteps.
-
-\subsubsection{Comparison With Other Models}
-
-Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffers from its own specific drawbacks.
-Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
-To get a better comparison between performances of different generative models, we report results on two different metrics as well as a qualitative comparison of different generative models.
 \begin{figure}[!t]

### 2015-11-01 iclr-paper-new.tex
fix the wording for write

@@
 which is added to a cumulative canvas matrix $c_t \in \real^{h \times w}$, with $c_0=0$ (\Eqref{eq:writeop}). 
-The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ with locations and scales predicted from the generative LSTM hidden state $\hdec_t$ (same as defined in \cite{gregor_draw}).
+The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ whose filter locations and scales are computed from the generative LSTM hidden state $\hdec_t$ (same as defined in \cite{gregor_draw}).
 The Gaussian filter banks apply to the generated $p$-by-$p$ image patch $\WriteFunc(\hdec_t) \in \real^{p\times p}$, placing it onto the canvas $c_t$:

### 2015-11-01 iclr-paper-new.tex
a

--- 2015-11-01 iclr-paper-new.tex
@@
 \end{center}
-\caption{Examples of novel scene compositions.}
+\caption{\small Examples of novel scene compositions.}
 \label{fig:genimages4}
@@
 \end{center}
-\caption{AlignDRAW model for generating images by learning an alignment between the input captions
+\caption{\small AlignDRAW model for generating images by learning an alignment between the input captions
 and generating canvas. The caption is encoded using 
@@
 the $\decoder$ takes the previous hidden state $\hdec_{t-1}$ and
-combines it with the input from both the latent sample $z_t$ and the sentence representation $s_t$.
-%combines it with the input from both the latent sample $z_t$ and 
-%sentence representation $s_t$, where
-%$[v_1, v_2]$ denotes the concatenation of $v_1$ and $v_2$ into the single vector.
+combines it with the input from both the latent sample $z_t$ and 
+sentence representation $s_t$, where
+$[v_1, v_2]$ denotes the concatenation of $v_1$ and $v_2$ into the single vector.
 
@@
 \end{center}
-\caption{\textbf{Bottom}: Examples of changing the object while keeping the caption fixed. \textbf{Top}: Examples of changing the color while keeping the caption fixed.}
+\caption{\small \textbf{Bottom}: Examples of changing the object while keeping the caption fixed. \textbf{Top}: Examples of changing the color while keeping the caption fixed.}
 \label{fig:genimages3}
@@
 \end{center}
-\caption{Results of different models.}
+\caption{Retrieval results of different models. \bf{ R$\at$K} is Recall@K (high is good). Med r is the median rank (low is good).
+Best results overall are bold while best results without OxfordNet features are underlined}
 \label{tab:results}

### 2015-11-01 iclr-paper-new.tex
a

@@
 \end{center}
-\caption{\small Examples of novel scene compositions.}
+\caption{\small Examples of novel scene compositions. Best viewed in colour.}
 \label{fig:genimages4}
@@
 The output of the $\decoder$ function $\hdec_t$ is then passed through the $\writeop$ operator 
-which is added to a cumulative canvas matrix $c_t \in \real^{h \times w}$, with $c_0=0$ (\Eqref{eq:writeop}). 
+which is added to a cumulative canvas matrix $c_t \in \real^{h \times w}$ (\Eqref{eq:writeop}). 
 The $\writeop$ operator produces two arrays of 1D Gaussian filter banks $\gFilterx(\hdec_t)\in\real^{h\times p}$ and $\gFiltery(\hdec_t)\in\real^{w\times p}$ whose filter locations and scales are computed from the generative LSTM hidden state $\hdec_t$ (same as defined in \cite{gregor_draw}).
-The Gaussian filter banks apply to the generated $p$-by-$p$ image patch $\WriteFunc(\hdec_t) \in \real^{p\times p}$, placing it onto the canvas $c_t$:
+The Gaussian filter banks are then applied to the generated $p$-by-$p$ image patch $\WriteFunc(\hdec_t) \in \real^{p\times p}$, placing it onto the canvas:
 \beqa
@@
 \end{center}
-\caption{\small \textbf{Bottom}: Examples of changing the object while keeping the caption fixed. \textbf{Top}: Examples of changing the color while keeping the caption fixed.}
+\caption{\small
+ \textbf{Top}: Examples of changing the color while keeping the caption fixed.
+ \textbf{Bottom}: Examples of changing the object while keeping the caption fixed. }
 \label{fig:genimages3}
@@
 \end{center}
-\caption{Retrieval results of different models. \bf{ R$\at$K} is Recall@K (high is good). Med r is the median rank (low is good).
-Best results overall are bold while best results without OxfordNet features are underlined}
+\caption{Retrieval results of different models. 
+\textbf{R@K} is Recall@K
+    (high is good). \textbf{Med} {\it r} is the median rank (low is
+    good). 
+}
 \label{tab:results}

### 2015-11-01 iclr-paper-new.tex
a

@@
 \section{Introduction}
-%\vspace{-0.2cm}
+\vspace{-0.1in}
 Statistical natural image modelling remains a fundamental problem in computer vision and image understanding. 
@@
 
-\section{Related Work}
-
 \begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
+\vspace{-0.3in}
 \begin{center}
@@
 \label{fig:genimages4}
-\vspace{-0.3cm}
+\vspace{-0.2in}
 \end{figure}
 
+\section{Related Work}
+\vspace{-0.1in}
 Deep Neural Networks have achieved significant success in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech}, and machine 
@@
 \section{The AlignDRAW Model}
+\vspace{-0.1in}
 \label{sec:model}
@@
 \subsection{Language Model: the Bidirectional Attention RNN}
+\vspace{-0.05in}
 \label{sec:lang}
@@
 \begin{figure}[t!]
+\vspace{-0.3in}
 \captionsetup[subfigure]{labelformat=empty}
@@
 \end{center}
+\vspace{-0.1in}
 \caption{\small AlignDRAW model for generating images by learning an alignment between the input captions
@@
 \label{fig:figmodel}
-\vspace{-0.3cm}
+\vspace{-0.2in}
 \end{figure}
@@
 \subsection{Image Model: the Conditional DRAW Network}
+\vspace{-0.05in}
 \label{sec:image_model}
@@
 $\Lat_{1:T}$ and the input caption $\icaption$.\footnote{We also experimented with a conditional Gaussian observation model, but
-it worked much worse compared to the Benoulli model.}
+it worked worse compared to the Benoulli model.}
 
 \subsection{Learning}
+\vspace{-0.05in}
 \label{sec:learning}
@@
 \subsection{Generating Images from Captions}
-
+\vspace{-0.05in}
 During the image generation step, we discard the inference network and instead sample from the prior distribution. 
@@
 
-\section{Experiments}
-\subsection{Microsoft COCO}
-
 \begin{figure}[!t]
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.3cm}
+\vspace{-0.2in}
 \begin{center}
@@
 \label{fig:genimages3}
-%\vspace{-0.3cm}
+\vspace{-0.1in}
 \end{figure}
 
+\section{Experiments}
+\vspace{-0.05in}
+\subsection{Microsoft COCO}
+\vspace{-0.05in}
 Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
@@
 \subsubsection{Analysis of Generated Images}
+\vspace{-0.05in}
 The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
@@
 \captionsetup[subfigure]{labelformat=empty}
-%\vspace{-0.5cm}
+\vspace{-0.2in}
 \begin{center}
@@
 \label{fig:genimages2}
-%\vspace{-0.3cm}
+\vspace{-0.1in}
 \end{figure}
@@
 \subsubsection{Analysis of Attention}
+\vspace{-0.05in}
 After flipping sets of words in the captions, we further explored which words the model attended to when generating images.
@@
 The attention values of words in sentences helped us interpret the reasons why the model made the changes it did when we flipped certain words.
-For example, in \Figref{fig:diffmodels} we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
+For example, in \Figref{fig:diffmodels}, top row, we can see that when we flipped the word ``desert'' to ``forest'', the attention over words in the sentence did not change drastically.
 This suggests that, in their respective sentences, the model looked at ``desert'' and ``forest'' with relatively equal probability, and thus made the correct changes.
@@
 \captionsetup[subfigure]{labelformat=empty}
-%\vspace{-0.3cm}
+\vspace{-0.2in}
 \begin{center}
@@
 \end{center}
-\caption{\textbf{Bottom}: Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.} \textbf{Top}: Examples of most attended words while changing the background in the caption.}
+\caption{ \textbf{Top}: Examples of most attended words while changing the background in the caption. 
+\textbf{Bottom}: Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
 \label{fig:diffmodels}
-%\vspace{-0.3cm}
+\vspace{-0.1in}
 \end{figure}
@@
 \subsubsection{Comparison With Other Models}
+\vspace{-0.05in}
 Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffers from its own specific drawbacks.
@@
 \label{tab:results}
-%\vspace{-0.3cm}
+\vspace{-0.1in}
 \end{table}
 
-In \Figref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
+In \Figref{fig:diffmodels}, bottom row, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards".
 While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with least squares cost function have a watercolor effect on the images.
@@
 \section{Discussion}
-
+\vspace{-0.1in}
 In this paper, we demonstrated that the alignDRAW model, a combination of a recurrent variational autoencoder with an alignment model over words, succeeded in generating images that correspond to a given input caption.

### 2015-11-01 iclr-paper-new.tex
a

@@
 The images were resized to $32 \times 32$ dimension for consistency with other tiny image datasets \citep{krizhevsky_cifar}.
-In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
+In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more
+generated images, visit \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
 
@@
 
-We compared the performance of the proposed model to the DRAW model conditioned on captions without $align$ function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors \citep{kiros_skipthought} (skipthoughtDRAW).
+We compared the performance of the proposed model to the DRAW model conditioned on captions without $align$ function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors of \cite{kiros_skipthought} (skipthoughtDRAW).
 All of the conditional DRAW models were trained with binary cross-entropy cost function.
@@
 \hline
-\multicolumn{7}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
+\multicolumn{7}{c}{\textbf{Microsoft COCO (prior to sharpening)}} \\
 \hline
-& \multicolumn{5}{c||}{Image Search} & \multicolumn{1}{c}{Image Similarity} \\
+& \multicolumn{5}{c||}{Image Retrieval} & \multicolumn{1}{c}{Image Similarity} \\
 %\cline{2-6}
@@
 To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
-The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.
+The shortlist of test images used for evaluation can be downloaded from \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.
 Since there are ``easy'' images for which the model assigns high log-probabilities independent of the query caption, 

### 2015-11-01 supp.tex
a

--- 2015-10-30 supp.tex
@@
-\section{Training Details}
+\section*{Appendix A: MNIST With Captions}
+
+\begin{figure}[!t]
+\captionsetup[subfigure]{labelformat=empty}
+\begin{center}
+\includegraphics[width=0.99\textwidth]{figures/new/mnist/test3.pdf}\quad
+%
+\end{center}
+\caption{Examples of generating $60 \times 60$ MNIST images corresponding to respective captions.
+The captions on the \textbf{left column} were part of the training set.
+The digits described in the captions on the \textbf{right column} were hidden during training for the respective configurations.}
+\label{fig:figmnist}
+\vspace{-0.3cm}
+\end{figure}
+
+As an additional experiment, we trained our model on the MNIST dataset with artificial captions.
+Either one or two digits from the MNIST training dataset were placed on a $60 \times 60$ blank image.
+One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image.
+Two digits were either placed horizontally or vertically in nonoverlapping fashion.
+The corresponding artificial captions specified the
+identity of each digit along with their relative positions, e.g. ``The digit two is on the top  
+of the digit zero'', or ``The digit seven is at the bottom left of the image''.
+
+The generated images together with the attention alignments are displayed in Figure~\ref{fig:figmnist}.
+The model correctly displayed the specified digits at the described positions and even managed to generalize
+reasonably well to the configurations that were never present during training.
+In the case of generating two digits, the model would dynamically attend 
+to the digit in the caption it was drawing at that particular time-step. 
+Similarly, in the setting where the caption specified only a single digit, the model would correctly 
+attend to the digit in the caption during the whole generation process. 
+In both cases, the model placed small attention values on the words describing the position of digits in the images.
+
+\section*{Appendix B: Training Details}
 \label{sec:training_details}
 
-\subsection{Hyperparameters}
+\subsection*{Hyperparameters}
 
-Each parameter in alignDRAW was initialized by sampling from the Gaussian distribution of mean $0$ and standard deviation of $0.01$.
+Each parameter in alignDRAW was initialized by sampling from the Gaussian distribution with 
+mean $0$ and standard deviation $0.01$.
 The model was trained using \textit{RMSprop} with initial learning rate of $0.001$.
@@
 
-\subsection{Evaluation}
+\subsection*{Evaluation}
 
-On Table~\ref{tab:nll}, you can see that the alignDRAW model didn't overfit on the training data as well as managed to achieve a better lower bound of log likelihood on the test data.
-Also notice that, the results worsened dramatically after sharpening test images.
+Table~\ref{tab:nll} shows the estimated variational lower bound on the average train/validation/test 
+log-probabilities.
+Note that the alignDRAW model does not suffer much from overfitting. 
+The results worsened dramatically after sharpening test images.
 
@@
 \end{tabulary}
-\caption{The lower bound of log likelihood of conditional DRAW models on the Microsoft COCO.}
+\caption{The lower bound on the average test log-probabilities of conditional DRAW models, trained on the Microsoft COCO dataset.}
 \label{tab:nll}

### 2015-11-01 iclr-paper-new.tex
some changes

@@
 \cite{gregor_draw} introduced the Deep Recurrent Attention Writer (DRAW), extending the VAE approach by incorporating a novel differentiable attention mechanism.
+
 Generative Adversarial Networks (GANs) \citep{goodfellow_gan} are another type of generative models that use noise-contrastive estimation \citep{gutmann_nce} to avoid calculating an intractable normalization constant.
@@
 the hidden state of the generative LSTM. 
-Similar to \cite{bachman_sdm}, we have observed that 
+Similar to \citep{bachman_sdm}, we have observed that 
 the model performance is improved by including dependencies between latent variables.
@@
 %\alpha_{k}^{t}=\frac{\exp({\hlang_k}^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b))}{\sum_{j=1}^{N}
-\alpha_{k}^{t} \propto \exp\left({\hlang_k}^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)\right),
+\alpha_{k}^{t} \propto \exp\left(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)\right),
 \nonumber 
 \eeqa
-where $U \in \mathbb{R}^{m \times m}$, $W \in \mathbb{R}^{m \times n}$ and 
-$b \in \mathbb{R}^{m} $ are the learned model parameters of the alignment model.
+where $\vv \in \mathbb{R}^{l}$, $U \in \mathbb{R}^{l \times m}$, $W \in \mathbb{R}^{l \times n}$ and 
+$b \in \mathbb{R}^{l} $ are the learned model parameters of the alignment model.
 
@@
 $\Lat_{1:T}$ and the input caption $\icaption$.\footnote{We also experimented with a conditional Gaussian observation model, but
-it worked worse compared to the Benoulli model.}
+it worked worse compared to the Bernoulli model.}
 
@@
 \vspace{-0.05in}
-Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions.
+Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with at least 5 captions.
 The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model
@@
 
-We compared the performance of the proposed model to the DRAW model conditioned on captions without $align$ function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors of \cite{kiros_skipthought} (skipthoughtDRAW).
+We compared the performance of the proposed model to the DRAW model conditioned on captions without $align$ function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors of \citep{kiros_skipthought} (skipthoughtDRAW).
 All of the conditional DRAW models were trained with binary cross-entropy cost function.
 We also compared our model with Fully-Connected (Fully-Conn) and Convolutional-Deconvolutional (Conv-Deconv) Variational Autoencoders which were trained with the least squares cost function.
-The LAPGAN model of~\cite{denton_lapgan} was trained on a two level Laplacian Pyramid with a GAN as a top layer generator and all stages were conditioned on the same skipthought vector.
-
-\begin{table}[!t]
-\begin{center}
-\begin{tabulary}{\linewidth}{c || c c c c c || c c}
-\hline
-\multicolumn{7}{c}{\textbf{Microsoft COCO (prior to sharpening)}} \\
-\hline
-& \multicolumn{5}{c||}{Image Retrieval} & \multicolumn{1}{c}{Image Similarity} \\
-%\cline{2-6}
-\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
-\hline
-\hline
-LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
-\hline
-Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
-\end{tabulary}
-\end{center}
-\caption{Retrieval results of different models. 
-\textbf{R@K} is Recall@K
-    (high is good). \textbf{Med} {\it r} is the median rank (low is
-    good). 
-}
-\label{tab:results}
-\vspace{-0.1in}
-\end{table}
+The LAPGAN model of~\citep{denton_lapgan} was trained on a two level Laplacian Pyramid with a GAN as a top layer generator and all stages were conditioned on the same skipthought vector.
 
@@
 \vspace{-0.1in}
+\begin{table}[!t]
+\begin{center}
+\begin{tabulary}{\linewidth}{c || c c c c c || c c}
+\hline
+\multicolumn{7}{c}{\textbf{Microsoft COCO (prior to sharpening)}} \\
+\hline
+& \multicolumn{5}{c||}{Image Retrieval} & \multicolumn{1}{c}{Image Similarity} \\
+%\cline{2-6}
+\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
+\hline
+\hline
+LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
+\hline
+Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+\end{tabulary}
+\end{center}
+\caption{Retrieval results of different models. 
+\textbf{R@K} is Recall@K
+    (high is good). \textbf{Med} {\it r} is the median rank (low is
+    good). \textbf{SSI} is Structural Similarity Index, which is between $-1$ and $1$ (high is good).
+}
+\label{tab:results}
+\vspace{-0.1in}
+\end{table}
 In this paper, we demonstrated that the alignDRAW model, a combination of a recurrent variational autoencoder with an alignment model over words, succeeded in generating images that correspond to a given input caption.

### 2015-11-01 supp.tex
some changes

@@
 The norm of the gradients was clipped to $10.0$ during training to deal with exploding gradients.
-For all tasks, $\overrightarrow{h}^{lang}_{n}$ and $\overleftarrow{h}^{lang}_{n}$ in the language model had $128$ units, and vector $v$ in \textit{align} perceptron had $512$ units.
+We used the vocabulary size of $K = 25323$ and $K = 22$ for Microsoft COCO and MNIST with Captions respectively.
+All capital letters in the words were changed to small letters to reduce the long tail.
+For all tasks, $\overrightarrow{h}^{lang}_{n}$ and $\overleftarrow{h}^{lang}_{n}$ in the language model had $128$ units.
+The parameters in \textit{align} operator had dimensionality of $l = 512$, $\vv \in \mathbb{R}^{512}$, $U \in \mathbb{R}^{512 \times 128}$, $W \in \mathbb{R}^{512 \times n^{gen}}$ and $b \in \mathbb{R}^{512}$.
 For the Microsoft COCO task, the alignDRAW model was trained for $18$ epochs and the learning rate was reduced by the factor of $0.1$ after $11$ epochs.
@@
 
-\begin{table}[!h]
+\begin{table}[!t]
 \begin{center}

### 2015-11-02 supp.tex

@@
 \end{table}
+
+\newpage
+\section*{Appendix C: Effect of Sharpening Images.}
+\label{sec:post_processing}
+
+\begin{figure}[!h]
+\captionsetup[subfigure]{labelformat=empty}
+\vspace{-0.27in}
+\begin{center}
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-blurry.png}}\\
+\vspace{-0.45cm}
+%
+\subfloat[%A stop sign is flying in blue skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
+%
+\subfloat[%A herd of elephants flying in the blue skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
+%
+\subfloat[%A toilet seat sits open in the grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
+%
+\subfloat[%A person skiing on sand clad vast desert.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\\
+%
+
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-blurry.png}}\\
+\vspace{-0.45cm}
+%
+\subfloat[%A yellow school bus parked in a parking lot.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[%A red school bus parked in a parking lot.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[%A green school bus parked in a parking lot.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+%
+\subfloat[%A blue school bus parked in a parking lot.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\\
+%
+
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-blurry.png}}\\
+\vspace{-0.45cm}
+%
+\subfloat[%The decadent chocolate desert is on the table.
+]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
+%
+\subfloat[%A bowl of bananas is on the table.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
+%
+\subfloat[%A vintage photo of a cat.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
+%
+\subfloat[%A vintage photo of a dog.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\\
+%
+
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-blurry.png}}\\
+\vspace{-0.45cm}
+%
+\subfloat[%A very large commercial plane flying in blue skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
+%
+\subfloat[%A very large commercial plane flying in rainy skies.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
+%
+\subfloat[%A herd of elephants walking across a dry grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
+%
+\subfloat[%A herd of elephants walking across a green grass field.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\\
+%
+
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-blurry.png}}\quad
+%
+\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-blurry.png}}\\
+\vspace{-0.45cm}
+%
+\subfloat[%A rider on a blue motorcycle in the desert.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
+%
+\subfloat[%A rider on a blue motorcycle in the forest.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
+%
+\subfloat[%A surfer, a woman, and a child walk on the beach.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
+%
+\subfloat[%A surfer, a woman, and a child walk on the sun.
+]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\\
+\end{center}
+%
+%\caption{\small Effect of sharpening generated images shown in the publication \textbf{Bottom}: Sharpened Images. \textbf{Top}: Original Images}
+%\vspace{-0.1in}
+\end{figure}

### 2015-11-03 iclr-paper-new.tex
a
changes

@@
 of 1-of-K encoded words 
-${\bf y} = \{y_{1}, y_{2}, ..., y_{N}\}$, 
+${\bf y} = (y_{1}, y_{2}, ..., y_{N})$, 
 where $K$ is the size of the vocabulary and $N$ is the length of the sequence.
@@
 %\alpha_{k}^{t}=\frac{\exp({\hlang_k}^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b))}{\sum_{j=1}^{N}
-\alpha_{k}^{t} \propto \exp\left(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)\right),
+\alpha_{k}^{t} = \frac{\exp\left(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)\right)}
+  {\sum_{i=1}^N \exp\left(\vv^{\top}\tanh(U\hlang_{i} + W\hdec_{t-1} + b)\right)},
 \nonumber 
@@
 Training details, hyperparameter settings,
-and the overall model architecture are specified in Appendix. 
+and the overall model architecture are specified in Appendix~\ref{sec:training_details}. 
 We will also release our code with pretrained models.
@@
 very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
-The images were resized to $32 \times 32$ dimension for consistency with other tiny image datasets \citep{krizhevsky_cifar}.
+The images were resized to $32 \times 32$ pixels for consistency with other tiny image datasets \citep{krizhevsky_cifar}.
 In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more
@@
 \vspace{-0.05in}
-The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc.
-and then use that knowledge to generate relevant images.
+The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc., and then use that knowledge to generate relevant images.
 To examine the understanding of our model, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
@@
 
-We compared the performance of the proposed model to the DRAW model conditioned on captions without $align$ function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors of \citep{kiros_skipthought} (skipthoughtDRAW).
-All of the conditional DRAW models were trained with binary cross-entropy cost function.
+We compared the performance of the proposed model to the DRAW model conditioned on captions without the $align$ function (noalignDRAW) as well as the DRAW model conditioned on the skipthought vectors of \citep{kiros_skipthought} (skipthoughtDRAW).
+All of the conditional DRAW models were trained with a binary cross-entropy cost function, i.e.
+they had Bernoulli conditional likelihoods.
 We also compared our model with Fully-Connected (Fully-Conn) and Convolutional-Deconvolutional (Conv-Deconv) Variational Autoencoders which were trained with the least squares cost function.
@@
 and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.).
-Unsurprisingly, generative models did not perform well on the image retrieval task.
+Perhaps unsurprisingly, generative models did not perform well on the image retrieval task.
 To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.

### 2015-11-03 supp.tex
changes

@@
 
-Each parameter in alignDRAW was initialized by sampling from the Gaussian distribution with 
+Each parameter in alignDRAW was initialized by sampling from a Gaussian distribution with 
 mean $0$ and standard deviation $0.01$.
-The model was trained using \textit{RMSprop} with initial learning rate of $0.001$.
-The norm of the gradients was clipped to $10.0$ during training to deal with exploding gradients.
-We used the vocabulary size of $K = 25323$ and $K = 22$ for Microsoft COCO and MNIST with Captions respectively.
-All capital letters in the words were changed to small letters to reduce the long tail.
+The model was trained using \textit{RMSprop} with an initial learning rate of $0.001$.
+The norm of the gradients was clipped at a maximum of $10.0$ during training to deal with exploding gradients.
+We used a vocabulary size of $K = 25323$ and $K = 22$ for Microsoft COCO and MNIST with Captions respectively.
+All capital letters in the words were converted to small letters as a preprocessing step.
 For all tasks, $\overrightarrow{h}^{lang}_{n}$ and $\overleftarrow{h}^{lang}_{n}$ in the language model had $128$ units.
-The parameters in \textit{align} operator had dimensionality of $l = 512$, $\vv \in \mathbb{R}^{512}$, $U \in \mathbb{R}^{512 \times 128}$, $W \in \mathbb{R}^{512 \times n^{gen}}$ and $b \in \mathbb{R}^{512}$.
+The parameters in the \textit{align} operator had a dimensionality of $l = 512$, $\vv \in \mathbb{R}^{512}$, $U \in \mathbb{R}^{512 \times 128}$, $W \in \mathbb{R}^{512 \times n^{gen}}$ and $b \in \mathbb{R}^{512}$.
 For the Microsoft COCO task, the alignDRAW model was trained for $18$ epochs and the learning rate was reduced by the factor of $0.1$ after $11$ epochs.
@@
 
-The GAN model used for sharpening had the same configuration as the $28 \times 28$ model trained by \cite{denton_lapgan} on the edge residuals of CIFAR dataset.
+The GAN model used for sharpening had the same configuration as the $28 \times 28$ model trained by \cite{denton_lapgan} on the edge residuals of the CIFAR dataset.
 The configuration can be found at \url{https://gist.github.com/soumith/e3f722173ea16c1ea0d9}.
 The model was trained for $6$ epochs.
+For the LAPGAN model, we used a data augmentation step where we took 5 crops (top-left, top-right, bottom-left, bottom-right, center) of a downsampled COCO image and assigned each crop one of the 5 captions that described the uncropped image.
 

### 2015-11-03 iclr-paper-new.tex
spelling

@@
 Finally, each entry $\canv_{T,i}$ from the final canvas matrix $\canv_T$ is transformed using a sigmoid function $\sigmoid$ to produce a conditional Bernoulli 
-distribution with the mean vector $\sigmoid(\canv_{T})$ over the 
+distribution with mean vector $\sigmoid(\canv_{T})$ over the 
 $h\times w$ image pixels~$\oimage$ given the latent variables 

### 2015-11-03 supp.tex
fixed appendix
small changes

@@
 The model was trained for $6$ epochs.
-For the LAPGAN model, we used a data augmentation step where we took 5 crops (top-left, top-right, bottom-left, bottom-right, center) of a downsampled COCO image and assigned each crop one of the 5 captions that described the uncropped image.
 
@@
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.27in}
+\vspace{-0.15in}
 \begin{center}
@@
 ]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\\
+\vspace{-0.25cm}
 %
-
 \subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
@@
 ]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\\
+\vspace{-0.25cm}
 %
-
 \subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-blurry.png}}\quad
@@
 ]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\\
+\vspace{-0.25cm}
 %
-
 \subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-blurry.png}}\quad
@@
 ]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\\
+\vspace{-0.25cm}
 %
-
 \subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-blurry.png}}\quad
@@
 %
-%\caption{\small Effect of sharpening generated images shown in the publication \textbf{Bottom}: Sharpened Images. \textbf{Top}: Original Images}
+\caption{\small Effect of sharpening images previously shown in the publication.}
 %\vspace{-0.1in}

### 2015-11-04 iclr-paper-new.tex
small changes
changes based on michaels suggestions

@@
 \newcommand{\hlang}{h^{lang}}
-\newcommand{\hlangall}{\boldsymbol{h^{lang}}}
+%\newcommand{\hlangall}{\boldsymbol{h^{lang}}}
+\newcommand{\hlangall}{h^{lang}}
 \newcommand{\hdec}{h^{gen}}
@@
 \newcommand{\lat}{z}
-\newcommand{\vv}{{\bf v}}
+%\newcommand{\vv}{{\bf v}}
+\newcommand{\vv}{v}
 \newcommand{\Lat}{Z}
@@
 
-\section{The AlignDRAW Model}
+\section{Model}
 \vspace{-0.1in}
@@
 \vspace{-0.1in}
-\caption{\small AlignDRAW model for generating images by learning an alignment between the input captions
+\caption{\small alignDRAW model for generating images by learning an alignment between the input captions
 and generating canvas. The caption is encoded using 
@@
 \canv_t &= \canv_{t-1} + \writeop(\hdec_t), \\
-\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \textrm{Bern}(\sigmoid(\canv_{T,i})). 
+\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \sigmoid(\canv_{T}). 
+%\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \textrm{Bern}(\sigmoid(\canv_{T,i})). 
 \label{eq:write}
@@
 Training details, hyperparameter settings,
-and the overall model architecture are specified in Appendix~\ref{sec:training_details}. 
+and the overall model architecture are specified in Appendix. 
 We will also release our code with pretrained models.

### 2015-11-05 supp.tex

@@
 For all tasks, $\overrightarrow{h}^{lang}_{n}$ and $\overleftarrow{h}^{lang}_{n}$ in the language model had $128$ units.
-The parameters in the \textit{align} operator had a dimensionality of $l = 512$, $\vv \in \mathbb{R}^{512}$, $U \in \mathbb{R}^{512 \times 128}$, $W \in \mathbb{R}^{512 \times n^{gen}}$ and $b \in \mathbb{R}^{512}$.
+The parameters in the \textit{align} operator had a dimensionality of $l = 512$, $\vv \in \mathbb{R}^{512}$, $U \in \mathbb{R}^{512 \times 256}$, $W \in \mathbb{R}^{512 \times n^{gen}}$ and $b \in \mathbb{R}^{512}$.
 For the Microsoft COCO task, the alignDRAW model was trained for $18$ epochs and the learning rate was reduced by the factor of $0.1$ after $11$ epochs.

### 2015-11-05 iclr-paper-new.tex
a

@@
 \newcommand{\sampleIdx}{l}
-\newcommand{\LatSample}{\tilde{Z}}
+\newcommand{\LatSample}{\tilde{z}}
 \newcommand{\icaption}{{\bf{y}}}
@@
 \end{center}
-\caption{\small Examples of novel scene compositions. Best viewed in colour.}
+\caption{\small 
+%Examples of novel scene compositions. Best viewed in colour.
+Examples of generated images based on captions that describe 
+novel scene compositions that are highly unlikely to occur in real life. 
+The captions describe a common object doing unusual things or set in a strange location. 
+}
 \label{fig:genimages4}
@@
 \canv_t &= \canv_{t-1} + \writeop(\hdec_t), \\
-\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \sigmoid(\canv_{T}). 
-%\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \textrm{Bern}(\sigmoid(\canv_{T,i})). 
+%\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \sigmoid(\canv_{T}). 
+\tilde{\oimage} &\sim P(\oimage\given\icaption, \Lat_{1:T}) = \prod_i P(x_i\given\icaption, \Lat_{1:T}) = \prod_i \textrm{Bern}(\sigmoid(\canv_{T,i})). 
 \label{eq:write}
@@
 \alpha_{1}^{t}\hlang_{1} + \alpha_{2}^{t}\hlang_{2} + ... + \alpha_{N}^{t}\hlang_{N}.
-\nonumber 
 \eeqa
@@
 \beqa
-%\alpha_{k}^{t}=\frac{\exp({\hlang_k}^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b))}{\sum_{j=1}^{N}
+\label{eq:align2}
 \alpha_{k}^{t} = \frac{\exp\left(\vv^{\top}\tanh(U\hlang_{k} + W\hdec_{t-1} + b)\right)}
   {\sum_{i=1}^N \exp\left(\vv^{\top}\tanh(U\hlang_{i} + W\hdec_{t-1} + b)\right)},
-\nonumber 
 \eeqa
@@
 $h\times w$ image pixels~$\oimage$ given the latent variables 
-$\Lat_{1:T}$ and the input caption $\icaption$.\footnote{We also experimented with a conditional Gaussian observation model, but
-it worked worse compared to the Bernoulli model.}
+$\Lat_{1:T}$ and the input caption $\icaption$\footnote{We also experimented with a conditional Gaussian observation model, but
+it worked worse compared to the Bernoulli model.}. In practice, when generating an 
+image $\oimage$, instead of sampling from the conditional Bernoulli
+distribution, we simply use the conditional mean $\oimage = \sigmoid(\canv_{T})$.
 
@@
 \vspace{-0.05in}
-During the image generation step, we discard the inference network and instead sample from the prior distribution. 
+During the image generation step, 
+we discard the inference network and instead sample from the prior distribution. 
 Due to the blurriness of samples generated by the DRAW model, we perform an additional post processing step where we use an 
@@
  \textbf{Top}: Examples of changing the color while keeping the caption fixed.
- \textbf{Bottom}: Examples of changing the object while keeping the caption fixed. }
+ \textbf{Bottom}: Examples of changing the object while keeping the caption fixed.
+ The shown images are the probabilities $\sigmoid(\canv_{T})$. Best viewed in colour.}
 \label{fig:genimages3}
@@
 Perhaps unsurprisingly, generative models did not perform well on the image retrieval task.
-To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
+To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest
+Euclidean distance in the top convolutional feature-space of a VGG-like model~\citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
 The shortlist of test images used for evaluation can be downloaded from \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.
@@
 \textbf{R@K} is Recall@K
-    (high is good). \textbf{Med} {\it r} is the median rank (low is
-    good). \textbf{SSI} is Structural Similarity Index, which is between $-1$ and $1$ (high is good).
+    (higher is better). \textbf{Med} {\it r} is the median rank (lower is better). 
+    \textbf{SSI} is Structural Similarity Index, which is between $-1$ and $1$ 
+    (higher is better).
 }
@@
 
-Because the alignDRAW model tends to output slightly blurry samples, likely due to underfitting on the training data, we augmented the model with a sharpening post-processing step in which GAN generated edges which were added to the alignDRAW samples.
+Because the alignDRAW model tends to output slightly blurry samples, we augmented the model with a sharpening post-processing step in which GAN generated edges which were added to the alignDRAW samples.
 Unfortunately, this is not an ideal solution due to the fact that the whole model was not trained in an end-to-end fashion.

### 2015-11-05 supp.tex
a

@@
 The model was trained using \textit{RMSprop} with an initial learning rate of $0.001$. 
-The norm of the gradients was clipped at a maximum of $10.0$ during training to deal with exploding gradients.
-We used a vocabulary size of $K = 25323$ and $K = 22$ for Microsoft COCO and MNIST with Captions respectively.
+For the Microsoft COCO task, we trained our model for $18$ epochs. The learning rate was 
+reduced to $0.0001$ after $11$ epochs.
+\FIXME{For the MNIST with Captions task, the model was trained for ??? epochs and the 
+learnign rate was reduced to ?? after ?? epochs}.   
+The norm of the gradients was clipped at $10$ during training 
+to avoid the exploding gradients problem.
+
+We used a vocabulary size 
+of $K = 25323$ and $K = 22$ for the Microsoft COCO and MNIST with Captions datasets respectively. 
 All capital letters in the words were converted to small letters as a preprocessing step. 
-For all tasks, $\overrightarrow{h}^{lang}_{n}$ and $\overleftarrow{h}^{lang}_{n}$ in the language model had $128$ units.
-The parameters in the \textit{align} operator had a dimensionality of $l = 512$, $\vv \in \mathbb{R}^{512}$, $U \in \mathbb{R}^{512 \times 256}$, $W \in \mathbb{R}^{512 \times n^{gen}}$ and $b \in \mathbb{R}^{512}$.
-For the Microsoft COCO task, the alignDRAW model was trained for $18$ epochs and the learning rate was reduced by the factor of $0.1$ after $11$ epochs.
-The architectural configurations of alignDRAW models are shown on Table~\ref{tab:drawhyper}.
+For all tasks, the hidden states $\overrightarrow{h}^{lang}_{i}$ and $\overleftarrow{h}^{lang}_{i}$ 
+in the language model had $128$ units. Hence the dimensionality of the 
+concatenated state of the Bidirectional LSTM
+$\hlang_{i} = [\overrightarrow{h}^{lang}_{i}, \overleftarrow{h}^{lang}_{i}]$ was~256. 
+The parameters in the \textit{align} operator (\Eqref{eq:align2}) 
+had a dimensionality of $l = 512$, so that 
+$\vv \in \mathbb{R}^{512}$, $U \in \mathbb{R}^{512 \times 256}$, $W \in \mathbb{R}^{512 \times n^{gen}}$ and $b \in \mathbb{R}^{512}$. 
+The architectural configurations of the alignDRAW models are shown in Table~\ref{tab:drawhyper}.
 
@@
 \hline
-\multicolumn{7}{c}{\textbf{alignDRAW Model}} \\
 \hline
-Task & \#glimpses & Inference \#$h$ & Generative \#$h$ & \#$z$ & Read Size & Write Size\\
+ & & \multicolumn{3}{c}{\textbf{alignDRAW Model}}  & &  \\
 \hline
-Microsoft COCO & 32 & 550 & 550 & 275 & 9 & 9\\
-MNIST Captions & 32 & 300 & 300 & 150 & 8 & 8\\
+Task & \#glimpses & Inference  & Generative  & Latent & Read Size & Write Size\\
+     & \ T & Dim of $h^{infer}$ & Dim of $h^{gen}$  & Dim of $Z$  & $p$ & $p$\\
+\hline
+MS COCO & 32 & 550 & 550 & 275 & 9 & 9\\
+MNIST   & 32 & 300 & 300 & 150 & 8 & 8\\
 \end{tabulary}
@@
 \subsection*{Evaluation}
-
-Table~\ref{tab:nll} shows the estimated variational lower bound on the average train/validation/test 
+Table~\ref{tab:nll} shows the estimated variational lower bounds on the average train/validation/test 
 log-probabilities.
 Note that the alignDRAW model does not suffer much from overfitting. 
-The results worsened dramatically after sharpening test images.
+The results substantially worsen after sharpening test images.
 
@@
 \label{sec:post_processing}
-
+\vspace{-0.1in}
+Some examples of generated images before (top row) and after (bottom row) sharpening images 
+using an adversarial network trained on residuals of a Laplacian 
+pyramid conditioned on the skipthought vectors of the captions.
 \begin{figure}[!h]
@@
 \begin{center}
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-blurry.png}}\\
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-blurry.png}}\\
 \vspace{-0.45cm}
@@
 \subfloat[%A stop sign is flying in blue skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
 %
 \subfloat[%A herd of elephants flying in the blue skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
 %
 \subfloat[%A toilet seat sits open in the grass field.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
 %
 \subfloat[%A person skiing on sand clad vast desert.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\\
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\\
 \vspace{-0.25cm}
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-blurry.png}}\\
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-blurry.png}}\\
 \vspace{-0.45cm}
@@
 \subfloat[%A yellow school bus parked in a parking lot.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
 %
 \subfloat[%A red school bus parked in a parking lot.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
 %
 \subfloat[%A green school bus parked in a parking lot.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
 %
 \subfloat[%A blue school bus parked in a parking lot.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\\
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\\
 \vspace{-0.25cm}
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-vintage-photo-of-a-cat-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-blurry.png}}\\
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-vintage-photo-of-a-dog-blurry.png}}\\
 \vspace{-0.45cm}
@@
 \subfloat[%The decadent chocolate desert is on the table.
-]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
 %
 \subfloat[%A bowl of bananas is on the table.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
 %
 \subfloat[%A vintage photo of a cat.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
 %
 \subfloat[%A vintage photo of a dog.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\\
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\\
 \vspace{-0.25cm}
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-blurry.png}}\\
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-blurry.png}}\\
 \vspace{-0.45cm}
@@
 \subfloat[%A very large commercial plane flying in blue skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
 %
 \subfloat[%A very large commercial plane flying in rainy skies.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
 %
 \subfloat[%A herd of elephants walking across a dry grass field.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
 %
 \subfloat[%A herd of elephants walking across a green grass field.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\\
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\\
 \vspace{-0.25cm}
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-blurry.png}}\quad
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-blurry.png}}\quad
 %
-\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-blurry.png}}\\
+\subfloat[]{\includegraphics[width=0.21\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-blurry.png}}\\
 \vspace{-0.45cm}
@@
 \subfloat[%A rider on a blue motorcycle in the desert.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
 %
 \subfloat[%A rider on a blue motorcycle in the forest.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
 %
 \subfloat[%A surfer, a woman, and a child walk on the beach.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
 %
 \subfloat[%A surfer, a woman, and a child walk on the sun.
-]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\\
+]{\includegraphics[width=0.21\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\\
 \end{center}
 %
-\caption{\small Effect of sharpening images previously shown in the publication.}
+\caption{\small Effect of sharpening images.}
 %\vspace{-0.1in}

### 2015-11-05 iclr-paper-new.tex
a

@@
 Training details, hyperparameter settings,
-and the overall model architecture are specified in Appendix. 
+and the overall model architecture are specified in Appendix B. 
 We will also release our code with pretrained models.
@@
 estimate the variational lower bound accordingly.
-%Additionally, we noticed that sampling from the mean of the uniform distribution generated samples with less noise than if we had otherwise sampled from the uniform distribution itself.
 
@@
 \vspace{-0.05in}
-Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with at least 5 captions.
+Microsoft COCO \citep{mscoco} is a very large dataset of \FIXME{how many images exactly}  $83,000$ images, each annotated with at least 5 captions.
 The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model
@@
 \vspace{-0.05in}
-The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc., and then use that knowledge to generate relevant images.
+The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, and then use that knowledge to generate relevant images.
 To examine the understanding of our model, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.
@@
 Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.
+Some examples of generated images before and after sharpening images are shown in Appendix~C.
 

### 2015-11-05 iclr-paper-new.tex
new changes

--- 2015-11-05 iclr-paper-new.tex
@@
 
-\title{Generating Images From Captions\\ With Attention}
+\title{Generating Images from Captions\\ with Attention}
 
@@
 To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest
-Euclidean distance in the top convolutional feature-space of a VGG-like model~\citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
+Euclidean distance in the last fully-connected feature space of a VGG-like model~\citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
 The shortlist of test images used for evaluation can be downloaded from \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.

### 2015-11-05 supp.tex
readded data augmentation description
new changes

@@
 reduced to $0.0001$ after $11$ epochs.
-\FIXME{For the MNIST with Captions task, the model was trained for ??? epochs and the 
-learnign rate was reduced to ?? after ?? epochs}.   
+For the MNIST with Captions task, the model was trained for 150 epochs and the 
+learning rate was reduced to $0.0001$ after 110 epochs. During each epoch, randomly created $10000$ training samples were used for training.
 The norm of the gradients was clipped at $10$ during training 
@@
 \hline
-\hline
+%\hline
  & & \multicolumn{3}{c}{\textbf{alignDRAW Model}}  & &  \\

### 2015-11-06 iclr-paper-new.tex
changed task of in itself to task in itself

@@
 \vspace{-0.05in}
-Quantitatively evaluating generative models remains a challenging task in of itself as each method of evaluation suffers from its own specific drawbacks.
+Quantitatively evaluating generative models remains a challenging task in itself as each method of evaluation suffers from its own specific drawbacks.
 Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.

### 2015-11-08 iclr-paper-new.tex
highlighted changes

--- 2015-11-06 iclr-paper-new.tex
@@
 \vspace{-0.05in}
-Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with at least 5 captions.
+Microsoft COCO \citep{mscoco} is a very large dataset of \FIXME{$82,783$ images}, each annotated with at least 5 captions.
 The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model
@@
 \vspace{-0.05in}
-Quantitatively evaluating generative models remains a challenging task in itself as each method of evaluation suffers from its own specific drawbacks.
+Quantitatively evaluating generative models remains \FIXME{a challenging task in itself} as each method of evaluation suffers from its own specific drawbacks.
 Compared to reporting classification accuracies in discriminative models, the measures defining generative models are intractable most of the times and might not correctly define the real quality of the model.
@@
 To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest
-Euclidean distance in the last fully-connected feature space of a VGG-like model~\citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
+Euclidean distance in the \FIXME{last fully-connected feature space} of a VGG-like model~\citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}.
 The shortlist of test images used for evaluation can be downloaded from \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}.

### 2015-11-08 supp.tex
highlighted changes

@@
 reduced to $0.0001$ after $11$ epochs.
-For the MNIST with Captions task, the model was trained for 150 epochs and the 
-learning rate was reduced to $0.0001$ after 110 epochs. During each epoch, randomly created $10000$ training samples were used for training.
+\FIXME{For the MNIST with Captions task, the model was trained for 150 epochs and the 
+learning rate was reduced to $0.0001$ after 110 epochs. During each epoch, randomly created $10,000$ training samples were used for training.}
 The norm of the gradients was clipped at $10$ during training 

### 2015-11-09 iclr-paper-new.tex

@@
 \vspace{-0.05in}
-Microsoft COCO \citep{mscoco} is a very large dataset of \FIXME{$82,783$ images}, each annotated with at least 5 captions.
+Microsoft COCO \citep{mscoco} is a large dataset of $82,783$ images, each annotated with at least 5 captions.
 The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model

### 2015-11-09 iclr-paper-new.tex
a

--- 2015-11-06 iclr-paper-new.tex
@@
 
-%There are two primary directions in learning a generative model of image and text.  
 There are numerous ways to learn a generative model over both image and text modalities. One approach is to learn a generative model of text conditioned on images, known as caption generation
-%A significant amount of recent work has focused on generating captions from images
-\citep{karpathy_captions,vinyals_captions,xu_captions}.
+\citep{kiros_icml14,karpathy_captions,vinyals_captions,xu_captions}.
 These models take an image descriptor and generate unstructured texts using a recurrent decoder.
@@
 taking textual descriptions as input and using them to generate relevant images.
-%By contrast, learning a generative model for image and text may also be studied by generating images correctly interpreting the text description. 
 Generating high dimensional realistic images from their descriptions 
-%is a more difficult approach that 
 combines the two challenging components of language modelling and image generation, and can be considered to be more difficult than caption generation.
-%Namely, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate pixel intensities of the image.
 
@@
 \caption{\small 
-%Examples of novel scene compositions. Best viewed in colour.
 Examples of generated images based on captions that describe 
@@
 While most of the recent success has been achieved by discriminative models, generative models have not yet enjoyed the same level of success.
-Most of the previous work in generative models has focused on variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}.
-While these models are very powerful, each iteration of training requires a computationally costly step of MCMC to approximate an intractable normalization constant, making it difficult to scale them to large datasets.
+Most of the previous work in generative models has focused on variants of Boltzmann Machines \citep{smolensky_rbm,russ_dbm} and Deep Belief Networks~\citep{hinton_dbn}.
+While these models are very powerful, each iteration of training requires a computationally costly step of MCMC to approximate derivatives of an
+intractable partition function (normalization constant), making it difficult to scale them to large datasets.
 
@@
 The encoder is used to approximate a posterior distribution and the decoder is used to stochastically reconstruct the data from latent variables.
-%The model's efficient inference and learning procedures allow it to scale to large datasets. 
-\cite{gregor_draw} introduced the Deep Recurrent Attention Writer (DRAW), extending the VAE approach by incorporating a novel differentiable attention mechanism.
+\cite{gregor_draw} further introduced the Deep Recurrent Attention Writer (DRAW), extending the VAE approach by incorporating a novel differentiable attention mechanism.
 
-Generative Adversarial Networks (GANs) \citep{goodfellow_gan} are another type of generative models that use noise-contrastive estimation \citep{gutmann_nce} to avoid calculating an intractable normalization constant.
+Generative Adversarial Networks (GANs) \citep{goodfellow_gan} are another type of generative models that use noise-contrastive estimation \citep{gutmann_nce} to avoid calculating an intractable partition function.
 The model consists of a generator that generates samples using a uniform distribution and a discriminator that discriminates between real and generated images.
-%Both networks can be seen as playing a game against each other, where the generator tries to produce samples that look real and the discriminator tries not to be fooled by the generator. 
 Recently, \cite{denton_lapgan} have scaled those models by training conditional GANs at each level of a Laplacian pyramid of images.
-%While their model has generated sharp looking samples, the generated images have lacked a clear structure. 
-%Compared to other mentioned generative models, GANs are more unstable and are harder to train.
 
-While all of the previous work has been focused on unconditional models or models conditioned on labels, %to the best of our knowledge 
+While many of the previous approaches have focused on unconditional models or models conditioned on labels, 
 in this paper we develop a generative model of images conditioned on captions.
@@
 The model can be viewed as a part of the sequence-to-sequence framework \citep{ilya_mt,cho_mt,nitish_video}.
-%Our proposed model can be viewed as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$.
-Let $\icaption$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $\oimage$ be the image corresponding to that caption.
 
@@
 where $K$ is the size of the vocabulary and $N$ is the length of the sequence.
-%and $\oimage$ be the output image. 
 We obtain the caption sentence representation by first 
@@
 \vspace{-0.1in}
-\caption{\small alignDRAW model for generating images by learning an alignment between the input captions
+\caption{\small AlignDRAW model for generating images by learning an alignment between the input captions
 and generating canvas. The caption is encoded using 
 the Bidirectional RNN (left). The generative RNN takes a latent sequence $z_{1:T}$ sampled from the 
-prior along with dynamic caption representation $s_{1:T}$ to generate the canvas matrix $c_T$, which is then
+prior along with the dynamic caption representation $s_{1:T}$ to generate the canvas matrix $c_T$, which is then
 used to generate the final image $\oimage$ (right).
@@
 representation $\hlangall$ at each step, as shown in \Figref{fig:figmodel}.
-The conditional DRAW network is a stochastic recurrent neural network (LSTM)
+The conditional DRAW network is a stochastic recurrent neural network
 that consists of a sequence of latent variables $\Lat_t \in \mathbb{R}^D$,
@@
 The corresponding alignment probability $\alpha_{k}^{t}$ for the $k^{th}$ word in the caption is obtained using the caption representation $\hlangall$ and the current hidden state of the generative model $\hdec_{t-1}$:
-%first calculating energies using a $\tanh$ perceptron scaled by vector $\vv$
-%and then normalizing them: %, as in \citep{bahdanau_mt}:
 \beqa
@@
 combines it with the input from both the latent sample $z_t$ and the sentence representation $s_t$.
-%combines it with the input from both the latent sample $z_t$ and 
-%sentence representation $s_t$, where
-%$[v_1, v_2]$ denotes the concatenation of $v_1$ and $v_2$ into the single vector.
 
@@
 \label{sec:learning}
-%The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. 
 The model is trained to maximize a variational lower bound $\loss$ 
@@
 \eeqa
-%The posterior inference over the latent variables $\Lat_{1:T}$ is approximated using the inference RNN. 
 Similar to the DRAW model, the inference recurrent network 
-produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from image using two arrays of 1D Gaussian filters (inverse of $\writeop$ from section \ref{sec:image_model}), over the training image at each step.
+produces an approximate posterior $Q(\Lat_{1:T}\given\oimage,\icaption)$ via a $\readop$ operator, which reads a patch from an input image $\oimage$ using two arrays of 1D Gaussian filters (inverse of $\writeop$ from section \ref{sec:image_model}) at each time-step $t$.
 Specifically,
@@
 \end{align}
-%The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$. 
-%The reconstruction loss $\rloss$ equals to $\frac{1}{L}\sum_{l=1}^{L}(log\,p(x_{t}|y,z)$ where $L$ is the number of samples used during training, which was set to $1$ in our experiments.
-%The latent loss $\lloss$ is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term.
-%Since the prior distribution at time $t$ is dependent on the sufficient statistics of the prior distribution at time $t-1$.
-Namely, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$.
-%\begin{align}
-%\mu_{t}^{prior} &= tanh(W_{\mu}\hdec_{t-1})\\
-%\sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
-%\end{align}
 where $\hat{\oimage}$ is the \textit{error} image and $\henc_0$ is initialized to the learned bias $b$. 
@@
 generative decoder $\hdec_{t-1}$, which depends on the latent sample history $z_{1:t-1}$ and 
-dynamic sentence representation $s_t$ (see \Eqref{eq:decoder}). 
+dynamic sentence representation $s_{t-1}$ (see \Eqref{eq:decoder}). 
 Hence, the approximate posterior $Q$ will depend on the input image $\oimage$,
@@
 \captionsetup[subfigure]{labelformat=empty}
-\vspace{-0.2in}
+\vspace{-0.3in}
 \begin{center}
@@
 \label{fig:genimages3}
-\vspace{-0.1in}
+\vspace{-0.2in}
 \end{figure}
@@
 \vspace{-0.05in}
-Microsoft COCO \citep{mscoco} is a very large dataset of \FIXME{how many images exactly}  $83,000$ images, each annotated with at least 5 captions.
+Microsoft COCO \citep{mscoco} is a large dataset containing $82,783$ images, each annotated with at least 5 captions.
 The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model
-%conditioned on a caption 
 very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
 The images were resized to $32 \times 32$ pixels for consistency with other tiny image datasets \citep{krizhevsky_cifar}.
-In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more
-generated images, visit \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.
+In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models.\footnote{To see more
+generated images, visit \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}
+Appendix A further reports some additional experiments using the MNIST dataset.
 
@@
 \label{fig:diffmodels}
-\vspace{-0.1in}
+\vspace{-0.2in}
 \end{figure}
@@
 While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with least squares cost function have a watercolor effect on the images.
-We also found that the quality of generated samples was very similar among different variants of conditional DRAW models.
+We found that the quality of generated samples was similar among different variants of conditional DRAW models.
 
@@
 Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.
-Some examples of generated images before and after sharpening images are shown in Appendix~C.
+Some examples of generated images before and after sharpening are shown in Appendix~C.
 
@@
 As you can see on Table~\ref{tab:results}, SSI scores achieved by variational models were higher compared to SSI score achieved by LAPGAN.
-%As for another metric of image similarity, we calculated euclidian distance between generated images and real images in the feature space of VGG-like model trained on CIFAR dataset.
-This is not necessarily a good metric, since the statistics of generated images were vastly different from statistics of training images.
-Nevertheless, the results are reported on Table~\ref{tab:results}.
 
-\section{Discussion}
-\vspace{-0.1in}
 \begin{table}[!t]
+\vspace{-0.3in}
 \begin{center}
@@
 \end{table}
+
+\section{Discussion}
+\vspace{-0.1in}
 In this paper, we demonstrated that the alignDRAW model, a combination of a recurrent variational autoencoder with an alignment model over words, succeeded in generating images that correspond to a given input caption.
@@
 We would like to thank developers of Theano \citep{theano}, the authors of \citep{denton_lapgan} for open sourcing their code, and Ryan Kiros and Nitish Srivastava for helpful discussions. 
-%We will be releasing our code with pretrained models.
 }

### 2015-11-09 supp.tex
a

--- 2015-11-05 supp.tex
@@
 \includegraphics[width=0.99\textwidth]{figures/new/mnist/test3.pdf}\quad
-%
 \end{center}
@@
 For the MNIST with Captions task, the model was trained for 150 epochs and the 
-learning rate was reduced to $0.0001$ after 110 epochs. During each epoch, randomly created $10000$ training samples were used for training.
+learning rate was reduced to $0.0001$ after 110 epochs. During each epoch, 
+randomly created $10,000$ training samples were used for learning. 
 The norm of the gradients was clipped at $10$ during training 
@@
 \hline
-%\hline
  & & \multicolumn{3}{c}{\textbf{alignDRAW Model}}  & &  \\
@@
 \end{center}
-%\vspace{-0.3cm}
 \end{table}
@@
 \hline
-Model & Train LL & Valid LL & Test LL & Test LL (after sharpening)\\
+Model & Train & Validation  & Test  & Test  (after sharpening)\\
 \hline
@@
 \caption{\small Effect of sharpening images.}
-%\vspace{-0.1in}
 \end{figure}

### 2015-11-09 supp.tex
fixed spelling

@@
 One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image.
-Two digits were either placed horizontally or vertically in nonoverlapping fashion.
+Two digits were either placed horizontally or vertically in non-overlapping fashion.
 The corresponding artificial captions specified the
-identity of each digit along with their relative positions, e.g. ``The digit two is on the top  
-of the digit zero'', or ``The digit seven is at the bottom left of the image''.
+identity of each digit along with their relative positions, e.g. ``The digit three is at the top  
+of the digit one'', or ``The digit seven is at the bottom left of the image''.
 

### 2015-11-09 iclr-paper-new.tex
a

@@
 \vspace{-0.05in}
-Microsoft COCO \citep{mscoco} is a large dataset containing $82,783$ images, each annotated with at least 5 captions.
+Microsoft COCO \citep{mscoco} is a large dataset containing 82,783 images, each annotated with at least 5 captions.
 The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model
@@
 We also found that the lower bound on the test log-probabilities decreased for sharpened images, and that sharpening considerably hurt the retrieval results.
-Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.
+Since sharpening changes the statistics of images, the estimated log-probabilities of image pixels is not necessarily a good metric.
 Some examples of generated images before and after sharpening are shown in Appendix~C.
@@
 {\small
-{\bf Acknowledgements:}
+{\bf Acknowledgments:}
 This work was supported by Samsung and IARPA, Raytheon BBN Contract No. D11PC20071.

### 2015-12-28 rebuttal.txt
adding response to the reviewers

Dear reviewers,
We thank you for very constructive reviews.

----
Reviewer 1

Q. It would also be nice to know how their align function compares to the standard content-based attention.

A. Not sure what to write to him/her. Not clear what reviewer means by "standard content-based attention".

----
Reviewer 2

Q. The results of "picture of a cat" vs. "picture of a dog" highlight my concern - I see no noticeable difference between the two sets of generated images.

A. Yes this example was meant to highlight the limitation of our model.

Q. The use of SSI and retrieval for comparison of the generative models is interesting, but I worry that the regime in which these models operate (SSI 0-0.2) is not a place where SSI is really useful or particularly interpretable - is a relative gain of 0.08 to 0.156 immense, or "in the noise". What is the standard deviation of the measurements for each row?

A. We agree that SSI results may not be particularly interpretable, however it was one of the only good ways to quantitatively compare performances of LAPGAN and our model that we could think of. The standard deviation for each model is around 0.1

Good quantitative evaluation of generative models still remains an open problem that was recently adressed in Theis et al. 2015.

Q. Given the recent (e.g. under submission in this conference!) improvements to generative models, I would expect this work will be extended to much higher quality in the near future. This paper lays solid groundwork for that, and is interesting in its own right, though the generated image quality is the main a limiting factor in giving a higher score.

A. As for the recent work on Generative Adversarial Networks (GANs) by Radford et al. 2015, their model was trained on a very narrow domain of images, for example millions of images of just bedrooms from LSUN dataset, where it is not clear whether the model is overfitting or not. In comparison, Microsoft COCO is much smaller and much more diverse dataset. Conditioning the model on captions provides a clear qualitative way to see whether the model has some notion of generalization. Being able to generalize to weird and novel captions not present in dataset is in itself a very challenging problem. At the same time, we hope that other researchers will try to come up with better generative models for this task.

----
Reviewer 3

Q. In the experiments, the size of the image is quite small (i.e 32 by 32). This means that the groundtruth images have already been blurred before the training of the model. This might be one of the reasons why the generated images are blurry. Can you train your model on images with common sizes (e,g, 100 by 100 or even 224 by 224)? Is there any obstacle to overcome for the proposed model to handle larger size of images than 32 by 32?

A. The images were resized to 32x32 dimension using nearest neighbor interpolation which doesn't make images blurry. The image size is not a limiting factor of the model, however it takes a bit more than a week to train the model on 32x32 images using Nvidia Titan GPUs due to large size of the model (both in depth and in time). Due to time constraints, we made the dimension of images in the dataset consistent with other popular tiny images datasets (CIFAR-10 and CIFAR-100).

Q. How does the model perform without the post-processing GAN model quantitatively, compared to other models with GAN shown in Table 1?

The results in Table 1 are shown without the post-processing GAN step. With post-processing step, where reconstruction error is the difference between sharpened images and real images instead of the difference between blurry images and real images, the ranking results become much worse. Median rank drops from 31 to 50, where ranking essentially becomes random. The bad quality of pixel-wise error motivated us to try better image similarity metrics such as SSI. In comparison SSI drops only by 0.02 points after sharpening images.

Q. It would be great if GAN post-processing step is merged into the model shown in Figure 2 so that the whole model can be trained and tested in an end-to-end way.

A. We tried combining VAE and GAN cost functions so that the model could be trained in an end-to-end fashion on CIFAR-10 and MS COCO datasets. We didn't have any success in that because the model started emphasizing GAN objective function and the samples started looking more like the samples from original GAN model from Goodfellow et al. 2014. Scaling both objectives and doing other tricks didn't help either. Note that Lotter et al. 2015 (Unsupervised Learning of Visual Structure using Predictive Generative Network submitted to this conference) have also tried combining both objectives in the similar way and had a limited success.

We also tried experimenting with other types of reconstruction error, for example in Fourier domain, all with no major success. Generating sharp and very realistic images from autoencoder without overfitting is a very hard problem and is part of an ongoing research (see Mathieu et al. 2015 Deep multi-scale video prediction beyond mean square error; Ridgeway et al. 2015 Learning to generate images with perceptual similarity metrics all submitted to this conference).

Q. In Eqn (15), only one sample is used for parameter learning. Will more samples make a difference in the learning process?

A. We would expect that using more samples would improve the lower bound of log likelihood as well as the quality of generated samples, which was shown recently in Burda et al. 2015 Importance Weighted Autoencoders (also submitted to this conference). However, using more samples would increase the training time and in our case training already takes quite a while on Nvidia Titan GPUs.

### 2015-12-30 rebuttal.txt
adding response to the reviewers
a

@@
-Dear reviewers,
-We thank you for very constructive reviews.
+We thank all reviewer for their valuable reviews.
 
-----
-Reviewer 1
-
-Q. It would also be nice to know how their align function compares to the standard content-based attention.
-
-A. Not sure what to write to him/her. Not clear what reviewer means by "standard content-based attention".
-
-----
 Reviewer 2
 
-Q. The results of "picture of a cat" vs. "picture of a dog" highlight my concern - I see no noticeable difference between the two sets of generated images.
+Regarding the results of "picture of a cat" vs. "picture of a dog" with no noticeable difference:
+Indeed, when swapping two objects that are visually similar, it is difficult to discriminate solely from the generated samples.
+This example was meant to highlight a limitation of our model in that it has difficulty generating the fine-grained details of objects.
 
-A. Yes this example was meant to highlight the limitation of our model.
+The use of SSI and retrieval for comparison:
+We agree that SSI results may not be particularly interpretable, however it was one of the only good ways to quantitatively compare performances of LAPGAN and our model that we could think of.
+We note that good quantitative evaluation of generative models still remains an open problem as was recently discussed in Theis et al.
+2015.
 
-Q.
-The use of SSI and retrieval for comparison of the generative models is interesting, but I worry that the regime in which these models operate (SSI 0-0.2) is not a place where SSI is really useful or particularly interpretable - is a relative gain of 0.08 to 0.156 immense, or "in the noise".
-What is the standard deviation of the measurements for each row?
+The standard deviation of SSI scores for variational models is 0.1 and the standard deviation for LAPGAN is 0.08. We will add these numbers to the paper.
 
-A.
-We agree that SSI results may not be particularly interpretable, however it was one of the only good ways to quantitatively compare performances of LAPGAN and our model that we could think of.
-The standard deviation for each model is around 0.1
+On the the recent improvements to generative models and the generated image quality remaining the main a limiting factor:
+We agree that there has been a lot of recent work on improving generative models. However, even today, generating sharp and globally consistent images of arbitrary scenes remains an open problem.
 
-Good quantitative evaluation of generative models still remains an open problem that was recently adressed in Theis et al. 2015.
+For example, in the recent work on Generative Adversarial Networks (GANs) by Radford et al. 2015, 
+they train their model on a very narrow domain of images -- millions of images of just 
+bedroom scenes from LSUN bedroom dataset. In this setting, it is may not clear how much overfitting takes place. 
+In comparison, Microsoft COCO is a far more diverse dataset. Conditioning the model on 
+novel captions provides a clear qualitative way to evaluate whether the model has some notion of generalization. 
+Being able to generalize to novel captions not present in dataset, such as 
+"A herd of elephants flying
+in the blue skies" is in itself a very challenging problem. 
+At the same time, we also hope that other researchers will try to come up with better generative models for this task.
 
-Q.
-Given the recent (e.g.
-under submission in this conference!) improvements to generative models, I would expect this work will be extended to much higher quality in the near future.
-This paper lays solid groundwork for that, and is interesting in its own right, though the generated image quality is the main a limiting factor in giving a higher score.
-
-A.
-As for the recent work on Generative Adversarial Networks (GANs) by Radford et al.
-2015, their model was trained on a very narrow domain of images, for example millions of images of just bedrooms from LSUN dataset, where it is not clear whether the model is overfitting or not.
-In comparison, Microsoft COCO is much smaller and much more diverse dataset.
-Conditioning the model on captions provides a clear qualitative way to see whether the model has some notion of generalization.
-Being able to generalize to weird and novel captions not present in dataset is in itself a very challenging problem.
-At the same time, we hope that other researchers will try to come up with better generative models for this task.
-
-----
 Reviewer 3
 
-Q.
-In the experiments, the size of the image is quite small (i.e 32 by 32).
-This means that the groundtruth images have already been blurred before the training of the model.
-This might be one of the reasons why the generated images are blurry.
-Can you train your model on images with common sizes (e,g, 100 by 100 or even 224 by 224)?
-Is there any obstacle to overcome for the proposed model to handle larger size of images than 32 by 32?
+In regard to small size images and training images with more common sizes: 
+The images were resized to 32x32 dimension using bi-cubic interpolation and the resulting images do not look 
+blurry. The image size is not a limiting factor of the model, however, it takes about a week 
+to train our model on 32x32 images on a single Nvidia TitanX GPU due to large size of the model 
+(both in depth and in time). Due to time constraints, we made the dimension of images in the dataset consistent with other popular tiny images datasets (CIFAR-10 and CIFAR-100). 
+We also believe that scaling our model up to modeling 64x64 or 100x100 images 
+will not necessarily significantly improve the quality of generated images.
 
-A.
-The images were resized to 32x32 dimension using nearest neighbor interpolation which doesn't make images blurry.
-The image size is not a limiting factor of the model, however it takes a bit more than a week to train the model on 32x32 images using Nvidia Titan GPUs due to large size of the model (both in depth and in time).
-Due to time constraints, we made the dimension of images in the dataset consistent with other popular tiny images datasets (CIFAR-10 and CIFAR-100).
+Model performance without the post-processing GAN step:   
+The results in Table 1 are shown without the post-processing GAN step.
+We found that with the post-processing GAN step, the ranking results became much worse.
+Median rank drops from 31 to 50, in which
+case the ranking essentially becomes random. The bad quality of pixel-wise error motivated 
+us to try better image similarity metrics such as SSI. In comparison, SSI drops only by 0.02 points 
+after sharpening images.
 
-Q. How does the model perform without the post-processing GAN model quantitatively, compared to other models with GAN shown in Table 1?
+In regard to merging GAN post-processing step into the whole model:
+We tried combining VAE and GAN cost functions so that the model could be trained in an end-to-end fashion on CIFAR-10 and MS COCO datasets. 
+So far, we did not have much success in that because the model started emphasizing GAN objective function and the samples started to look more like the samples from the GAN model.
+Scaling both objectives and doing other tricks
+did not help either.
+Note that Lotter et al.
+2015 (Unsupervised Learning of Visual Structure using Predictive Generative Network submitted to this conference) have also tried combining both objectives in a similar way and had a limited success.
 
-The results in Table 1 are shown without the post-processing GAN step.
-With post-processing step, where reconstruction error is the difference between sharpened images and real images instead of the difference between blurry images and real images, the ranking results become much worse.
-Median rank drops from 31 to 50, where ranking essentially becomes random.
-The bad quality of pixel-wise error motivated us to try better image similarity metrics such as SSI.
-In comparison SSI drops only by 0.02 points after sharpening images.
+We also tried experimenting with other types of reconstruction error, for example in Fourier domain, all with no major success so far.
+Generating sharp and very realistic images remains challenging and is part of an ongoing research (see Mathieu et al.; Ridgeway et al.
+2015 all submitted to this conference).
 
-Q. It would be great if GAN post-processing step is merged into the model shown in Figure 2 so that the whole model can be trained and tested in an end-to-end way.
-
-A.
-We tried combining VAE and GAN cost functions so that the model could be trained in an end-to-end fashion on CIFAR-10 and MS COCO datasets.
-We didn't have any success in that because the model started emphasizing GAN objective function and the samples started looking more like the samples from original GAN model from Goodfellow et al.
-2014.
-Scaling both objectives and doing other tricks didn't help either.
-Note that Lotter et al.
-2015 (Unsupervised Learning of Visual Structure using Predictive Generative Network submitted to this conference) have also tried combining both objectives in the similar way and had a limited success.
-
-We also tried experimenting with other types of reconstruction error, for example in Fourier domain, all with no major success.
-Generating sharp and very realistic images from autoencoder without overfitting is a very hard problem and is part of an ongoing research (see Mathieu et al.
-2015 Deep multi-scale video prediction beyond mean square error; Ridgeway et al.
-2015 Learning to generate images with perceptual similarity metrics all submitted to this conference).
-
-Q. In Eqn (15), only one sample is used for parameter learning. Will more samples make a difference in the learning process?
-
-A.
-We would expect that using more samples would improve the lower bound of log likelihood as well as the quality of generated samples, which was shown recently in Burda et al.
-2015 Importance Weighted Autoencoders (also submitted to this conference).
-However, using more samples would increase the training time and in our case training already takes quite a while on Nvidia Titan GPUs.
+Eq (15): 1 vs. many samples:
+We would expect that using more samples could improve the lower bound on the log-likelihood 
+as well as the quality of generated samples, especially when using importance sampling as 
+was shown recently in Burda et al. 2015. However, using more samples would increase the training 
+time, which in our case already takes quite a while (using TitanX GPUs).

### 2015-12-30 rebuttal.txt
a

@@
 
-Reviewer 2
+Reviewer 1:
+Thank you for your review, but we were not quite sure what the reviewer meant by the standard content-based attention.
+In our paper, we compared the performance of our model to the DRAW model conditioned on captions
+without attention and the DRAW model conditioned on
+the skipthought vectors of Kiros et al., 2015.
+We generally found that the
+quality of generated samples was similar among different variants of conditional DRAW models.
+
+Reviewer 2:
 
@@
 
-Reviewer 3
+Reviewer 3:
 
@@
 blurry. The image size is not a limiting factor of the model, however, it takes about a week 
-to train our model on 32x32 images on a single Nvidia TitanX GPU due to large size of the model 
+to train our model on 32x32 images on a single Nvidia Titan GPU due to large size of the model 
 (both in depth and in time). Due to time constraints, we made the dimension of images in the dataset consistent with other popular tiny images datasets (CIFAR-10 and CIFAR-100). 
@@
 was shown recently in Burda et al. 2015. However, using more samples would increase the training 
-time, which in our case already takes quite a while (using TitanX GPUs).
+time, which in our case already takes quite a while (using Titan GPUs).

### 2015-12-31 rebuttal.txt
a

@@
-We thank all reviewer for their valuable reviews.
+We thank all reviewer for their valuable feedback.
 
 Reviewer 1:
-Thank you for your review, but we were not quite sure what the reviewer meant by the standard content-based attention.
+Thank you for your review.
+We were not quite sure what the reviewer meant by the standard content-based attention.
 In our paper, we compared the performance of our model to the DRAW model conditioned on captions
-without attention and the DRAW model conditioned on
-the skipthought vectors of Kiros et al., 2015.
-We generally found that the
-quality of generated samples was similar among different variants of conditional DRAW models.
+without attention and the DRAW model conditioned on the skipthought vectors of Kiros et al., 2015.
+We generally found that the quality of generated samples was similar among different variants of conditional DRAW models.
 
@@
 The use of SSI and retrieval for comparison:
-We agree that SSI results may not be particularly interpretable, however it was one of the only good ways to quantitatively compare performances of LAPGAN and our model that we could think of.
+We agree that SSI results may not be particularly interpretable.
+However, it was one of the only good ways to quantitatively compare performances of LAPGAN and our model that we could think of.
 We note that good quantitative evaluation of generative models still remains an open problem as was recently discussed in Theis et al.
@@
 
-On the the recent improvements to generative models and the generated image quality remaining the main a limiting factor:
+On the recent improvements to generative models and the quality of generated images:
 We agree that there has been a lot of recent work on improving generative models. However, even today, generating sharp and globally consistent images of arbitrary scenes remains an open problem.
 
-For example, in the recent work on Generative Adversarial Networks (GANs) by Radford et al. 2015, 
-they train their model on a very narrow domain of images -- millions of images of just 
+For example, in the recent work on Generative Adversarial Networks (GANs) by Radford et al. 2015, they train their model on a very narrow domain of images -- millions of images of just 
 bedroom scenes from LSUN bedroom dataset. In this setting, it is may not clear how much overfitting takes place. 
@@
 Being able to generalize to novel captions not present in dataset, such as 
-"A herd of elephants flying
-in the blue skies" is in itself a very challenging problem. 
+"A herd of elephants flying in the blue skies" is in itself a very challenging problem.
 At the same time, we also hope that other researchers will try to come up with better generative models for this task.
@@
 
-In regard to small size images and training images with more common sizes: 
-The images were resized to 32x32 dimension using bi-cubic interpolation and the resulting images do not look 
-blurry. The image size is not a limiting factor of the model, however, it takes about a week 
-to train our model on 32x32 images on a single Nvidia Titan GPU due to large size of the model 
+In regard to training on small size images:
+The images were resized to 32x32 using bi-cubic interpolation and the resulting images did not look blurry.
+The image size is not a limiting factor of the model.
+However, even on a high-end GPU it takes about a week
+to train our model on 32x32 images due to the large size of the model 
 (both in depth and in time). Due to time constraints, we made the dimension of images in the dataset consistent with other popular tiny images datasets (CIFAR-10 and CIFAR-100). 
@@
 We found that with the post-processing GAN step, the ranking results became much worse.
-Median rank drops from 31 to 50, in which
-case the ranking essentially becomes random. The bad quality of pixel-wise error motivated 
+Median rank dropped from 31 to 50, in which
+case the ranking essentially became random. The bad quality of pixel-wise error motivated 
 us to try better image similarity metrics such as SSI. In comparison, SSI drops only by 0.02 points 
@@
 We tried combining VAE and GAN cost functions so that the model could be trained in an end-to-end fashion on CIFAR-10 and MS COCO datasets. 
-So far, we did not have much success in that because the model started emphasizing GAN objective function and the samples started to look more like the samples from the GAN model.
+So far, we did not have much success because the model started emphasizing GAN objective function and the samples started to look more like the samples from the GAN model.
 Scaling both objectives and doing other tricks
@@
 
-We also tried experimenting with other types of reconstruction error, for example in Fourier domain, all with no major success so far.
+We have also tried experimenting with other types of reconstruction error, for example in Fourier domain, all with no major success so far.
 Generating sharp and very realistic images remains challenging and is part of an ongoing research (see Mathieu et al.; Ridgeway et al.
@@
 as well as the quality of generated samples, especially when using importance sampling as 
-was shown recently in Burda et al. 2015. However, using more samples would increase the training 
-time, which in our case already takes quite a while (using Titan GPUs).
+was recently demonstrated by Burda et al. 2015. However, using more samples would increase the training 
+time, which in our case already takes quite a while.

### 2016-01-04 rebuttal.txt
small changes. already submitted rebuttal to cmt

@@
-We thank all reviewer for their valuable feedback.
+We thank all reviewers for their valuable feedback.
 
 Reviewer 1:
+
 Thank you for your review.
-We were not quite sure what the reviewer meant by the standard content-based attention.
-In our paper, we compared the performance of our model to the DRAW model conditioned on captions
-without attention and the DRAW model conditioned on the skipthought vectors of Kiros et al., 2015.
+We were not quite sure what you meant by the standard content-based attention.
+In our paper, we compared the performance of our model to the DRAW model conditioned on captions without attention and the DRAW model conditioned on the skipthought vectors of Kiros et al., 2015.
 We generally found that the quality of generated samples was similar among different variants of conditional DRAW models.
@@
 Regarding the results of "picture of a cat" vs. "picture of a dog" with no noticeable difference:
+
 Indeed, when swapping two objects that are visually similar, it is difficult to discriminate solely from the generated samples.
@@
 The use of SSI and retrieval for comparison:
+
 We agree that SSI results may not be particularly interpretable.
@@
 2015.
-
 The standard deviation of SSI scores for variational models is 0.1 and the standard deviation for LAPGAN is 0.08. We will add these numbers to the paper.
@@
 On the recent improvements to generative models and the quality of generated images:
-We agree that there has been a lot of recent work on improving generative models. However, even today, generating sharp and globally consistent images of arbitrary scenes remains an open problem.
 
-For example, in the recent work on Generative Adversarial Networks (GANs) by Radford et al. 2015, they train their model on a very narrow domain of images -- millions of images of just 
-bedroom scenes from LSUN bedroom dataset. In this setting, it is may not clear how much overfitting takes place. 
-In comparison, Microsoft COCO is a far more diverse dataset. Conditioning the model on 
-novel captions provides a clear qualitative way to evaluate whether the model has some notion of generalization. 
-Being able to generalize to novel captions not present in dataset, such as 
-"A herd of elephants flying in the blue skies" is in itself a very challenging problem.
+We agree that there has been a lot of recent work on improving generative models.
+However, even today, generating sharp and globally consistent images of arbitrary scenes remains an open problem.
+For example, in the recent work on Generative Adversarial Networks (GANs) by Radford et al.
+2015, they train their model on a very narrow domain of images -- millions of images of just bedroom scenes from LSUN bedroom dataset.
+In this setting, it is not clear how much overfitting takes place.
+In comparison, Microsoft COCO is a far more diverse dataset.
+Conditioning the model on novel captions provides a clear qualitative way to evaluate whether the model has some notion of generalization.
+Being able to generalize to novel captions not present in dataset, such as "A herd of elephants flying in the blue skies" is in itself a very challenging problem.
 At the same time, we also hope that other researchers will try to come up with better generative models for this task.
@@
 In regard to training on small size images:
+
 The images were resized to 32x32 using bi-cubic interpolation and the resulting images did not look blurry.
 The image size is not a limiting factor of the model.
-However, even on a high-end GPU it takes about a week
-to train our model on 32x32 images due to the large size of the model 
-(both in depth and in time). Due to time constraints, we made the dimension of images in the dataset consistent with other popular tiny images datasets (CIFAR-10 and CIFAR-100). 
-We also believe that scaling our model up to modeling 64x64 or 100x100 images 
-will not necessarily significantly improve the quality of generated images.
+However, even on a high-end GPU it takes about a week to train our model on 32x32 images due to the large size of the model (both in depth and in time).
+Due to time constraints, we made the dimension of images in the dataset consistent with other popular tiny images datasets (CIFAR-10 and CIFAR-100).
+We also believe that scaling our model up to modeling 64x64 or 100x100 images will not necessarily significantly improve the quality of generated images.
 
 Model performance without the post-processing GAN step:
+
 The results in Table 1 are shown without the post-processing GAN step.
 We found that with the post-processing GAN step, the ranking results became much worse.
-Median rank dropped from 31 to 50, in which
-case the ranking essentially became random. The bad quality of pixel-wise error motivated 
-us to try better image similarity metrics such as SSI. In comparison, SSI drops only by 0.02 points 
-after sharpening images.
+Median rank dropped from 31 to 50, in which case the ranking essentially became random.
+The bad quality of pixel-wise error motivated us to try better image similarity metrics such as SSI.
+In comparison, SSI drops only by 0.02 points after sharpening images.
 
 In regard to merging GAN post-processing step into the whole model:
+
 We tried combining VAE and GAN cost functions so that the model could be trained in an end-to-end fashion on CIFAR-10 and MS COCO datasets.
-So far, we did not have much success because the model started emphasizing GAN objective function and the samples started to look more like the samples from the GAN model.
-Scaling both objectives and doing other tricks
-did not help either.
+So far, we did not have much success because the model started emphasizing GAN objective function and the samples started to look more like the samples from the GAN model shown in Goodfellow et.
+al.
+2014.
+Scaling both objectives and doing other tricks did not help either.
 Note that Lotter et al.
@@
 We have also tried experimenting with other types of reconstruction error, for example in Fourier domain, all with no major success so far.
-Generating sharp and very realistic images remains challenging and is part of an ongoing research (see Mathieu et al.; Ridgeway et al.
+Generating sharp and very realistic images without overfitting remains challenging and is part of an ongoing research (see Mathieu et al.; Ridgeway et al.
 2015 all submitted to this conference).
@@
 Eq (15): 1 vs. many samples:
-We would expect that using more samples could improve the lower bound on the log-likelihood 
-as well as the quality of generated samples, especially when using importance sampling as 
-was recently demonstrated by Burda et al. 2015. However, using more samples would increase the training 
-time, which in our case already takes quite a while.
+
+We would expect that using more samples could improve the lower bound on the log-likelihood as well as the quality of generated samples, especially when using importance sampling as was recently demonstrated by Burda et al.
+2015.
+However, using more samples would increase the training time, which in our case already takes quite a while.

### 2016-02-25 iclr-paper-new.tex
small changes for final revision

@@
 and the overall model architecture are specified in Appendix B. 
-We will also release our code with pretrained models.
+The code is available at \url{https://github.com/emansim/text2image}.
+%We will also release our code with pretrained models.
 
@@
 \hline
-LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
+LAPGAN & - & - & - & - & - & 0.08 $\pm$ 0.07 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
 \hline
-Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
-Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
-skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
-noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
-alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
+Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 $\pm$ 0.10  \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
+Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 $\pm$ 0.10 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
+skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 $\pm$ 0.11 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
+noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 $\pm$ 0.11 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
+alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 $\pm$ 0.11 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
 \end{tabulary}
